---
title: "Analyses for testing the association betwween isolation trajectories and outcomes - Three trajectories"
author: "Katherine N Thompson"
date: "19th Jan 2021"
output:
  word_document:
    toc: yes
    toc_depth: '2'
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
    number_sections: no
    highlight: monochrome
    theme: cerulean
code_folding: hide
---

outcomes_isolation_THREE_trajectories_KT

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      comment = NA,
                      prompt = FALSE,
                      cache = FALSE,
                      message = FALSE,
                      warning = FALSE)

options(bitmapType = 'quartz') # to render fonts better
```

```{r Clear global environment}
remove(list = ls())
```

```{r colour palletes for graphs}
palette2 <- c("#78D9C5","#F5BE5E")
palette2b <- c("#78D9C5","#EEB6E9")
```

```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(skimr)
library(ggpubr)
library(tidyr)
library(gplots)
library(Rmisc)
library(MplusAutomation)
library(stringr)
library(stringi)
library(filesstrings)
library(texreg)
library(relimp)
library(nnet)
library(ggplot2)
library(tidyverse)
library(dplyr) #conflicts with tidyverse for e.g. rename and row_number
```

```{r source the data file path}
#source raw data directory: data_raw and data included
source("../../../isolation_trajectories_data_path.R")
```

# Theme for plotting

```{r theme for all plots}
theme.antecedent <- theme(panel.grid.major.x = element_line(size = 0.5,
                                          linetype = 'dashed',
                                          colour = "gray"),
        plot.title = element_text(size = 15, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        axis.title.x = element_text(size=12,face="bold"),
        axis.title.y = element_blank(),
        axis.text.x = element_text(colour = "black", size = 9),
        axis.text.y = element_text(colour = "black", size = 12),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        panel.background = element_blank(),
        legend.position = "bottom",
        strip.text.y = element_blank())
```

# Variable lists

```{r lists for outcome regressions}
outcome.dependent <- c( # mental health
                      "depression_diagnosis_12mo_18", 
                       "depression_current_scale_18",
                       "anxiety_diagnosis_18", 
                       "anxiety_current_scale_18",
                       "ADHD_diagnosis_18", 
                       "inattentive_hyperactive_symptoms_total_18",
                       "conduct_disorder_moderate_18", 
                       "conduct_disorder_symptoms_18", 
                       "alcohol_dependence_18", 
                       "alcohol_abuse_18", # combine with the above?
                       "alcohol_symptom_scale_18",
                       "cannabis_dependence_18", 
                       "cannabis_symptom_scale_18", 
                       "PTSD_diagnosis_lifetime_18", 
                       "PTSD_diagnosis_current_18", 
                       "cat_psychosis_experiences_scale_18", 
                       "suicide_attempt_selfharm_18", 
                       "service_use_18", 
                      # physical health
                       "BMI_18", 
                       "CRP_log_18",
                       "physical_activity_18",
                       "smoking_current_number_18",
                      # coping and functioning
                       "loneliness_18",
                       "life_satisfaction_18",
                       "technology_use_18",
                       "coping_with_stress_18",
                       "PSQI_global_score_18",
                      # employment
                       "not_in_employment_education_18", 
                       "job_preparedness_skills_18",
                       "job_preparedness_attributes_18",
                       "optimism_18",
                       "job_search_activities_count_18")

outcome.dependent.numeric <- c( # mental health
                               "depression_diagnosis_12mo_numeric_18", 
                               "depression_current_scale_18",
                               "anxiety_diagnosis_numeric_18", 
                               "anxiety_current_scale_18",
                               "ADHD_diagnosis_numeric_18", 
                               "inattentive_hyperactive_symptoms_total_18",
                               "conduct_disorder_moderate_numeric_18", 
                               "conduct_disorder_symptoms_18", 
                               "alcohol_dependence_numeric_18", 
                               "alcohol_abuse_numeric_18", # combine with the above?
                               "alcohol_symptom_scale_18",
                               "cannabis_dependence_numeric_18", 
                               "cannabis_symptom_scale_18", 
                               "PTSD_diagnosis_lifetime_numeric_18", 
                               "PTSD_diagnosis_current_numeric_18", 
                               "cat_psychosis_experiences_scale_numeric_18", 
                               "suicide_attempt_selfharm_numeric_18", 
                               "service_use_numeric_18", 
                               "BMI_18", # physical health
                               "CRP_log_18",
                               "physical_activity_18",
                               "smoking_current_number_18",
                               "loneliness_18", # coping and functioning
                               "life_satisfaction_18",
                               "technology_use_18",
                               "coping_with_stress_18",
                               "PSQI_global_score_18",
                               "not_in_employment_education_numeric_18", # employment
                               "job_preparedness_skills_18",
                               "job_preparedness_attributes_18",
                               "optimism_18",
                               "job_search_activities_count_18")

name.list <- c(
              "Depression diagnosis: Yes",
              "Current depression",
              "Anxiety diagnosis: Yes",
              "Current anxiety",
              "ADHD diagnosis: Yes",
              "Current inattentive and hyperactive symptoms",
              "Moderate conduct disorder: Yes",
              "Current conduct disorder symptoms",
              "Alcohol dependence: Yes",
              "Alcohol abuse: Yes",
              "Current alcohol dependence symptoms",
              "Canabis dependence: Yes",
              "Current canabis dependence symptoms",
              "PTSD lifetime diagnosis: Yes",
              "PTSD current diagnosis: Yes",
              "Expereinces of psychosis",
              "Suicide attempt: Yes",
              "Service use: Yes",
              "BMI",
              "Log CRP",
              "Physical activity",
              "Current number of cigarettes smoked",
              "Loneliness",
              "Life satisfaction",
              "Technology use",
              "Coping with stress",
              "PSQI score",
              "Not in employment or education: Yes",
              "Job preparedness: skills",
              "Job preparedness: attributes",
              "Optimism",
              "Job search activities")

outcome.dependent.linear <- c(
  "depression_current_scale_18", # mental health
  "anxiety_current_scale_18",
  "inattentive_hyperactive_symptoms_total_18",
  "conduct_disorder_symptoms_18", 
  "alcohol_symptom_scale_18",
  "cannabis_symptom_scale_18", 
  "cat_psychosis_experiences_scale_18", 
  "BMI_18", # physical health
  "CRP_log_18",
  "physical_activity_18",
  "smoking_current_number_18",
  "loneliness_18", #coping and functioning
  "life_satisfaction_18",
  "technology_use_18",
  "coping_with_stress_18",
  "PSQI_global_score_18",
  "job_preparedness_skills_18", #employment
  "job_preparedness_attributes_18",
  "optimism_18",
  "job_search_activities_count_18"     
)

outcome.dependent.linear.name.list <- c(
                                        "Current depression",
                                        "Current anxiety",
                                        "Current inattentive and hyperactive symptoms",
                                        "Current conduct disorder symptoms",
                                        "Current alcohol dependence symptoms",
                                        "Current canabis dependence symptoms",
                                        "Expereinces of psychosis",
                                        "BMI",
                                        "Log CRP",
                                        "Physical activity",
                                        "Current number of cigarettes smoked",
                                        "Loneliness",
                                        "Life satisfaction",
                                        "Technology use",
                                        "Coping with stress",
                                        "PSQI score",
                                        "Job preparedness: skills",
                                        "Job preparedness: attributes",
                                        "Optimism",
                                        "Job search activities")


outcome.dependent.logistic <- c(
  "depression_diagnosis_12mo_18", # mental health 
  "anxiety_diagnosis_18", 
  "ADHD_diagnosis_18",
  "conduct_disorder_moderate_18",
  "alcohol_dependence_18",
  "alcohol_abuse_18", 
  "cannabis_dependence_18", 
  "PTSD_diagnosis_lifetime_18", 
  "PTSD_diagnosis_current_18", 
  "suicide_attempt_selfharm_18", 
  "service_use_18",
  "not_in_employment_education_18" #employment
)

outcome.dependent.logistic.name.list <- c(
  "Depression diagnosis: Yes",
  "Anxiety diagnosis: Yes",
  "ADHD diagnosis: Yes",
  "Moderate conduct disorder: Yes",
  "Alcohol dependence: Yes",
  "Alcohol abuse: Yes",
  "Canabis dependence: Yes",
  "PTSD lifetime diagnosis: Yes",
  "PTSD current diagnosis: Yes",
  "Suicide attempt: Yes",
  "Service use: Yes",
  "Not in employment or education: Yes"
)
```

# Read in data

```{r read in prepped regression data and check variable names, include=FALSE}
dat <- readRDS(file = paste0(data_path, "regression_data_THREE_traj.rds"))
colnames(dat)
```

# Test for multicollinearity fo routcome variables
Create correlation matrix to look for variables that are highly correlated.
Here parental antisocial behaviour and parental alcoholism have a correlation of 0.94 (Pearson). *Parental alcoholism has been excluded.*

```{r test for multicollinearity for outcome variables}
# create data frame with just numeric antecedent variables
outcome_data_frame <- as.data.frame(dat[,outcome.dependent.numeric])
colnames(outcome_data_frame) <- name.list

# get correlation matrix
isolation_cor_matrix <- cor(outcome_data_frame, method = "pearson", use = "complete.obs")

reorder_cor_matrix <- function(cor_matrix){
 dd <- as.dist((1-cor_matrix)/2) #Use correlation between variables as distance
 hc <- hclust(dd)
 cormat <-cor_matrix[hc$order, hc$order]
 }

# Reorder the correlation matrix
cor_matrix_full <- reorder_cor_matrix(isolation_cor_matrix)
  
#melt the values
metled_cor_matrix <- reshape::melt(cor_matrix_full, na.rm = TRUE)

#correlation heat map
correlation_heat_map <- ggplot(metled_cor_matrix, aes(X2, X1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white",
   midpoint = 0, limit = c(-1,1), space = "Lab",
    name="Pearson\nCorrelation") +
  theme_minimal() +
  labs(y = "",
       x = "",
       title = "Correlation heat map for outcomes")+
 theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),
       axis.text.y = element_text(size = 12),
       plot.margin=grid::unit(c(0,0,0,0), "mm"),
       plot.title = element_text(size = 20),
       )+
 coord_fixed() +
  geom_text(aes(X2, X1, label = round(value, digits = 2)), color = "black", size = 3)

correlation_heat_map

#save
ggsave(
  "outcomes_correlation_heat_map.png",
  plot = correlation_heat_map,
  device = "png",
  path = paste0(graph_save_data_path, "outcomes/analysis_interpretation"),
  width = 15,
  height = 15
)
```


# Association with outcome variables at age 18

ANALYSIS PLAN 

1. **Individual outcome regressions**: single regressions are computed to see if the class can predict the the occurrence of each variable (variable ~ class). This is done through a loop. Relative Risk Ratios (RRR) and their confidence intervals are calculated. Statistics for all variables are combined and RRR (95% CI) are plotted. FDR multiple testing correction is accounted for independent tests in each block. Will be a mix of linear and logistic regressions dependent on the variable. 

OR 

Is it best to look at group differences between the classes using Chi square? 

Linear regressions (continuous variables):
depression_current_scale_18, anxiety_current_scale_18, inattentive_hyperactive_symptoms_total_18, conduct_disorder_symptoms_18, alcohol_symptom_scale_18, cannabis_symptom_scale_18,   cat_psychosis_experiences_scale_18, BMI_18, CRP_log_18, physical_activity_18, smoking_current_number_18, loneliness_18, life_satisfaction_18, technology_use_18, coping_with_stress_18, PSQI_global_score_18, job_preparedness_skills_18, job_preparedness_attributes_18, optimism_18, job_search_activities_count_18

Logistic regressions (binary variables):
depression_diagnosis_12mo_18, anxiety_diagnosis_18, ADHD_diagnosis_18, conduct_disorder_moderate_18, alcohol_dependence_18, alcohol_abuse_18, cannabis_dependence_18, PTSD_diagnosis_lifetime_18, PTSD_diagnosis_current_18,suicide_attempt_selfharm_18, service_use_18, not_in_employment_education_numeric_18,

## Logistic regression loop

```{r logistic regression loop}
logistic.stats.outcome <- list() #create empty list to store the information in

for (i in 1:length(outcome.dependent.logistic))
{ cat(i)
  logistic.formula.outcome <- as.formula(sprintf("%s ~ class_reordered", outcome.dependent.logistic[i]))

  logistic.reg.outcome <- glm(logistic.formula.outcome, data = dat, family = "binomial")

  logistic.output.outcome.full <- broom::tidy(logistic.reg.outcome, conf.int = TRUE, exponentiate = TRUE, conf.level = 0.95)[-c(1),] %>%
    dplyr::rename(
           OR = estimate,
           OR.conf.low = conf.low,
           OR.conf.high = conf.high)
 
  if (i < 12) {
    logistic.output.outcome.full <- logistic.output.outcome.full %>%
      mutate(block = "Mental health block") %>%
      mutate(Variable = outcome.dependent.logistic.name.list[[i]]) %>%
                  mutate(p.value.adjusted = p.adjust(p.value, method = "fdr", n = 18)) # change depending if we combine a few variables 
  } else {
   logistic.output.outcome.full <- logistic.output.outcome.full %>%
      mutate(block = "Employment block") %>%
      mutate(Variable = outcome.dependent.logistic.name.list[[i]]) %>%
                  mutate(p.value.adjusted = p.adjust(p.value, method = "fdr", n = 5))
  }
  
   logistic.output.outcome.full <- logistic.output.outcome.full %>% 
    mutate(
    Significance = 
      if_else(
        p.value.adjusted < 0.05, 
        1,
        0
      ) %>%
    recode_factor(
        "0" = "Non-significant",
        "1" = "Significant"
      ))
    
  logistic.stats.outcome[[i]] <- logistic.output.outcome.full
}

# set the names of the list to match the variable names
names(logistic.stats.outcome) <- outcome.dependent.logistic

for (i in 1:length(outcome.dependent.logistic))
{ 
  assign(paste0("logistic.", names(logistic.stats.outcome[i]), ".outcome"), as.tibble(logistic.stats.outcome[[i]]))
}
```

## Linear regression loop

```{r linear regression loop}
linear.stats.outcome <- list() #create empty list to store the information in

for (i in 1:length(outcome.dependent.linear))
{ cat(i)
  linear.formula.outcome <- as.formula(sprintf("%s ~ class_reordered", outcome.dependent.linear[i]))

  linear.reg.outcome <- lm(linear.formula.outcome, data = dat)

  linear.output.outcome.full <- broom::tidy(linear.reg.outcome, conf.int = TRUE, conf.level = 0.95)[-c(1),] 
 
  if (i < 8) {
    linear.output.outcome.full <- linear.output.outcome.full %>%
      mutate(block = "Mental health block") %>%
      mutate(Variable = outcome.dependent.linear.name.list[[i]]) %>%
                  mutate(p.value.adjusted = p.adjust(p.value, method = "fdr", n = 18))
  } else if (i < 12 ) {
   linear.output.outcome.full <- linear.output.outcome.full %>%
      mutate(block = "Physical health block") %>%
      mutate(Variable = outcome.dependent.linear.name.list[[i]]) %>%
                  mutate(p.value.adjusted = p.adjust(p.value, method = "fdr", n = 4))
  } else if (i < 17) {
    linear.output.outcome.full <- linear.output.outcome.full %>%
      mutate(block = "Coping and functioning block") %>%
      mutate(Variable = outcome.dependent.linear.name.list[[i]]) %>%
                  mutate(p.value.adjusted = p.adjust(p.value, method = "fdr", n = 5))
  } else {
    linear.output.outcome.full <- linear.output.outcome.full %>%
      mutate(block = "Employment block") %>%
      mutate(Variable = outcome.dependent.linear.name.list[[i]]) %>%
                  mutate(p.value.adjusted = p.adjust(p.value, method = "fdr", n = 5))
  }
  
   linear.output.outcome.full <- linear.output.outcome.full %>% 
    mutate(
    Significance = 
      if_else(
        p.value.adjusted < 0.05, 
        1,
        0
      ) %>%
    recode_factor(
        "0" = "Non-significant",
        "1" = "Significant"
      ))
    
  linear.stats.outcome[[i]] <- linear.output.outcome.full
}
# set the names of the list to match the variable names
names(linear.stats.outcome) <- outcome.dependent.linear

for (i in 1:length(outcome.dependent.logistic))
{ 
  assign(paste0("linear.", names(linear.stats.outcome[i]), ".outcome"), as.tibble(linear.stats.outcome[[i]]))
}
```















# Pairwise comparisons - ADAPT FOR OUTCOMES
**We have decided this doesn't make sense for antecedents, as this is only measuring the differences in the intercept rather than the slope. This will be used at the outcome stage and not for antecedent analyses**

## Create new dataset dropping those who are in the lowstable class 

```{r create pairwise dataset and order class variable}
# dat.pairwise <- dat %>%
#   filter(class_renamed != "Low stable") %>%
#   mutate_if(is.factor, ~forcats::fct_drop(., only = "Low stable"))
# 
# dat.pairwise <- dat.pairwise %>%
#   mutate(
#     class_binary = 
#       as.factor(class_renamed) %>%
#       relevel(ref = "Increasing", 
#               first = TRUE,
#               collapse = "+",
#               xlevels = TRUE)
#   )
```

```{r loop for running antecedent regressions, include=FALSE}
# logistic.stats.ant <- list() #create empty list to store the information in
# 
# for (i in 1:length(antecedent.independent))
# { cat(i)
#   logistic.formula.ant <- as.formula(sprintf("class_binary ~ %s", antecedent.independent[i]))
# 
#   logistic.reg.ant <- glm(logistic.formula.ant, data = dat.pairwise, family = "binomial")
# 
#   logistic.output.ant.full <- broom::tidy(logistic.reg.ant, conf.int = TRUE, exponantiate = TRUE)[-c(1),] %>%
#     add_column(Variable = name.list[[i]]) %>%
#     add_column(OR = exp(coef(logistic.reg.ant))[2]) %>% 
#     add_column(`Lower CI` = exp(confint(logistic.reg.ant))[2,1]) %>%
#     add_column(`Upper CI` = exp(confint(logistic.reg.ant))[2,2])
#  
#   if (i < 7) {
#     logistic.output.ant.full <- logistic.output.ant.full %>%
#       mutate(block = "Social block") %>%
#                   mutate(p.value.adjusted = p.adjust(p.value, method = "fdr", n = 6))
#   } else if (i < 13 ) {
#    logistic.output.ant.full <- logistic.output.ant.full %>%
#       mutate(block = "Home block") %>%
#                   mutate(p.value.adjusted = p.adjust(p.value, method = "fdr", n = 7))
#   } else if (i < 21) {
#     logistic.output.ant.full <- logistic.output.ant.full %>%
#       mutate(block = "Parent block") %>%
#                   mutate(p.value.adjusted = p.adjust(p.value, method = "fdr", n = 8))
#   } else if (i < 24) {
#     logistic.output.ant.full <- logistic.output.ant.full %>%
#       mutate(block = "Neuro block") %>%
#                   mutate(p.value.adjusted = p.adjust(p.value, method = "fdr", n = 3))
#   } else {
#     logistic.output.ant.full <- logistic.output.ant.full %>%
#       mutate(block = "Emo/behave block") %>%
#                   mutate(p.value.adjusted = p.adjust(p.value, method = "fdr", n = 4))
#   }
#   
#    logistic.output.ant.full <- logistic.output.ant.full %>% 
#     mutate(
#     Significance = 
#       if_else(
#         p.value.adjusted < 0.05, 
#         1,
#         0
#       ) %>%
#     recode_factor(
#         "0" = "Non-significant",
#         "1" = "Significant"
#       ))
#   
#   logistic.stats.ant[[i]] <- logistic.output.ant.full
# }
# # set the names of the list to match the variable names
# names(logistic.stats.ant) <- antecedent.independent
```

```{r call statistics for all continuous antecedent regressions, include=FALSE}
# for (i in 1:length(antecedent.independent))
# { 
#   assign(paste0("logistic.", names(logistic.stats.ant[i])), as.tibble(logistic.stats.ant[[i]]))
# }
# 
# logistic.stats.all <- bind_rows(logistic.SES_binary_releveled, logistic.acorn_recoded_05, logistic.vandalism_05, logistic.problems_neighbours_05, logistic.number_children_school_05, logistic.number_children_school_free_meals_05, logistic.child_harm_recoded_05, logistic.total_siblings_05, logistic.total_social_support_05, logistic.total_activities_with_mum_05, logistic.mum_notlived_biodad_sincebirth_05, logistic.any_domestic_violence_05, logistic.maternal_depression_lifetime_05, logistic.maternal_personality_openness_05, logistic.maternal_personality_conscientiousness_05, logistic.maternal_personality_extroversion_05, logistic.maternal_personality_agreeableness_05, logistic.maternal_personality_neuroticism_05, logistic.antisocial_behaviour_parent_05, logistic.alcoholism_parent_05, logistic.IQ_05, logistic.executive_function_05, logistic.theory_of_mind_05, logistic.externalising_combined_05, logistic.internalising_combined_excl_sis_05, logistic.ADHD_combined_05, logistic.prosocial_behaviours_combined_05 )
# 
# logistic.stats.all.table <- kable(logistic.stats.all)
```

```{r pairwise plot}
# pairwise.all.plot <- ggplot(
#   logistic.stats.all,
#   aes(x = Variable,
#       y = OR,
#       ymin = `Lower CI`,
#       ymax = `Upper CI`)
#   ) +
#   scale_fill_manual(values = palette2b) +
#   scale_color_manual(values = palette2b) +
#   scale_shape_manual(values = c(21, 19)) +
#   geom_pointrange(aes(shape = Significance), position = position_dodge(width = 0.7)) +
#   labs(y = "OR (95% CI) \n OR>1 shows association with being in the decreasing class compared to increasing",
#        x = "Symptom",
#        title = "Antecedents and pairwise class membership",
#        subtitle = paste("N(Total) = ", length(dat.pairwise$sex),
#                         "; N(Increasing) = ", sum(!is.na(dat.pairwise$class_binary[dat.pairwise$class_binary=="Increasing"])),
#                         "; N(Decreasing) = ", sum(!is.na(dat.pairwise$class_binary[dat.pairwise$class_binary=="Decreasing"])), sep = ""),
#        color = "",
#        shape = "") +
#   theme(panel.grid.major.x = element_line(size = 0.5,
#                                           linetype = 'dashed',
#                                           colour = "gray"),
#         plot.title = element_text(size = 12, face = "bold", hjust=0.5),
#         plot.subtitle = element_text(hjust = 0.5),
#         axis.title.x = element_text(size=10,face="bold"),
#         axis.title.y = element_blank(),
#         axis.text.x = element_text(colour = "black", size = 9),
#         axis.text.y = element_text(colour = "black", size = 10),
#         axis.ticks.x = element_blank(),
#         axis.ticks.y = element_blank(),
#         panel.background = element_blank(),
#         legend.position = "bottom") +
#   geom_hline(linetype = "dashed", yintercept = 1) +
#   scale_y_continuous(trans = scales::log10_trans(), breaks=c(0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1, 1.5, 2, 2.5, 3, 3.5, 4)) +
#   coord_flip()
# 
# pairwise.all.plot
# 
# #save
# ggsave(
#   "pairwise.all.plot.png",
#   plot = pairwise.all.plot,
#   device = "png",
#   path = paste0(graph_save_data_path, "antecedents/"),
#   width = 10,
#   height = 7
# )

```










