---
title: "Associations between social isolation trajectories and antecedents"
output:  
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: flatly
    code_folding: hide
    includes:
      after_body: footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = NA,
                      prompt = FALSE,
                      cache = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      results = 'asis')

options(bitmapType = 'quartz') # to render fonts better
```

```{r Clear global environment, include=FALSE}
remove(list = ls())
```

```{r colour palletes for graphs, include=FALSE}
palette2 <- c("#78D9C5","#F5BE5E")
palette2b <- c("#78D9C5","#EEB6E9")
```

```{r Load packages, include=FALSE}
library(summarytools)
library(MplusAutomation)
library(filesstrings)
library(nnet)
library(ggplot2)
library(knitr)
library(tidyverse)
library(dplyr) #conflicts with tidyverse for e.g. rename and row_number
```

```{r source the data file path, include=FALSE}
#source raw data directory: data_raw and data included
source("../../../isolation_trajectories_data_path.R")
```

```{r read in prepped regression data and check variable names, include=FALSE}
dat <- readRDS(file = paste0(data_path, "class_joined_preprocessed_isolation_data_full_sample.rds"))
#colnames(dat)
```

```{r variable character lists, include=FALSE}
antecedent.independent.numeric <- c("SES_binary_releveled_numeric",
                                    "acorn_continuous_05",
                                    "vandalism_05",
                                    "problems_neighbours_05",
                                    "number_children_school_05",
                                    "number_children_school_free_meals_05",
                                    "child_harm_recoded_numeric_05",
                                    "total_siblings_05",
                                    "total_social_support_05",
                                    "total_activities_with_mum_05",
                                    "mum_notlived_biodad_sincebirth_numeric_05",
                                    "any_domestic_violence_numeric_05",
                                    "maternal_warmth_continuous_05",
                                    "maternal_depression_lifetime_numeric_05",
                                    "maternal_personality_openness_05",
                                    "maternal_personality_conscientiousness_05",
                                    "maternal_personality_extroversion_05",
                                    "maternal_personality_agreeableness_05",
                                    "maternal_personality_neuroticism_05",
                                    "antisocial_behaviour_parent_05",
                                    "alcoholism_parent_05", 
                                    "IQ_05",
                                    "executive_function_05",
                                    "theory_of_mind_05",
                                    "externalising_combined_05",
                                    "internalising_combined_excl_sis_05",
                                    "ADHD_combined_05",
                                    "prosocial_behaviours_combined_05")

antecedent.independent.continuous <- c("acorn_continuous_05",
                                      "vandalism_05",
                                      "problems_neighbours_05",
                                      "number_children_school_05",
                                      "number_children_school_free_meals_05",
                                      "total_siblings_05",
                                      "total_social_support_05",
                                      "total_activities_with_mum_05",
                                      "maternal_warmth_continuous_05",
                                      "maternal_personality_openness_05",
                                      "maternal_personality_conscientiousness_05",
                                      "maternal_personality_extroversion_05",
                                      "maternal_personality_agreeableness_05",
                                      "maternal_personality_neuroticism_05",
                                      "antisocial_behaviour_parent_05",
                                      "alcoholism_parent_05",
                                      "IQ_05",
                                      "executive_function_05",
                                      "theory_of_mind_05",
                                      "externalising_combined_05",
                                      "internalising_combined_excl_sis_05",
                                      "ADHD_combined_05",
                                      "prosocial_behaviours_combined_05")
                                 
name.list <- c("SES: Low",
               "ACORN", 
               "Vandalism", 
               "Problems with neighbours", 
               "Number of children in school", 
               "Number of children eligible for free school meals",
               "Child harm: Harmed",
               "Total siblings",
               "Social support", 
               "Total activities with mum",
               "Mum not lived with biological dad since birth: Yes", 
               "Any domestic violence: Yes",
               "Maternal warmth",
               "Maternal lifetime depression: Yes",
               "Maternal personality: Openness",
               "Maternal personality: Conscientiousness",
               "Maternal personality: Extroversion",
               "Maternal personality: Agreeableness",
               "Maternal personality: Neuroticism",
               "Parental antisocial behaviour", 
               "Parental alcoholism", 
               "Child IQ",
               "Child executive function",
               "Child theory of mind",
               "Child externalising behaviours",
               "Child internalising behaviours",
               "Child ADHD behaviours",
               "Child prosocial behaviours")
```

***

This page displays the results from multiple multinomial logisitc regressions used to assess how the antecedents are associated with trajectory membership. The multinomial regressions were run in STATA to enable clustering for familyid. Antecedents are sorted into domains of "Social factors", "Home factors", "Parent characteristics", "Neurological child characteristics", and "Emotional/behavioural child characteristics". 

***

# Functions 

To run the following code more clearly, I have created three functions:

* multiple.testing() applies the BH multiple testing correction. You define the data, class, method and number of tests. 
* paper.table() creates a table that labels the domain of the variable, the comparison class, Relative risk ratios (RRR), confidence intervals, and p value from the multinomial regression. 
* order.variables() orders the data whereby points in a plot are arranged by RRR and domain. 

```{r script functions, class.source = 'fold-show'}
multiple.testing <- function(data, class, method = "BH", n.tests){
  data <- data %>%
    filter(Class == class) %>%
  mutate(
    p.value.adjusted = 
           stats::p.adjust(p.value, 
                    method = method,
                    n = n.tests) 
  ) %>%
  mutate(Significance.adjusted = 
           if_else(
            p.value.adjusted < 0.05, 
            1,
            0
      ) %>%
        recode_factor(
          `1` = "Significant",
          `0` = "Non-significant")
         )  
  return(data)
}

paper.table <- function(data){
  
  table <- data %>%
  arrange(match(Domain, c("Social domain", "Home domain", "Parent domain", "Neuro domain", "Emo/behave domain"))) %>%
  select(
    Class,
    Variable,
    RRR,
    `95% CI low` = RRR.conf.low,
    `95% CI high` = RRR.conf.high,
    `p value` = p.value,
    Domain,
    -Significance) %>%
  kable(digits = 2)
  
  return(table)
}

order.variables <- function(data, domain){
  data <- filter(data, Domain == domain) %>%
    arrange(desc(RRR))
  
  data$Variable <- factor(data$Variable, 
                          levels = paste(unique(paste(data$Variable)), sep = ","))
  return(data)
}
```

```{r theme for all plots, include=FALSE}
theme.antecedent <- theme(panel.grid.major.x = element_line(size = 0.5,
                                          linetype = 'dashed',
                                          colour = "gray"),
                          plot.title = element_text(size = 12,face = "bold"),
                          plot.subtitle = element_text(hjust = 0.5),
                          axis.title.x = element_text(size = 10, face = "bold"),
                          axis.title.y = element_blank(),
                          axis.text.x = element_text(size = 8, colour = "black"),
                          axis.text.y = element_text(colour = "black", size = 10),
                          axis.ticks.x = element_blank(),
                          axis.ticks.y = element_blank(),
                          panel.background = element_blank(),
                          legend.position = "bottom",
                          strip.text.y = element_blank(),
                          strip.text.x = element_text(size = 10, color = "black", face = "bold"),
                          strip.background = element_rect(color = "black", fill = "white", size = 0.5, linetype = "solid")          
  )
```

# Test for multicollinearity 

We created a correlation matrix to look for variables that are too highly correlated (multicollinearity). 

```{r test for multicollinearity}
# create data frame with just numeric antecedent variables
antececent_data_frame <- as.data.frame(dat[,antecedent.independent.numeric])
colnames(antececent_data_frame) <- name.list

# get correlation matrix
isolation_cor_matrix <- cor(antececent_data_frame, method = "pearson", use = "pairwise.complete.obs")

reorder_cor_matrix <- function(cor_matrix){
 dd <- as.dist((1-cor_matrix)/2) #Use correlation between variables as distance
 hc <- hclust(dd)
 cormat <-cor_matrix[hc$order, hc$order]
 }

# Reorder the correlation matrix
cor_matrix_full <- reorder_cor_matrix(isolation_cor_matrix)
  
#melt the values
metled_cor_matrix <- reshape::melt(cor_matrix_full, na.rm = TRUE)

#correlation heat map
correlation_heat_map <- ggplot(metled_cor_matrix, aes(X2, X1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white",
   midpoint = 0, limit = c(-1,1), space = "Lab",
    name="Pearson\nCorrelation") +
  theme_minimal() +
  labs(y = "",
       x = "",
       title = "Correlation heat map for antecedents")+
 theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),
       axis.text.y = element_text(size = 12),
       plot.margin=grid::unit(c(0,0,0,0), "mm"),
       plot.title = element_text(size = 20),
       )+
 coord_fixed() +
  geom_text(aes(X2, X1, label = round(value, digits = 2)), color = "black", size = 3)
```

```{r heat map fig, fig.width=15, fig.height=15}
correlation_heat_map
```

```{r save heatmap}
# ggsave(
#   "antecedent_correlation_heat_map.png",
#   plot = correlation_heat_map,
#   device = "png",
#   path = paste0(graph_save_data_path, "antecedents/analysis_interpretation"),
#   width = 15,
#   height = 15
# )
```

# Posterior sensitivity analysis 

The below few chunks are used to run the same analysis again in a reduced sample (only individuals who had a posterior probability of above 0.8 in all classes). 

**IMPORTANT**: We created the current Rmd script to be able to use the same script to run the same analysis again in a reduced sample as a sensitivity analysis (only individuals who had a posterior probability of above 0.8 in all classes). When the variable "posterior.cut" is set to TRUE the following code will run for the data set that has those probability<0.8 dropped.  

```{r posterior probability indicator, results='hide', class.source = 'fold-show'}
posterior.cut = FALSE
```

```{r posterior probability code, results='hide', class.source = 'fold-show'}
if(posterior.cut == TRUE){

# create new variable only with those who have a probability higher than 0.8 for each class
dat <- dat %>%
  mutate(class_renamed_cut =
           factor(
             case_when(
             prob2 > 0.8 & class_renamed == "Increasing" ~ "Increasing",
             prob3 > 0.8 & class_renamed == "Decreasing" ~ "Decreasing",
             prob1 > 0.8 & class_renamed == "Low stable" ~ "Low stable"
           ), 
           ordered = FALSE))

library(nnet)
# check the biggest group in the class to be the reference group
table(dat$class_renamed_cut)

# relevel class_factor
dat <- dat %>% 
  mutate(
    class_reordered_cut =
      relevel(
        class_renamed_cut,
        ref = "Low stable",
        first = TRUE, #levels in ref come first
        collapse = "+", #String used when constructing names for combined factor levels
        xlevels = TRUE #levels maintained even if not actually occurring
        
      )
  )
# check
table(dat$class_reordered_cut)

# create new data set with only these individuals
dat_prob_cut_out <- dat %>% filter(is.na(class_reordered_cut))
dat <- dat %>% filter(!is.na(class_reordered_cut))

# check
freq(dat$class_reordered)
}
```

Statistics for posterior probability sensitivity analyses:

* Low stable: Dropped 42 people. Average probability = 0.66, minimum probability = 0.48. 
* Increasing: Dropped 24 people. Average probability = 0.64, minimum probability = 0.49.
* Decreasing: Dropped 25 people. Average probability = 0.63, minimum probability = 0.51.

**Posterior sensitivity N = 2141** 

```{r descriptives for probabilites, results='hide', class.source = 'fold-show'}
if(posterior.cut == TRUE){
low.stable.cut.stats <- dat_prob_cut_out %>%
  filter(class_reordered == "Low stable") %>%
  descr(prob1)

increasing.cut.stats <- dat_prob_cut_out %>%
  filter(class_reordered == "Increasing") %>%
  descr(prob2)

decreasing.cut.stats <- dat_prob_cut_out %>%
  filter(class_reordered == "Decreasing") %>%
  descr(prob3)

low.stable.cut.stats
increasing.cut.stats
decreasing.cut.stats
}
```

# Antecedent analysis

The following variables were dropped for the analyses: multinom.class_size_average_05, multinom.resident_moves_05, multinom.temp_negative_affect_05, multinom.temp_impulsivity_05, multinom.temp_approach_05, multinom.temp_sluggishness_05, multinom.temp_wariness_05, multinom.temp_undercontrolled_05, multinom.temp_inhibited_05, multinom.temp_shy_05. 

**Analysis plan** 

1. *Individual antecedent regressions*: single regressions are computed for each variable (class ~ variable). This is done through a loop. Relative Risk Ratios (RRR) and their confidence intervals are calculated. Statistics for all variables are combined and RRR (95% CI) are plotted for the Increasing and Decreasing class. 

2. *Full model*: All variables that significantly predict class membership are controlled for. E.g. (class ~ social_variable_1 + home_variable_2 + neuro_variable_3 + emo_variable_4). RRR (95% CI) are calculated. Z scores for continuous variables are used. No multiple testing correction is applied at this step.

All regressions are first computed in STATA - multinomial regression with clustering is not currently doable in R. The coefficients are then imported into R to be interpreted. The RRR coefficient indicates the likelihood of having higher values on the variable of interest, for those following either the increasing and decreasing class compared to the low stable. 

Binary variables: SES (middle/high vs low), Child harm (No harm vs possible/definite), any domestic violence (Yes/No), maternal lifetime depression (Yes/No). ACORN (1-5) and maternal warmth (1-6) have been recoded to be continuous variables.

## STATA file preparation

Create *z scores* for all continuous variables

```{r scale continuous variables and combine to main dataset}
# create a scaled version of the continuous variables - z scores for all variables
scaled_continuous_data <- as.data.frame(scale(dat[antecedent.independent.continuous], center = TRUE, scale = TRUE))

# rename variables to show that they are z scores 
colnames(scaled_continuous_data) <- paste0("z_score.", colnames(scaled_continuous_data))

# combine the scaled data to the main dataset 
dat <- cbind(dat, scaled_continuous_data)
```

Create STATA file 

```{r create mlogit STATA file}
data.mlogit <- dat %>% # need to rename for stata format
  select(id,
         familyid,
         sex,
         classreordered = class_reordered,
         ses = SES_binary_releveled,
         acorn = z_score.acorn_continuous_05,
         vandalism = z_score.vandalism_05,
         probneighbours = z_score.problems_neighbours_05,
         numberchildschool = z_score.number_children_school_05,
         meals = z_score.number_children_school_free_meals_05,
         harm = child_harm_recoded_05,
         siblings = z_score.total_siblings_05,
         socialsupport = z_score.total_social_support_05,
         activities = z_score.total_activities_with_mum_05,
         biodad = mum_notlived_biodad_sincebirth_05,
         domesticviolence = any_domestic_violence_05,
         maternalwarmth = z_score.maternal_warmth_continuous_05,
         maternaldepression = maternal_depression_lifetime_05,
         openness = z_score.maternal_personality_openness_05,
         conscientiousness = z_score.maternal_personality_conscientiousness_05,
         extroversion = z_score.maternal_personality_extroversion_05,
         agreeableness = z_score.maternal_personality_agreeableness_05,
         neuroticism = z_score.maternal_personality_neuroticism_05,
         antisocial = z_score.antisocial_behaviour_parent_05,
         alcohol = z_score.alcoholism_parent_05,
         iq = z_score.IQ_05,
         execfunction = z_score.executive_function_05,
         theoryofmind = z_score.theory_of_mind_05,
         externalising = z_score.externalising_combined_05,
         internalising = z_score.internalising_combined_excl_sis_05,
         adhd = z_score.ADHD_combined_05,
         prosocial = z_score.prosocial_behaviours_combined_05)
```

Write data files for STATA to be used in multinomial regression. We have saved two different files depending on if we're doing the sensitivity analysis (reduced sample) or not. 

```{r write STATA file for multinom}
library(foreign)
if(posterior.cut == FALSE){
  write.dta(data.mlogit, "/Users/katiethompson/Documents/PhD/LISS-DTP_Louise_and_Tim/Social isolation trajectories_Paper 1/data_analysis/data_full/data_raw/multinom.reg.stata.dta")
}

if(posterior.cut == TRUE){
  write.dta(data.mlogit, "/Users/katiethompson/Documents/PhD/LISS-DTP_Louise_and_Tim/Social isolation trajectories_Paper 1/data_analysis/data_full/data_raw/multinom.reg.stata.POSTERIOR.dta")
}
```

## STATA multinomial regressions

The mlogit code was run in STATA to calculate RRR, CI and pvalues for all antecedents entered into a single multinomial regression model. If you would like to run this in STATA, you can open the STATA file *"antecedent_multinomial_regression.do"* or *"antecedent_multinomial_regression_POSTERIOR.do"* from the Github home page. The output from the STATA file is saved as a txt file containing all the information from the regressions. 

Three models were run in STATA:

1. Individual regressions. Each antecedent was regressed on class membership. 
2. Significant from univariate model. **This was NOT taken forward into analyses - only the individual and full model to avoid a stepwise approach**
3. All model. This includes all antecedents in one model predicting class membership. 

```{r, eval=FALSE, class.source = 'fold-show'}
// read in file
use "multinom.reg.stata.dta"

// create list of outcome variables
local outcomes "ses acorn vandalism probneighbours numberchildschool meals harm siblings socialsupport activities biodad domesticviolence maternalwarmth maternaldepression openness conscientiousness extroversion agreeableness neuroticism antisocial alcohol iq execfunction theoryofmind externalising internalising adhd prosocial"

// loop for multinomial regression clustered by familyid
foreach var of varlist `outcomes' {
mlogit classreordered `var' sex, cluster(familyid)

estimates store results_`var'
}

// install package estout
// ssc install estout

// save results, dropping estimates we don't need and saving the RRR and upper and lower CI for each variable 
estout results_* using multinomial_results_si.txt, drop(sex) cells(b(fmt(2)) ci_l(fmt(2)) ci_u(fmt(2)) p(fmt(8))) eform replace

// all model - predictors that are significant from UNIVARIATE models
mlogit classreordered ses acorn probneighbours meals harm siblings socialsupport activities biodad domesticviolence maternalwarmth maternaldepression openness conscientiousness extroversion agreeableness neuroticism antisocial alcohol iq execfunction theoryofmind externalising internalising adhd prosocial sex, cluster(familyid)
estimates store allmodelsiguni 
estout allmodelsiguni using multinomial_results_si.allmodel.SIGUNI.txt, drop(sex) cells(b(fmt(2)) ci_l(fmt(2)) ci_u(fmt(2)) p(fmt(8))) eform replace

// all model - all predictors in one model
mlogit classreordered ses acorn vandalism probneighbours numberchildschool meals harm siblings socialsupport activities biodad domesticviolence maternalwarmth maternaldepression openness conscientiousness extroversion agreeableness neuroticism antisocial alcohol iq execfunction theoryofmind externalising internalising adhd prosocial sex, cluster(familyid)
estimates store allmodel 
estout allmodel using multinomial_results_si.allmodel.txt, drop(sex) cells(b(fmt(2)) ci_l(fmt(2)) ci_u(fmt(2)) p(fmt(8))) eform replace
```

Th next bit of code reades the files and moves the txt files to the right place - once this has been done the code is hashed out. 

```{r run stata do file from r}
if(posterior.cut == FALSE){
# move txt file to data file 
# file.move(paste0(stata_data_path, "multinomial_results_si.txt"), data_path, overwrite = TRUE)
# file.move(paste0(stata_data_path,"multinomial_results_si.allmodel.SIGUNI.txt"), data_path, overwrite = TRUE)
# file.move(paste0(stata_data_path,"multinomial_results_si.allmodel.txt"), data_path, overwrite = TRUE)
  
# individual regressions
dat.stata.individual.raw <- read_tsv(paste0(data_path, "multinomial_results_si.txt")) 
# all model with only univariate sig variables
dat.stata.all.siguni.raw <- read_tsv(paste0(data_path, "multinomial_results_si.allmodel.SIGUNI.txt"))
# all model
dat.stata.all.raw <- read_tsv(paste0(data_path, "multinomial_results_si.allmodel.txt"))
}

if(posterior.cut == TRUE){
# move txt file to data file 
# file.move(paste0(stata_data_path, "multinomial_results_si_POSTERIOR.txt"), data_path, overwrite = TRUE)
# file.move(paste0(stata_data_path,"multinomial_results_si.allmodel.SIGUNI_POSTERIOR.txt"), data_path, overwrite = TRUE)
# file.move(paste0(stata_data_path,"multinomial_results_si.allmodel_POSTERIOR.txt"), data_path, overwrite = TRUE)

# individual regressions
dat.stata.individual.raw <- read_tsv(paste0(data_path, "multinomial_results_si_POSTERIOR.txt")) 
# all model with only univariate sig variables
dat.stata.all.siguni.raw <- read_tsv(paste0(data_path, "multinomial_results_si.allmodel.SIGUNI_POSTERIOR.txt"))
# all model
dat.stata.all.raw <- read_tsv(paste0(data_path, "multinomial_results_si.allmodel_POSTERIOR.txt"))
}
```

## Reformat STATA output {.tabset .tabset-fade} 

Layout from the STATA output file:

* b/min95/max95/p 

We now rearrange this file to create the antecedent multinomial tables needed for the manuscript.  

### Univariate regressions 

```{r reformat stata data frame}
dat.stata.individual <- dat.stata.individual.raw[c(120:231,237:348),-c(1)] %>%
  `colnames<-`(name.list) %>%
  mutate(Class =
    case_when(
    row_number() < 113 ~ "Increasing",
    row_number() > 112 ~ "Decreasing")) %>%
  mutate(Variable = 
           rep(c("RRR", "RRR.conf.low", "RRR.conf.high", "p.value"), 56)) %>%
  gather(key, value, `SES: Low`:`Child prosocial behaviours`, factor_key = TRUE) %>%
  drop_na() %>%
  spread(Variable, value) %>%
  dplyr::rename(Variable = key) %>%
  mutate(
    Significance = 
      if_else(
        p.value < 0.05, 
        1,
        0
      ) %>%
    recode_factor(
        "0" = "Non-significant",
        "1" = "Significant")) %>%
  mutate(
    block = 
      case_when(row_number() < 7 ~ "Social domain",
                row_number() < 14 ~ "Home domain",
                row_number() < 22 ~ "Parent domain",
                row_number() < 25 ~ "Neuro domain",
                row_number() < 29 ~ "Emo/behave domain",
                row_number() < 35 ~ "Social domain",
                row_number() < 42 ~ "Home domain",
                row_number() < 50 ~ "Parent domain",
                row_number() < 53 ~ "Neuro domain",
                row_number() > 52 ~ "Emo/behave domain")) %>%
  transmute(Class = Class,
            Variable = Variable,
            RRR = as.numeric(RRR),
            RRR.conf.low = as.numeric(RRR.conf.low),
            RRR.conf.high = as.numeric(RRR.conf.high),
            p.value = as.numeric(p.value),
            Significance = Significance,
            Domain = block)
```

#### Table 

```{r table for stata coefficients}
univariate.reg.table <- paper.table(dat.stata.individual)
univariate.reg.table
```

```{r multiple testing correction for univariate associations, include=FALSE}
dat.stata.individual.decreasing <- multiple.testing(data = dat.stata.individual,
                                                    class = "Decreasing",
                                                    method = "BH",
                                                    n.tests = 28)      

dat.stata.individual.increasing <- multiple.testing(data = dat.stata.individual,
                                                    class = "Increasing",
                                                    method = "BH",
                                                    n.tests = 28)
dat.stata.individual.fdr <- rbind(
  dat.stata.individual.decreasing,
  dat.stata.individual.increasing
)

dat.stata.individual.fdr %>%
  filter(Significance == "Significant" & Significance.adjusted == "Non-significant")
```

Bind *z score* trasformed values to order the graph according to block of variables

```{r bind and order rows for the plot}
individual.home <- order.variables(data = dat.stata.individual, domain = "Home domain")
individual.social <- order.variables(data = dat.stata.individual, domain = "Social domain")
individual.parent <- order.variables(data = dat.stata.individual, domain = "Parent domain")
individual.neuro <- order.variables(data = dat.stata.individual, domain = "Neuro domain")
individual.emo <- order.variables(data = dat.stata.individual, domain = "Emo/behave domain")

# join all
dat.stata.individual.ordered <- rbind(
  individual.social,
  individual.home,
  individual.parent,
  individual.neuro,
  individual.emo
)
```

#### Plot for univariate regressions 

```{r stata plot to be used for results}
stata.plot.individual <- ggplot(
  dat.stata.individual.ordered, 
  aes(x = Variable,
      y = RRR,
      ymin = RRR.conf.low,
      ymax = RRR.conf.high,)
  ) +
  scale_fill_manual(values = "black") +
  scale_color_manual(values = "black") +
  scale_shape_manual(values = c(19)) +
  geom_pointrange(position = position_dodge(width = 0.7), 
                  show.legend = FALSE, 
                  size = 0.5) +
 facet_grid(Domain ~ Class, drop = TRUE, scales = "free_y", space = "free") +
  labs(y = "RRR (95% CI)",
       x = "names",
       shape = "") +
  theme.antecedent +
  geom_hline(linetype = "dashed", yintercept = 1) +
 scale_y_continuous(trans = scales::log10_trans(), breaks=c(0.4,0.6,0.8, 1, 1.5, 2, 2.5, 3, 3.5, 4)) +
  coord_flip() 
```
```{r individual plot, fig.height=6.5, fig.width=11}
stata.plot.individual
```

```{r save final antecedent plot, include=FALSE}
# if(posterior.cut == FALSE){
# #save
# ggsave(
#   "STATA_All_antecedents_individual_regressions_using_zscores_final.png",
#   plot = stata.plot.individual,
#   device = "png",
#   path = paste0(graph_save_data_path, "antecedents/final"),
#   width = 11,
#   height = 6.5
# )
# }
```

### Multivariate model: all variables included

```{r reformat stata data all model}
dat.stata.all <- dat.stata.all.raw[c(120:231,237:348),] %>%
  mutate(Class =
    case_when(
    row_number() < 113 ~ "Increasing",
    row_number() > 112 ~ "Decreasing")) %>%
  mutate(key = 
           rep(c("RRR", "RRR.conf.low", "RRR.conf.high", "p.value"), 56)) %>%
  mutate(Variable =
    case_when(row_number() < 5  ~ "SES: Low",
              row_number() < 9  ~ "ACORN",
              row_number() < 13 ~ "Vandalism",
              row_number() < 17 ~ "Problems with neighbours",
              row_number() < 21 ~ "Number of children in school",
              row_number() < 25 ~ "Number of children eligible for free school meals",
              row_number() < 29 ~ "Child harm: Harmed",
              row_number() < 33 ~ "Total siblings",
              row_number() < 37 ~ "Social support",
              row_number() < 41 ~ "Total activities with mum",
              row_number() < 45 ~ "Mum not lived with biological dad since birth: Yes",
              row_number() < 49 ~ "Any domestic violence: Yes" ,
              row_number() < 53 ~ "Maternal warmth",
              row_number() < 57 ~ "Maternal lifetime depression: Yes",
              row_number() < 61 ~ "Maternal personality: Openness",
              row_number() < 65 ~ "Maternal personality: Conscientiousness",
              row_number() < 69 ~ "Maternal personality: Extroversion",
              row_number() < 73 ~ "Maternal personality: Agreeableness",
              row_number() < 77 ~ "Maternal personality: Neuroticism",
              row_number() < 81 ~ "Parental antisocial behaviour",
              row_number() < 85 ~ "Parental alcoholism",
              row_number() < 89 ~ "Child IQ",
              row_number() < 93 ~ "Child executive function",
              row_number() < 97 ~ "Child theory of mind",
              row_number() < 101 ~ "Child externalising behaviours" ,
              row_number() < 105 ~ "Child internalising behaviours",
              row_number() < 109 ~ "Child ADHD behaviours",
              row_number() < 113 ~ "Child prosocial behaviours",
              row_number() < 117  ~ "SES: Low",
              row_number() < 121  ~ "ACORN",
              row_number() < 125 ~ "Vandalism",
              row_number() < 129 ~ "Problems with neighbours",
              row_number() < 133 ~ "Number of children in school",
              row_number() < 137 ~ "Number of children eligible for free school meals",
              row_number() < 141 ~ "Child harm: Harmed",
              row_number() < 145 ~ "Total siblings",
              row_number() < 149 ~ "Social support",
              row_number() < 153 ~ "Total activities with mum",
              row_number() < 157 ~ "Mum not lived with biological dad since birth: Yes",
              row_number() < 161 ~ "Any domestic violence: Yes" ,
              row_number() < 165 ~ "Maternal warmth",
              row_number() < 169 ~ "Maternal lifetime depression: Yes",
              row_number() < 173 ~ "Maternal personality: Openness",
              row_number() < 177 ~ "Maternal personality: Conscientiousness",
              row_number() < 181 ~ "Maternal personality: Extroversion",
              row_number() < 185 ~ "Maternal personality: Agreeableness",
              row_number() < 189 ~ "Maternal personality: Neuroticism",
              row_number() < 193 ~ "Parental antisocial behaviour",
              row_number() < 197 ~ "Parental alcoholism",
              row_number() < 201 ~ "Child IQ",
              row_number() < 205 ~ "Child executive function",
              row_number() < 209 ~ "Child theory of mind",
              row_number() < 213 ~ "Child externalising behaviours" ,
              row_number() < 217 ~ "Child internalising behaviours",
              row_number() < 221 ~ "Child ADHD behaviours",
              row_number() > 220 ~ "Child prosocial behaviours")) %>%
  select(-X1) %>%
  mutate(
    block = case_when(
              row_number() < 25  ~ "Social domain",
              row_number() < 53 ~ "Home domain",
              row_number() < 85 ~ "Parent domain",
              row_number() < 97 ~ "Neuro domain",
              row_number() < 113 ~ "Emo/behave domain",
              row_number() < 137 ~ "Social domain",
              row_number() < 165 ~ "Home domain",
              row_number() < 197 ~ "Parent domain",
              row_number() < 209 ~ "Neuro domain",
              row_number() > 208 ~ "Emo/behave domain")) %>%
  spread(key, allmodel) %>%
  mutate(
    Significance = 
      if_else(
        p.value < 0.05, 
        1,
        0
      ) %>%
    recode_factor(
        "0" = "Non-significant",
        "1" = "Significant")) %>%
  transmute(Class = Class,
            Variable = Variable,
            RRR = as.numeric(RRR),
            RRR.conf.low = as.numeric(RRR.conf.low),
            RRR.conf.high = as.numeric(RRR.conf.high),
            p.value = as.numeric(p.value),
            Significance = Significance,
            Domain = block) 
```

#### Table 

```{r stata all table}
paper.table.all.variables <- paper.table(dat.stata.all)
paper.table.all.variables
```

```{r bind and order rows multivariate sig plot}
dat.stata.all.plot <- dat.stata.all %>% 
  filter(Significance == "Significant",
         Variable != "Child executive function")

individual.sig.social.plot <- order.variables(data = dat.stata.all.plot, domain = "Social domain")
individual.sig.parent.plot <- order.variables(data = dat.stata.all.plot, domain = "Parent domain")
individual.sig.emo.plot <- order.variables(data = dat.stata.all.plot, domain = "Emo/behave domain")

# join all
dat.stata.all.plot.ordered <- rbind(
  individual.sig.social.plot,
  individual.sig.parent.plot,
  individual.sig.emo.plot
)
```

Plot of significant variables from the multivariate analysis:

```{r stata plot full model}
stata.plot.all_model <- ggplot(
  dat.stata.all.plot.ordered, 
  aes(x = Variable,
      y = RRR,
      ymin = RRR.conf.low,
      ymax = RRR.conf.high,
      color = Class)
  ) +
  scale_fill_manual(values = c("black", "black")) +
  scale_color_manual(values = c("black", "black")) +
  scale_shape_manual(values = c(19)) +
  geom_pointrange(aes(shape = Significance), position = position_dodge(width = 0.7)) +
  facet_grid(Domain ~ Class, drop = TRUE, scales = "free_y", space = "free") +
  labs(y = "RRR (95% CI)",
       x = "names",
       title = "Association between antecedents and social isolation trajectories",
       subtitle = paste("N(Total) = ", length(dat$sex),
                        "; N(Increasing) = ", sum(!is.na(dat$class_renamed[dat$class_renamed=="Increasing"])),
                        "; N(Decreasing) = ", sum(!is.na(dat$class_renamed[dat$class_renamed=="Decreasing"])), sep = ""),
       color = "",
       shape = "") +
  theme.antecedent +
  theme(legend.position = "none") +
  geom_hline(linetype = "dashed", yintercept = 1) +
 scale_y_continuous(trans = scales::log10_trans(), breaks=c(0.5, 0.6,0.8, 1, 1.5, 2, 2.5, 3, 4)) +
  coord_flip()

stata.plot.all_model 
```

```{r save multivariate plot, include=FALSE}
# if(posterior.cut == FALSE){
# #save
# ggsave(
#   "multivariate_antecedent_STATA_SGDP_CENTRE.png",
#   plot = stata.plot.all_model,
#   device = "png",
#   path = paste0(graph_save_data_path, "antecedents/"),
#   width = 8,
#   height = 3
# )
# }
```

