---
title: "Mplus 3-Step sensitivity analysis interpretation"
output:  
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: flatly
    code_folding: hide
    includes:
      after_body: footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = NA,
                      prompt = FALSE,
                      cache = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      results = 'asis')

options(bitmapType = 'quartz') # to render fonts better
```

```{r Clear global environment}
remove(list = ls())
```

```{r colour palletes for graphs}
palette <- c("#78D9C5","#F5BE5E","#EEB6E9","#DBDB73","#FFED98","#BFD2EB")
```
 
```{r source the data file path}
#source raw data directory: data_raw and data included
source("../../../isolation_trajectories_data_path.R")
```

```{r Mplus packages}
library(MplusAutomation)
library(stringr)
library(stringi)
library(filesstrings)
library(texreg)
library(relimp)
library(rhdf5)
library(knitr)
library(ggplot2)
library(tidyverse)
library(plyr) #conflicts with tidyverse for e.g. rename and row_number
```

***

This page displays the results from the 3-step sensitivity analysis used to assess how social isolation trajectories are associated with outcomes and antecedents while accounting for classification errors in class trajectories. From this sensitivity analysis we conclude that there was very limited difference in accounting for class classification error rates (the results were replicated), therefore the hard classification was presented in the manuscript. 

***

```{r MplusAutomation citation, include=FALSE}
citation("MplusAutomation") # need the citation for the paper
```

***

# Functions 

To run the following code more clearly, I have created these functions:

* add.significance.variable() which assesses the p value and creates a new variable based on if the result is significant or not. 

```{r functions, class.source = 'fold-show'}
add.significance.variable <- function(data){
  data <- data %>%
    mutate(
    Significance = 
      if_else(
        p < 0.05, 
        1,
        0
      ) %>%
    recode_factor(
        "0" = "Non-significant",
        "1" = "Significant"
      ))
  return(data)
}
```


# Antecedents

In Mplus, to assess the association with antecedents, I manually added the code to account for Logits for the Classification Probabilities in the original GMM. 

```{bash, eval=FALSE, class.source = 'fold-show'}
    %OVERALL%
    
    C ON seswq35 P5CACORNCategoryrecoded vndngdm5 socprbm5 nchildren00e5 
	schmeals00e5 harm3em5recoded tsibl5 tssupm5 actvm5 nobiodl5 anyviom5 warme5 fdepmm5 
	bfiom5 bficm5 bfiem5 bfiam5 bfinm5 antisocialparent alcoholismparent iqe5 exfunce5 
	tomtote5 totexte5 intisoe5 totadde5 totproe5 ;
     
    %C#1%
     [class#1@5.092]; [class#2@-0.194];
     
    %C#2%
     [class#1@2.193]; [class#2@4.204];
     
    %C#3%
     [class#1@-2.233]; [class#2@-5.270];
```

*READING IN THE DATA IN RMD DOES NOT PROVIDE THE RIGHT INFORMATION FOR THE ANTECEDENTS - need to look at the Mplus .out file*
You can only access the unstandardised paratimisations in R. To find the results, go to the section: "ODDS RATIO FOR THE ALTERNATIVE PARAMETERIZATIONS FOR THE CATEGORICAL LATENT VARIABLE REGRESSION: Parameterization using Reference Class 1".For now I have manually copied over the correct table needed. 

```{bash, eval=FALSE, class.source = 'fold-show'}
ODDS RATIO FOR THE ALTERNATIVE PARAMETERIZATIONS FOR THE CATEGORICAL LATENT VARIABLE REGRESSION

                                         (Est. - 1) Two-Tailed
                    Estimate       S.E.     / S.E.    P-Value

Parameterization using Reference Class 1

 C#2      ON
    SESWQ35            0.886      0.247     -0.463      0.643
    P5CACORNCA         1.199      0.477      0.418      0.676
    VNDNGDM5           0.855      0.102     -1.417      0.157
    SOCPRBM5           1.139      0.144      0.965      0.335
    NCHILDREN0         1.000      0.001      0.315      0.753
    SCHMEALS00         1.000      0.015     -0.028      0.978
    HARM3EM5RE         1.160      0.526      0.304      0.761
    TSIBL5             0.848      0.128     -1.185      0.236
    TSSUPM5            0.981      0.026     -0.729      0.466
    ACTVM5             0.867      0.076     -1.746      0.081
    NOBIODL5           1.241      0.670      0.359      0.719
    ANYVIOM5           0.742      0.271     -0.950      0.342
    WARME5             0.981      0.160     -0.119      0.906
    FDEPMM5            1.093      0.425      0.220      0.826
    BFIOM5             0.996      0.030     -0.119      0.905
    BFICM5             0.994      0.034     -0.163      0.871
    BFIEM5             1.010      0.038      0.251      0.802
    BFIAM5             0.988      0.028     -0.433      0.665
    BFINM5             0.989      0.032     -0.358      0.721
    ANTISOCIAL         1.000      0.010     -0.043      0.966
    ALCOHOLISM         0.967      0.022     -1.521      0.128
    IQE5               1.003      0.012      0.289      0.772
    EXFUNCE5           1.098      0.052      1.895      0.058
    TOMTOTE5           0.936      0.045     -1.427      0.154
    TOTEXTE5           1.020      0.014      1.399      0.162
    INTISOE5           1.056      0.035      1.603      0.109
    TOTADDE5           1.064      0.016      3.948      0.000
    TOTPROE5           1.000      0.030      0.002      0.998

 C#3      ON
    SESWQ35            1.000      0.391     -0.001      1.000
    P5CACORNCA         0.286      0.244     -2.933      0.003
    VNDNGDM5           0.851      0.158     -0.944      0.345
    SOCPRBM5           0.907      0.206     -0.453      0.651
    NCHILDREN0         1.004      0.002      2.406      0.016
    SCHMEALS00         1.029      0.024      1.221      0.222
    HARM3EM5RE         1.300      0.631      0.476      0.634
    TSIBL5             1.216      0.161      1.341      0.180
    TSSUPM5            0.964      0.040     -0.911      0.362
    ACTVM5             1.019      0.105      0.181      0.856
    NOBIODL5           0.876      0.532     -0.233      0.816
    ANYVIOM5           0.704      0.411     -0.720      0.472
    WARME5             1.160      0.220      0.727      0.467
    FDEPMM5            1.187      0.548      0.342      0.733
    BFIOM5             1.092      0.037      2.493      0.013
    BFICM5             0.981      0.030     -0.648      0.517
    BFIEM5             0.946      0.044     -1.205      0.228
    BFIAM5             1.004      0.046      0.094      0.925
    BFINM5             1.001      0.054      0.018      0.985
    ANTISOCIAL         1.003      0.012      0.258      0.796
    ALCOHOLISM         1.008      0.036      0.233      0.816
    IQE5               0.986      0.016     -0.869      0.385
    EXFUNCE5           0.886      0.046     -2.505      0.012
    TOMTOTE5           1.020      0.075      0.271      0.786
    TOTEXTE5           1.023      0.020      1.153      0.249
    INTISOE5           1.389      0.092      4.238      0.000
    TOTADDE5           1.056      0.021      2.643      0.008
    TOTPROE5           0.908      0.033     -2.755      0.006
```


# Outcomes {.tabset .tabset-fade}

We conducted the outcome associations two ways. First, the [Auxillary DE3STEP method](https://www.statmodel.com/examples/webnotes/webnote15.pdf), then the more recently developed [BCH approach](https://www.statmodel.com/examples/webnotes/webnote21.pdf). 

## Auxillary DE3STEP method 

First we need to read in the output file from Mplus, then extract all the means per variable. 

```{r read in data for outcome results, results='hide'}
outcome.model <- readModels(target = paste0(mplus_outcome_output_3step), recursive = TRUE)
```

```{r extract overall class means per variable}
# extract overall means for each class per variable
outcome.model_overall_means <- outcome.model[["X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.3step.output.outcomes..isolation_3traj_full_sample_clustered_3STEP_DE3STEP_outcomes.out"]][["lcCondMeans"]][["overall"]]

colnames(outcome.model_overall_means) <- c("Variable",
                                           "Low stable class mean",
                                           "Low stable class standard error",
                                           "Increasing class mean",
                                           "Increasing class standard error",
                                           "Decreasing class mean",
                                           "Decreasing class standard error",
                                           "Chi square",
                                           "df",
                                           "p value") 
outcome.model_overall_means
```

We are not interested in comparison from class 2 to 3, ware only interested in the comparisons to the low stable class (class 1). Significant results have been presented here, split by increasing and decreasing class, respectively.  

We also ran the analysis using the "DU3STEP" method in Mplus. But the DU3STEP option showed errors such as: PROBLEMS OCCURRED DURING THE ESTIMATION FOR THE DISTAL OUTCOME NEETE18. THE VARIANCE OF THE DISTAL OUTCOME IN CLASS 1 IS ZERO. THE PROBLEM CAN BE RESOLVED BY USING THE DE3STEP OPTION. **Therefore the DE3STEP option was used instead**

```{r extract pairwise comparisons - auxiliary}
outcome.model_all_pairwise_tests <- outcome.model[["X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.3step.output.outcomes..isolation_3traj_full_sample_clustered_3STEP_DE3STEP_outcomes.out"]][["lcCondMeans"]][["pairwise"]]

# Overall table
outcome.model_pairwise_tests <- outcome.model_all_pairwise_tests %>% 
  filter(classA == 1) %>%    # only get low stable comparisons
  select(!df) %>%            # remove df column (it was empty)
  add.significance.variable 
        
# Increasing comparisons - significant only
outcome.model_pairwise_tests_significant_increasing <- outcome.model_pairwise_tests %>%
  filter(classB == 2 & Significance == "Significant") %>%
  mutate(`Class comparison` = c("Increasing")) %>%
  select(Variable = var,
         `Class comparison`,
         `Chi square` = chisq,
         p)
outcome.model_pairwise_tests_significant_increasing
        
# Decreasing comparisons - significant only
outcome.model_pairwise_tests_significant_decreasing <- outcome.model_pairwise_tests %>%
  filter(classB == 3 & Significance == "Significant") %>%
  mutate(`Class comparison` = c("Decreasing")) %>%
  select(Variable = var,
         `Class comparison`,
         `Chi square` = chisq,
         p)
outcome.model_pairwise_tests_significant_decreasing

# All significant results
outcome.model_pairwise_tests_significant <- rbind(outcome.model_pairwise_tests_significant_increasing, 
                                                  outcome.model_pairwise_tests_significant_decreasing)
#outcome.model_pairwise_tests_significant
```

*Summary of variables that are significant using pairwise t-tests: Using "DE3STEP" in Mplus:* (Y indicates same as original analysis)

Increasing:

* Moderate conduct disorder diagnosis Y
* Not in employment or education Y
* ADHD symptoms Y
* Conduct disorder symptoms Y
* Loneliness Y
* Technology use Y
* Coping with stress Y 
* Sleep Y
* Job preparedness: attributes Y

Decreasing:

* Depression symptoms Y

Both:

* Psychotic experiences Y
* Physical activity Y
* Life satisfaction Y (but only for increasing)
* Job optimism Y

Missing but reported in original analysis:

* Self harm or suicide attempt
* Cannabis dependence - DUE TO ERROR ACCOUNTED IN MPLUS
* ADHD diagnosis 
* Service use


## BCH approach

Similar to the auxilary DE3STEP approach, the [BCH approach](https://www.statmodel.com/examples/webnotes/webnote21.pdf):

The BCH method, BCH for short, described in Vermunt (2010) and also in Bakk et al. (2013). For the distal outcome model that evaluates the means across classes for a continuous auxiliary variable these simulations show that the BCH method substantially outperforms Lanza’s method and the 3-step method. The BCH method avoids shifts in latent class in the final stage that the 3-step method is susceptible to. In its final stage the BCH method uses a weighted multiple group analysis, where the groups correspond to the latent classes, and thus the class shift is not possible because the classes are known. In addition, the BCH method performs well when the variance of the auxiliary variable differs substantially across classes, i.e., resolving the problems that Lanza’s method is susceptible to. The BCH method uses weights wij which reflect the measurement error of the latent class variable. In the estimation of the auxiliary model, the i-th observation in class/group j is assigned a weight of wij and the auxiliary model is estimated as a multiple group model using these weights. 

No errors were thrown when running this BCH model - all variables were thus included in the analyses. 

```{r extract pairwise comparisons - BCH}
outcome.model_all_pairwise_tests_bch <- outcome.model[["X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.3step.output.outcomes..isolation_3traj_full_sample_clustered_3STEP_BCH_outcomes.out"]][["lcCondMeans"]][["pairwise"]]

# Overall table
outcome.model_pairwise_tests_bch <- outcome.model_all_pairwise_tests_bch %>% 
  filter(classA == 1) %>%    # only get low stable comparisons
  select(!df) %>%            # remove df column (it was empty)
  add.significance.variable 


# Increasing comparisons - significant only
outcome.model_pairwise_tests_increasing_bch <- outcome.model_pairwise_tests_bch %>%
  filter(classB == 2) %>%
  mutate(`Class comparison` = c("Increasing")) %>%
  select(Variable = var,
         `Class comparison`,
         `Chi square` = chisq,
         p)
        
# Decreasing comparisons - significant only
outcome.model_pairwise_tests_decreasing_bch <- outcome.model_pairwise_tests_bch %>%
  filter(classB == 3) %>%
  mutate(`Class comparison` = c("Decreasing")) %>%
  select(Variable = var,
         `Class comparison`,
         `Chi square` = chisq,
         p)

# All significant results
outcome.model_pairwise_tests_bch_full <- rbind(outcome.model_pairwise_tests_increasing_bch, 
                                                  outcome.model_pairwise_tests_decreasing_bch)


kable(outcome.model_pairwise_tests_bch_full) # compare to auxilary 
# kable(outcome.model_pairwise_tests_significant)
```

*Summary of variables that were significant using pairwise t-tests: Using "BCH" in Mplus* (Y indicates same as original analysis)

Increasing:

* ADHD diagnosis Y
* Moderate conduct disorder diagnosis Y
* Cannabis dependence Y (both in original)
* Self harm or suicide attempt Y (both in original)
* Not in employment or education Y
* ADHD symptoms Y
* Conduct disorder symptoms Y
* Loneliness Y
* Technology use Y
* Coping with stress Y 
* Sleep Y
* Job preparedness: attributes Y

Decreasing:

* Depression symptoms Y
* Cannabis symptoms (diagnosis in original analysis)

Both:

* Psychotic experiences Y
* Physical activity Y
* Life satisfaction Y (but only for increasing)
* Job optimism Y
* Smoking (increasing in original)

Missing here but included in original:

* Service use (increasing)
