---
title: "Data preprosessing script for social isolation trajectories"
author: "K N Thompson"
date: Sys.Date()
output:  
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: cerulean
code_folding: show

html_notebook:
  theme: cerulean
toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      comment = NA,
                      prompt = FALSE,
                      cache = FALSE)


options(bitmapType = 'quartz') # to render fonts better
```

Clear global environment
```{r Clear global environment}
remove(list = ls())
```

Retrieve the current date to use it for file endings to not overwrite files when one exports files
```{r Recent date}
date = Sys.Date()
```

# Packages

```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(foreign) # Needed to load spss data files
library(tidyverse)
```

Source the data file path - all data is saved under one path. If you need access to this data_path please contact Katherine N Thompson on katherine.n.thompson@kcl.ac.uk. 

```{r source data file path}
#source raw data directory: data_raw and data included
source("../../../../../isolation_trajectories_data_path.R")
```

# Read in data

Read in full SPSS sav file.

```{r read in data (spss)}
dat <- read.spss(file = paste0(data.raw_path, "Katie_18Dec20.sav"),
                  use.value.labels = FALSE, #convert value labels into factors with those labels
                  to.data.frame = TRUE) #return data frame 

colnames(dat) # check colnames
```

# Item lists

Only using elder twin variables as each twin is included in the data set as an individual. 
Without doing twin analyses, the elder and younger variables show the same information. 

```{r prepare item lists}
#All the items in numeric form
combined_social_isolation <- c("sisoe5",
                              "sisoe7",
                              "sisoe10",
                              "sisoe12")
```

# Rename ID variable

When doing non-twin analyses - the atwin and btwin id give the same information.

```{r rename ID}
dat <- dat %>%
  mutate(
    id = atwinid
  )

dat <- dat %>%
  mutate(
    randomised_order = rorderp5
  )
#check
colnames(dat)
```

# Demographic variables

All variables below have been recoded into a factor for R purposes and renamed. Only the categorical variables need recoding to establish the levels of the variable. 

```{r twin order factor}
# twin order
dat <- dat %>%
  mutate(
    twin_order = 
      recode_factor(torder,
        "1" = "Elder",
        "2" = "Younger"))
```

```{r sample group factor}
# sample groups
dat <- dat %>%
  mutate(
    risk = # this represents mothers who had their first child under 20 years old
      recode_factor(risks,
        "0" = "Low risk",
        "1" = "High risk"))
```

```{r cohort factor}
# cohort
dat <- dat %>%
  mutate(
    cohort_binary = 
      recode_factor(cohort,
        "94" = "Born in 1994",
        "95" = "Born in 1995"))
```

```{r sex factor}
# sex
dat <- dat %>%
  mutate(
    sex = 
      recode_factor(sampsex,
        "1" = "Male",
        "2" = "Female"))
```

```{r zygosity factor}
# zygosity
dat <- dat %>%
  mutate(
    zygosity_binary = 
      recode_factor(zygosity,
        "1" = "MZ",
        "2" = "DZ"))
```

```{r SES factor}
# SES
dat <- dat %>%
  mutate(
    SES = 
      recode_factor(seswq35,
        "1" = "Low",
        "2" = "Middle", #this was missing in original SPSS file but still have 2s in the data set
        "3" = "High"))
```

Need to create a variable that has ordered SES to show that there are order differences between the labels. 

```{r SES ordered variable}
dat <- dat %>%
  mutate(
    SES_ordered = 
      ordered(SES,
              levels = c("Low",
                         "Middle",
                         "High")
      )
  )
```

```{r ethnicity factor}
# ethnicity
dat <- dat %>%
  mutate(
    ethnicity = 
      recode_factor(sethnic,
        "1" = "White",
        "2" = "Asian", 
        "3" = "Black",
        "4" = "Mixed race",
        "5" = "Other"))
```

# Social factors at age 5

```{r acorn factor}
# acorn_05
dat <- dat %>%
  mutate(
    acorn_05 = 
      recode_factor(P5CACORNCategory,
        "1" = "Wealthy Achievers",
        "2" = "Urban Prosperity",
        "3" = "Comfortably Off",
        "4" = "Moderate Means",
        "5" = "Hard Pressed"))
```

# Home environment at age 5

```{r mother not lived with biodad coding}
# mum_notlived_biodad_sincebirth_05
dat <- dat %>%
  mutate(
    mum_notlived_biodad_sincebirth_05 =
      recode_factor(nobiodl5,
                    "0" = "No",
                    "1" = "Yes")
  )
```

```{r any domestic violence coding}
# any_domestic_violence_05
dat <- dat %>%
  mutate(
    any_domestic_violence_05 =
      recode_factor(anyviom5,
                    "0" = "No",
                    "1" = "Yes")
  )
```

```{r child harm coding}
# child_harm_05
dat <- dat %>%
  mutate(
    child_harm_05 =
      recode_factor(harm3em5,
                    "0" = "No harm",
                    "1" = "Possible harm",
                    "2" = "Definite harm")
  )
```

```{r maternal warmth coding}
# maternal_warmth_05
dat <- dat %>%
  mutate(
    maternal_warmth_05 =
      recode_factor(warme5,
                    "0" = "No warmth",
                    "1" = "Very little warmth",
                    "2" = "Some warmth",
                    "3" = "Moderate warmth",
                    "4" = "Moderately high warmth",
                    "5" = "High warmth")
  )
```

# Parent characterisrics at age 5

```{r maternal depression lifetime}
# maternal_depression_lifetime_05
dat <- dat %>%
  mutate(
    maternal_depression_lifetime_05 =
      recode_factor(fdepmm5,
                    "0" = "No",
                    "1" = "Yes")
  )
```

# Mental health and service use at age 18

```{r depression diagnosis in last 12 months}
# depression_diagnosis_12mo_18
dat <- dat %>%
  mutate(
    depression_diagnosis_12mo_18 =
      recode_factor(dxmdee18,
                    "0" = "No",
                    "1" = "Yes")
  )
```

```{r anxiety diagnosis}
# anxiety_diagnosis_12mo_18
dat <- dat %>%
  mutate(
    anxiety_diagnosis_18 =
      recode_factor(dxgade18,
                    "0" = "No",
                    "1" = "Yes")
  )
```

```{r ADHD diagnosis}
# ADHD_diagnosis_12mo_18
dat <- dat %>%
  mutate(
    ADHD_diagnosis_18 =
      recode_factor(dxadhd5_18e,
                    "0" = "No",
                    "1" = "Yes")
  )
```

```{r conduct disorder diagnosis(???)}
# conduct_disorder_moderate_18
dat <- dat %>%
  mutate(
    conduct_disorder_moderate_18 =
      recode_factor(cdmode18,
                    "0" = "No",
                    "1" = "Yes")
  )
```

```{r alcohol dependence}
# alcohol_dependence_18
dat <- dat %>%
  mutate(
    alcohol_dependence_18 =
      recode_factor(dxalcdepe18,
                    "0" = "No",
                    "1" = "Yes")
  )
```

```{r alcohol abuse}
# alcohol_abuse_18
dat <- dat %>%
  mutate(
    alcohol_abuse_18 =
      recode_factor(dxalcabue18,
                    "0" = "No",
                    "1" = "Yes")
  )
```

```{r cannabis dependence}
# cannabis_dependence_18
dat <- dat %>%
  mutate(
    cannabis_dependence_18 =
      recode_factor(dxmarje18,
                    "0" = "No",
                    "1" = "Yes")
  )
```

```{r PTSD diagnosis lifetime}
# PTSD_diagnosis_lifetime_18
dat <- dat %>%
  mutate(
    PTSD_diagnosis_lifetime_18 =
      recode_factor(dxptsd5lfe18,
                    "0" = "No",
                    "1" = "Yes")
  )
```

```{r PTSD diagnosis current}
# PTSD_diagnosis_current_18
dat <- dat %>%
  mutate(
    PTSD_diagnosis_current_18 =
      recode_factor(dxptsd5cue18,
                    "0" = "No",
                    "1" = "Yes")
  )
```

```{r pychosis symptom count categorical}
# cat_psychosis_symptom_count_18
dat <- dat %>%
  mutate(
    cat_psychosis_symptom_count_18 =
      recode_factor(psysymp01e18,
                    "0" = "No symptoms",
                    "1" = "One or more symptoms")
  )
```

```{r categorical pychosis symptom count categorical}
# cat_psychosis_experiences_scale_18
dat <- dat %>%
  mutate(
    cat_psychosis_experiences_scale_18 =
      recode_factor(psyexpce18,
                    "0" = "No symptoms",
                    "1" = "One to two symptoms",
                    "2" = "Three to five symptoms",
                    "3" = "Six or more symptoms")
  )
```

```{r suicide attempt or self harm}
# suicide_attempt_selfharm_18
dat <- dat %>%
  mutate(
    suicide_attempt_selfharm_18 =
      recode_factor(sharmsuice18,
                    "0" = "No",
                    "1" = "Yes")
  )
```

```{r service use}
# service_use_18
dat <- dat %>%
  mutate(
    service_use_18 =
      recode_factor(srvusemhe18,
                    "0" = "No",
                    "1" = "Yes")
  )
```

# Physical health and health risks at age 18
*We decided to keep this as a continuous scale rather than categories - have hashed out*
```{r physical activity}
# physical_activity_18
# dat <- dat %>%
#   mutate(
#     physical_activity_18 =
#       recode_factor(phyacte18,
#                     "1" = "Inactive",
#                     "2" = "Light intensity",
#                     "3" = "Moderate intensity",
#                     "4" = "Hard intensity",
#                     "5" = "Very hard intensity")
#   )
```

# Employment prospects

```{r not in education, employment or training}
# not_in_employment_education_18
dat <- dat %>%
  mutate(
    not_in_employment_education_18 =
      recode_factor(neete18,
                    "0" = "No",
                    "1" = "Yes")
  )
```

```{r highest level of education}
# highest_education_18
dat <- dat %>%
  mutate(
    highest_education_18 =
      recode_factor(educachve18,
                    "0" = "No qualification",
                    "1" = "Level 1 (GCSE at grades D-G)",
                    "2" = "Level 2 (GCSE at grades A*-C)",
                    "3" = "Level 3 (A Level)"
                    )
  )
```

# Variable REcode based on analysis decisions

Some variables were edited further dependent on power and other factors. 

```{r ACORN recode, include=FALSE}
dat <- dat %>%
  mutate(acorn_recoded_05 =
           if_else(
             acorn_05 == "Moderate Means" | #Moderate Means and Hard Pressed were combined, Wealthy Achievers, Urban Prosperity, and Comfortably Off were combined
               acorn_05 == "Hard Pressed",
             1,
             0
           ) %>%
           recode_factor(
             "1" = "Deprived",
             "0" = "Relatively affluent"
           ) %>%
           relevel(ref = "Relatively affluent",
            first = TRUE, #levels in ref come first
            collapse = "+", #String used when constructing names for combined factor levels
            xlevels = TRUE) #levels maintained even if not actually occurring
      )
table(dat$acorn_recoded_05)

dat <- dat %>%
  mutate(
    acorn_continuous_05 =
      as.numeric(acorn_05)
  )
```

To have a binary option for SES, middle and high SES were combined. 

```{r SES recode - combine middle and high, include=FALSE}
dat <- dat %>%
  mutate(
    SES_binary_releveled =
      as.factor(fct_collapse(
        SES,
        "Middle to high" = c("Middle","High"),
        "Low" = c("Low")
      )
  ) %>%
    relevel(ref = "Middle to high",
            first = TRUE, #levels in ref come first
            collapse = "+", #String used when constructing names for combined factor levels
            xlevels = TRUE) #levels maintained even if not actually occurring
  )
table(dat$SES_binary_releveled)
```

Possible harm and Definite harm were combined to give a binary variable of Harm VS No harm.

```{r child harm recode, include=FALSE}
dat <- dat %>%
  mutate(child_harm_recoded_05 =
           if_else(
             child_harm_05 == "Possible harm" |
               child_harm_05 == "Definite harm",
             1,
             0
           ) %>%
           recode_factor(
             "1" = "Harm",
             "0" = "No harm"
           ) %>%
           relevel(ref = "No harm",
            first = TRUE, #levels in ref come first
            collapse = "+", #String used when constructing names for combined factor levels
            xlevels = TRUE)
      )
table(dat$child_harm_recoded_05)
```

Originally maternal warmth was condensed - this resulted in a large reduction in power. Therefore, maternal warmth was recoded to be continuous. 

```{r maternal warmth recode, include=FALSE}
# dat <- dat %>%
#   mutate(
#     maternal_warmth_recoded_05 =
#       as.factor(fct_collapse(
#         maternal_warmth_05,
#         "Little warmth" = c("No warmth","Very little warmth"),
#         "Some warmth" = c("Some warmth","Moderate warmth"),
#         "High warmth" = c("Moderately high warmth","High warmth")
#       )
#   ) %>%
#     relevel(ref = "Some warmth",
#         first = TRUE, #levels in ref come first
#         collapse = "+", #String used when constructing names for combined factor levels
#         xlevels = TRUE #levels maintained even if not actually occurring
#   ))
# table(dat$maternal_warmth_recoded_05)

dat <- dat %>%
  mutate(
    maternal_warmth_continuous_05 = 
      as.numeric(maternal_warmth_05)
  )
```

GSCE grades were collapsed: No qualification and Level 1 (GCSE at grades D-G) were combined. Level 2 (GCSE at grades A*-C) and Level 3 (A Level) were combined. 

```{r highest education recode, include=FALSE}
dat <- dat %>%
  mutate(
    highest_education_recode_18 =
      as.factor(fct_collapse(
        highest_education_18,
        "Level 1 or below (GCSE grades D-G or below)" = c("No qualification","Level 1 (GCSE at grades D-G)"),
        "Level 2 or 3 (GCSE grades A*-C or A-level)" = c("Level 2 (GCSE at grades A*-C)","Level 3 (A Level)")
      )
  ))
table(dat$highest_education_recode_18)
```

We combined maternal and paternal reports to form parental report for antisocial behaviour and alcoholism. 

```{r recode antisocial behaviour for parent}
# create total score for mum and data to make parent
dat <- dat %>%
  mutate(antisocial_behaviour_parent_05 =
           asbmm5 + asbfm5) # mum plus dad scores
```

```{r recode alcoholism for parent}
# create total score for mum and data to make parent
dat <- dat %>%
  mutate(alcoholism_parent_05 =
           alcmm5 + alcfm5) # mum plus dad scores
```

# Numeric versions for categorical variables

Create numeric versions of all categorical antecedent and outcome variables. This is done purely for the correlation matrix to calculate multicollinearity between variables. 

```{r numreic varions of categorical antecedent variables}
dat <- dat %>%
  mutate(SES_binary_releveled_numeric = as.numeric(SES_binary_releveled)) %>%
  mutate(child_harm_recoded_numeric_05 = as.numeric(child_harm_recoded_05)) %>%
  mutate(mum_notlived_biodad_sincebirth_numeric_05 = as.numeric(mum_notlived_biodad_sincebirth_05)) %>%
  mutate(any_domestic_violence_numeric_05 = as.numeric(any_domestic_violence_05)) %>%
  mutate(maternal_depression_lifetime_numeric_05 = as.numeric(maternal_depression_lifetime_05))
```

```{r numreic varions of categorical outcome variables}
dat <- dat %>%
  mutate(depression_diagnosis_12mo_numeric_18 = as.numeric(depression_diagnosis_12mo_18)) %>%
  mutate(anxiety_diagnosis_numeric_18 = as.numeric(anxiety_diagnosis_18)) %>%
  mutate(ADHD_diagnosis_numeric_18 = as.numeric(ADHD_diagnosis_18)) %>%
  mutate(conduct_disorder_moderate_numeric_18 = as.numeric(conduct_disorder_moderate_18)) %>%
  mutate(alcohol_dependence_numeric_18 = as.numeric(alcohol_dependence_18)) %>%
  mutate(alcohol_abuse_numeric_18 = as.numeric(alcohol_abuse_18)) %>%
  mutate(cannabis_dependence_numeric_18 = as.numeric(cannabis_dependence_18)) %>%
  mutate(PTSD_diagnosis_lifetime_numeric_18 = as.numeric(PTSD_diagnosis_lifetime_18)) %>%
  mutate(PTSD_diagnosis_current_numeric_18 = as.numeric(PTSD_diagnosis_current_18)) %>%
  mutate(cat_psychosis_experiences_scale_numeric_18 = as.numeric(cat_psychosis_experiences_scale_18)) %>%
  mutate(suicide_attempt_selfharm_numeric_18 = as.numeric(suicide_attempt_selfharm_18)) %>%
  mutate(service_use_numeric_18 = as.numeric(service_use_18)) %>%
  mutate(not_in_employment_education_numeric_18 = as.numeric(not_in_employment_education_18)) 
```

# Social isolation 

We needed to create several forms of the social isolation variables. 

## Log transformed social isolation

```{r log transform social isolation combined variables}
colnames(dat)
dat <- dat %>%
  mutate(log_sisoe5 =
           log1p(sisoe5))
dat <- dat %>%
  mutate(log_sisoe7 =
           log1p(sisoe7))
dat <- dat %>%
  mutate(log_sisoe10 =
           log1p(sisoe10))
dat <- dat %>%
  mutate(log_sisoe12 =
           log1p(sisoe12))

#check
colnames(dat)
```

## Binary social isolation

Using the same categorical cut offs as previously used in E-risk:
- social isolation <= to 1, give it a 0.
- social isolation > 1 and <= 2, give it a 1
- social isolation > 2, give it a 2. 

```{r binary social isolation variable}
dat <- dat %>%
  mutate(
    isolation.categorical.5 =
      case_when(sisoe5 > 2 ~ 2,
                sisoe5 > 1 & sisoe5 <= 2 ~ 1,
                sisoe5 <= 1 ~ 0)
  ) %>%
  mutate(
    isolation.categorical.7 =
      case_when(sisoe7 > 2 ~ 2,
                sisoe7 > 1 & sisoe7 <= 2 ~ 1,
                sisoe7 <= 1 ~ 0)
  ) %>%
  mutate(
    isolation.categorical.10 =
      case_when(sisoe10 > 2 ~ 2,
                sisoe10 > 1 & sisoe10 <= 2 ~ 1,
                sisoe10 <= 1 ~ 0)
  ) %>%
  mutate(
    isolation.categorical.12 =
      case_when(sisoe12 > 2 ~ 2,
                sisoe12 > 1 & sisoe12 <= 2 ~ 1,
                sisoe12 <= 1 ~ 0)
  ) 

# check
table(dat$isolation.categorical.5)
table(dat$isolation.categorical.7)
table(dat$isolation.categorical.10)
table(dat$isolation.categorical.12)
```

# Missing values

Make sue that any values coded as -9, -999 , -8, -7 or -1 are coded as NA

```{r recode missing values}
# Recode missing values to NAs in numeric variables
dat <- dat %>%
  mutate_if(is.numeric, ~na_if(., -9)) %>% # missing
  mutate_if(is.numeric, ~na_if(., -999)) %>% # missing
  mutate_if(is.numeric, ~na_if(., -1))  %>% # no question
  mutate_if(is.numeric, ~na_if(., -8))  %>% # missing specific to the question
  mutate_if(is.numeric, ~na_if(., -7)) # missing specific to the question
```

## Missingness for social isolation variables

```{r social isolation missingness}
# create the NA per person variable
dat <- dat %>%
  mutate(
    na.per.person.si =
      rowSums(
        is.na(dat[,colnames(dat) %in% combined_social_isolation])
      )
  )
```

# Create data set exluding those with 3 isolation data points missing 

```{r remove those that have 3 missing data points on combined isolation}
# only select those who who have less than 3 combined isolation variables missing
dat.3missing <- dat %>%
  filter(
    na.per.person.si < 3
  )

# check the new number of rows int he data set - 2079
nrow(dat.3missing)
```

# Select all renamed variables
Remember to select which data set you want exported - dat (full sample), or dat.3missing (those with 3 missing time points excluded)

Not using the following variables in analysis:
       acorn_05,
       child_harm_05,
       maternal_warmth_05,
       highest_education_18,
       psychosis_symptom_count_18, 
       cat_psychosis_symptom_count_18,
       psychosis_experiences_scale_18,
       isolation_mother_05,
       isolation_mother_07,
       isolation_mother_10,
       isolation_mother_12,
       isolation_teacher_05,
       isolation_teacher_07,
       isolation_teacher_10,
       isolation_teacher_12
       
## Data with those missing 3 time points excluded

```{r rename social isolation variables for 3 missing dataset}
# variables split by blocks
# those with an "=" are just renamed with no other coding needed. Coding all done above. 
# Have # the variables that aren't being used in analyses

dat.renamed.3missing <- dat.3missing %>%
  select(
    # Demographics
    id,
    familyid,
    sex,
    random_order = rorderp5,
    zygosity_binary,
    ethnicity,
    risk,
    cohort_binary,
    twin_order,
    # Social factors
    acorn_05,
    acorn_recoded_05,
    acorn_continuous_05,
    vandalism_05 = vndngdm5, # original data frame only reported 0 = low and 6 = high
    problems_neighbours_05 = socprbm5, # original data frame only reported 0 = low and 6 = high
    SES,
    SES_ordered,
    SES_binary_releveled,
    SES_binary_releveled_numeric,
    number_children_school_05 = nchildren00e5,
    number_children_school_free_meals_05 = schmeals00e5,
    class_size_average_05 = classsize00e5,
    # Home environment
    resident_moves_05 = nmovel5,
    mum_notlived_biodad_sincebirth_05,
    mum_notlived_biodad_sincebirth_numeric_05,
    total_siblings_05 = tsibl5,
    any_domestic_violence_05,
    any_domestic_violence_numeric_05,
    child_harm_05,
    child_harm_recoded_05,
    child_harm_recoded_numeric_05,
    total_social_support_05 = tssupm5,
    total_activities_with_mum_05 = actvm5,
    maternal_warmth_05,
    maternal_warmth_continuous_05,
    prosocial_behaviours_combined_05 = totproe5,
    # Parent characteristics
    maternal_depression_lifetime_05,
    maternal_depression_lifetime_numeric_05,
    maternal_personality_openness_05 = bfiom5,
    maternal_personality_conscientiousness_05 = bficm5,
    maternal_personality_extroversion_05 = bfiem5,
    maternal_personality_agreeableness_05 = bfiam5,
    maternal_personality_neuroticism_05 = bfinm5,
    antisocial_behaviour_mum_05 = asbmm5,
    antisocial_behaviour_dad_05 = asbfm5,
    antisocial_behaviour_parent_05,
    alcoholism_mum_05 = alcmm5,
    alcoholism_dad_05 = alcfm5,
    alcoholism_parent_05,
    # Child neurodevelopment
    IQ_05 = iqe5,
    executive_function_05 = exfunce5,
    theory_of_mind_05 = tomtote5,
    # Child emotional and behavioural development
    externalising_combined_05 = totexte5,
    internalising_combined_excl_sis_05 = intisoe5,
    ADHD_combined_05 = totadde5,
    temp_negative_affect_05 = irre5,
    temp_impulsivity_05 = impe5,
    temp_approach_05 = appe5,
    temp_sluggishness_05 = slue5,
    temp_wariness_05 = ware5,
    temp_undercontrolled_05 = unce5,
    temp_inhibited_05 = inhe5,
    temp_shy_05 = shye5,
    ## Age 18
    # Mental health and service use
    depression_diagnosis_12mo_18,
    depression_diagnosis_12mo_numeric_18,
    depression_current_scale_18 = mdesxe18, 
    anxiety_diagnosis_18,
    anxiety_diagnosis_numeric_18,
    anxiety_current_scale_18 = gadsxe18, 
    ADHD_diagnosis_18,
    ADHD_diagnosis_numeric_18,
    inattentive_hyperactive_symptoms_total_18 = SR_symtot18e,
    conduct_disorder_moderate_18, 
    conduct_disorder_moderate_numeric_18,
    conduct_disorder_symptoms_18 = cdsxe18,
    alcohol_dependence_18,
    alcohol_dependence_numeric_18,
    alcohol_abuse_18,
    alcohol_abuse_numeric_18,
    alcohol_symptom_scale_18 = alcsxe18,
    cannabis_dependence_18,
    cannabis_dependence_numeric_18,
    cannabis_symptom_scale_18 = marjsxe18,
    PTSD_diagnosis_lifetime_18,
    PTSD_diagnosis_lifetime_numeric_18,
    PTSD_diagnosis_current_18,
    PTSD_diagnosis_current_numeric_18,
  #  psychosis_symptom_count_18 = psysympe18,
  #  cat_psychosis_symptom_count_18,
  #  psychosis_experiences_scale_18 = psyexpe18,
    cat_psychosis_experiences_scale_18, # only including one psychosis variable for now 
    cat_psychosis_experiences_scale_numeric_18,
    suicide_attempt_selfharm_18,
    suicide_attempt_selfharm_numeric_18,
    service_use_18,
    service_use_numeric_18,
    # Physical health and health risks
    BMI_18 = bmie18,
    CRP_log_18 = lnCRP_E18_4SD,
    physical_activity_18 = phyacte18,
    smoking_current_number_18 = smkcnume18,
    # Coping and functioning
    loneliness_18 = lonelye18,
    life_satisfaction_18 = lifsate18,
    technology_use_18 = teche18,
    coping_with_stress_18 = copstrse18,
    PSQI_global_score_18 = psqie18,
    #Employment prospects
    not_in_employment_education_18,
    not_in_employment_education_numeric_18,
    highest_education_18,
    highest_education_recode_18,
    job_preparedness_skills_18 = jprepse18,
    job_preparedness_attributes_18 = jprepae18,
    optimism_18 = optime18,
    job_search_activities_count_18 = jbschacte18,
    ## Social isolation
    isolation_combined_05 = sisoe5, # only ELDER variables included here - need to add a chunk to select younger variables if decide to do twin analyses. 
    isolation_combined_07 = sisoe7,
    isolation_combined_10 = sisoe10,
    isolation_combined_12 = sisoe12,
    isolation_mother_05 = sisoem5,
    isolation_mother_07 = sisoem7,
    isolation_mother_10 = sisoem10,
    isolation_mother_12 = sisoem12,
    isolation_teacher_05 = sisoet5,
    isolation_teacher_07 = sisoet7,
    isolation_teacher_10 = sisoet10,
    isolation_teacher_12 = sisoet12,
    log_isolation_combined_05 = log_sisoe5, # log values needed for trajectory analyses
    log_isolation_combined_07 = log_sisoe7,
    log_isolation_combined_10 = log_sisoe10,
    log_isolation_combined_12 = log_sisoe12,
    isolation.categorical.5,
    isolation.categorical.7,
    isolation.categorical.10,
    isolation.categorical.12,
    na.per.person.si
  )
# check
colnames(dat.renamed.3missing)
```

## Full data set

```{r rename social isolation variables for full dataset}
# same order as the variable list for the concept paper
# those with an "=" are just renamed with no other coding needed. Coding all done above. 
dat.renamed.full <- dat %>%
  select(
    # Demographics
    id,
    familyid,
    sex,
    random_order = rorderp5,
    zygosity_binary,
    ethnicity,
    risk,
    cohort_binary,
    twin_order,
    # Social factors
    acorn_05,
    acorn_recoded_05,
    acorn_continuous_05,
    vandalism_05 = vndngdm5, # original data frame only reported 0 = low and 6 = high
    problems_neighbours_05 = socprbm5, # original data frame only reported 0 = low and 6 = high
    SES,
    SES_ordered,
    SES_binary_releveled,
    SES_binary_releveled_numeric,
    number_children_school_05 = nchildren00e5,
    number_children_school_free_meals_05 = schmeals00e5,
    class_size_average_05 = classsize00e5,
    # Home environment
    resident_moves_05 = nmovel5,
    mum_notlived_biodad_sincebirth_05,
    mum_notlived_biodad_sincebirth_numeric_05,
    total_siblings_05 = tsibl5,
    any_domestic_violence_05,
    any_domestic_violence_numeric_05,
    child_harm_05,
    child_harm_recoded_05,
    child_harm_recoded_numeric_05,
    total_social_support_05 = tssupm5,
    total_activities_with_mum_05 = actvm5,
    maternal_warmth_05,
    maternal_warmth_continuous_05,
    prosocial_behaviours_combined_05 = totproe5,
    # Parent characteristics
    maternal_depression_lifetime_05,
    maternal_depression_lifetime_numeric_05,
    maternal_personality_openness_05 = bfiom5,
    maternal_personality_conscientiousness_05 = bficm5,
    maternal_personality_extroversion_05 = bfiem5,
    maternal_personality_agreeableness_05 = bfiam5,
    maternal_personality_neuroticism_05 = bfinm5,
    antisocial_behaviour_mum_05 = asbmm5,
    antisocial_behaviour_dad_05 = asbfm5,
    antisocial_behaviour_parent_05,
    alcoholism_mum_05 = alcmm5,
    alcoholism_dad_05 = alcfm5,
    alcoholism_parent_05,
    # Child neurodevelopment
    IQ_05 = iqe5,
    executive_function_05 = exfunce5,
    theory_of_mind_05 = tomtote5,
    # Child emotional and behavioural development
    externalising_combined_05 = totexte5,
    internalising_combined_excl_sis_05 = intisoe5,
    ADHD_combined_05 = totadde5,
    temp_negative_affect_05 = irre5,
    temp_impulsivity_05 = impe5,
    temp_approach_05 = appe5,
    temp_sluggishness_05 = slue5,
    temp_wariness_05 = ware5,
    temp_undercontrolled_05 = unce5,
    temp_inhibited_05 = inhe5,
    temp_shy_05 = shye5,
    ## Age 18
    # Mental health and service use
    depression_diagnosis_12mo_18,
    depression_diagnosis_12mo_numeric_18,
    depression_current_scale_18 = mdesxe18, 
    anxiety_diagnosis_18,
    anxiety_diagnosis_numeric_18,
    anxiety_current_scale_18 = gadsxe18, 
    ADHD_diagnosis_18,
    ADHD_diagnosis_numeric_18,
    inattentive_hyperactive_symptoms_total_18 = SR_symtot18e,
    conduct_disorder_moderate_18, 
    conduct_disorder_moderate_numeric_18,
    conduct_disorder_symptoms_18 = cdsxe18,
    alcohol_dependence_18,
    alcohol_dependence_numeric_18,
    alcohol_abuse_18,
    alcohol_abuse_numeric_18,
    alcohol_symptom_scale_18 = alcsxe18,
    cannabis_dependence_18,
    cannabis_dependence_numeric_18,
    cannabis_symptom_scale_18 = marjsxe18,
    PTSD_diagnosis_lifetime_18,
    PTSD_diagnosis_lifetime_numeric_18,
    PTSD_diagnosis_current_18,
    PTSD_diagnosis_current_numeric_18,
  #  psychosis_symptom_count_18 = psysympe18,
  #  cat_psychosis_symptom_count_18,
  #  psychosis_experiences_scale_18 = psyexpe18,
    cat_psychosis_experiences_scale_18, # only including one psychosis variable for now 
    cat_psychosis_experiences_scale_numeric_18,
    suicide_attempt_selfharm_18,
    suicide_attempt_selfharm_numeric_18,
    service_use_18,
    service_use_numeric_18,
    # Physical health and health risks
    BMI_18 = bmie18,
    CRP_log_18 = lnCRP_E18_4SD,
    physical_activity_18 = phyacte18,
    smoking_current_number_18 = smkcnume18,
    # Coping and functioning
    loneliness_18 = lonelye18,
    life_satisfaction_18 = lifsate18,
    technology_use_18 = teche18,
    coping_with_stress_18 = copstrse18,
    PSQI_global_score_18 = psqie18,
    #Employment prospects
    not_in_employment_education_18,
    not_in_employment_education_numeric_18,
    highest_education_18,
    highest_education_recode_18,
    job_preparedness_skills_18 = jprepse18,
    job_preparedness_attributes_18 = jprepae18,
    optimism_18 = optime18,
    job_search_activities_count_18 = jbschacte18,
    ## Social isolation
    isolation_combined_05 = sisoe5, # only ELDER variables included here - need to add a chunk to select younger variables if decide to do twin analyses. 
    isolation_combined_07 = sisoe7,
    isolation_combined_10 = sisoe10,
    isolation_combined_12 = sisoe12,
    isolation_mother_05 = sisoem5,
    isolation_mother_07 = sisoem7,
    isolation_mother_10 = sisoem10,
    isolation_mother_12 = sisoem12,
    isolation_teacher_05 = sisoet5,
    isolation_teacher_07 = sisoet7,
    isolation_teacher_10 = sisoet10,
    isolation_teacher_12 = sisoet12,
    log_isolation_combined_05 = log_sisoe5, # log values needed for trajectory analyses
    log_isolation_combined_07 = log_sisoe7,
    log_isolation_combined_10 = log_sisoe10,
    log_isolation_combined_12 = log_sisoe12,
    isolation.categorical.5,
    isolation.categorical.7,
    isolation.categorical.10,
    isolation.categorical.12,
    na.per.person.si
  )
# check
colnames(dat.renamed.full)
```

# Select variables to be exported into rds and csv files 

Have hashed out the trajectory file - each time you run this line, you need to then manually change the the NAs to be . and remove the header so that Mplus can read the file. I have been using only one export and not rerunning this line 

```{r selct variables to export files}
#select variables for rds file
dat_rds.3missing <- dat.renamed.3missing 
dat_rds.full <- dat.renamed.full

#select variables for csv file (for Mplus) - uses numeric values rather than factors
dat_csv.3missing <- dat.3missing %>%
  select(
    id, 
    familyid, 
    rorderp5, 
    torder, 
    risks,
    cohort,
    sampsex,
    zygosity,
    seswq35,
    sisoe5,
    sisoe7,
    sisoe10,
    sisoe12,
    log_5 = log_sisoe5,
    log_7 = log_sisoe7,
    log_10 = log_sisoe10,
    log_12 = log_sisoe12)

#select variables for csv file (for Mplus) - uses numeric values rather than factors
dat_csv.full <- dat %>%
  select(
    id, 
    familyid, 
    rorderp5, 
    torder, 
    risks,
    cohort,
    sampsex,
    zygosity,
    seswq35,
    sisoe5,
    sisoe7,
    sisoe10,
    sisoe12,
    log_5 = log_sisoe5,
    log_7 = log_sisoe7,
    log_10 = log_sisoe10,
    log_12 = log_sisoe12)

# variables for categorical LCGA
dat_csv.full.cat <- dat %>%
  select(
    id, 
    familyid, 
    rorderp5, 
    torder, 
    risks,
    cohort,
    sampsex,
    zygosity,
    seswq35,
    sisoe5,
    sisoe7,
    sisoe10,
    sisoe12,
    sisoe5_cat = isolation.categorical.5,
    sisoe7_cat = isolation.categorical.7,
    sisoe10_cat = isolation.categorical.10,
    sisoe12_cat = isolation.categorical.12)

# mother report social isolation only
dat_csv.full.mother <- dat %>%
  select(
    id, 
    familyid, 
    rorderp5, 
    torder, 
    risks,
    cohort,
    sampsex,
    zygosity,
    seswq35,
    sisoem5,
    sisoem7,
    sisoem10,
    sisoem12)

# teacher report social isolation only
dat_csv.full.teacher <- dat %>%
  select(
    id, 
    familyid, 
    rorderp5, 
    torder, 
    risks,
    cohort,
    sampsex,
    zygosity,
    seswq35,
    sisoet5,
    sisoet7,
    sisoet10,
    sisoet12)
```

# Export the data as rds and csv (for Mplus) files 

```{r export new data files}
# save R data file
saveRDS(object = dat_rds.3missing, file = paste0(data_path, "preprocessed_isolation_trajectories_Jan2021_3missing.rds"))
saveRDS(object = dat_rds.full, file = paste0(data_path, "preprocessed_isolation_trajectories_Jan2021_full_sample.rds"))

# save csv file in two places

# write_csv(x = dat_csv.3missing, path = paste0(data_path, "Mplus_back_up/FOR_MPLUS_preprocessed_isolation_trajectories_Jan2021_3missing.csv"))
# write_csv(x = dat_csv.3missing, path = "/Users/katiethompson/Desktop/mplusisolationtrajectoriesJan2021_3missing.csv")
# write_csv(x = dat_csv.full, path = paste0(data_path, "Mplus_back_up/FOR_MPLUS_preprocessed_isolation_trajectories_Jan2021_full_sample.csv"))
# write_csv(x = dat_csv.full, path = "/Users/katiethompson/Desktop/mplusisolationtrajectoriesJan2021_full_sample.csv") 

# categorical LCGA
# write_csv(x = dat_csv.full.cat, path = paste0(data_path, "Mplus_back_up/FOR_MPLUS_preprocessed_isolation_trajectories_Jan2021_full_sample_CAT.csv"))
# write_csv(x = dat_csv.full.cat, path = "/Users/katiethompson/Desktop/mplusisolationtrajectoriesJan2021_full_sample_CAT.csv") 

# GMM for mother separately 
# write_csv(x = dat_csv.full.mother, path = "/Users/katiethompson/Desktop/mplusisolationtrajectoriesJan2021_full_sample_MOTHER.csv") 

# GMM for teacher separately 
# write_csv(x = dat_csv.full.teacher, path = "/Users/katiethompson/Desktop/mplusisolationtrajectoriesJan2021_full_sample_TEACHER.csv") 

#remember to delete the headings. There will be no NAs as they have been removed for the isolation variables but in future Mplus files need NA to be a fullstop (.)
```

