---
title: "GMM trajectory statistics and figures for quadratic models"
output:  
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: flatly
    code_folding: hide
    includes:
      after_body: footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = NA,
                      prompt = FALSE,
                      cache = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      results = 'asis')

options(bitmapType = 'quartz') # to render fonts better
```

```{r Clear global environment, include=FALSE}
remove(list = ls())
```

```{r colour palletes for graphs, include=FALSE}
palette1 <- c("#78D9C5")
palette2 <- c("#78D9C5","#F5BE5E")
palette3 <- c("#78D9C5","#F5BE5E","#EEB6E9")
palette4 <- c("#78D9C5","#F5BE5E","#EEB6E9","#DBDB73")
palette5 <- c("#78D9C5","#F5BE5E","#EEB6E9","#DBDB73","#FFED98")
palette6 <- c("#78D9C5","#F5BE5E","#EEB6E9","#DBDB73","#FFED98","#BFD2EB")
```
 
```{r source the data file path}
#source raw data directory: data_raw and data included
source("../../../isolation_trajectories_data_path.R")
```

```{r Packages, include=FALSE}
library(MplusAutomation)
library(summarytools)
library(stringr)
library(stringi)
library(filesstrings)
library(texreg)
library(relimp)
library(rhdf5)
library(knitr)
library(ggplot2)
library(tidyverse)
library(plyr) #conflicts with tidyverse for e.g. rename and row_number
```

# Quadratic GMM 

## Quadratic term significance for each model {.tabset .tabset-fade}

```{r read in data for growth curves, results='hide', message=FALSE}
si.quad.all <- readModels(paste0(mplus_GMM_clustered_full_output_QUAD_data_path), recursive = TRUE)
```

```{r extract summary stats for growth curves}
#extract the summary variables from the mplus output files. Assuming there are multiple files in the above directory, model summaries could be retained as a data.frame as follows:
si_quad_summaries <- do.call("rbind.fill",
                        sapply(si.quad.all,
                               "[", 
                               "summaries"))
```

Here we only reprt trsults for the two and three class models as the four to six class models did not converge. 

* 4th class originally had problems with negative S variances - set S@0 for this model, this then resulted in non-convergence. 
* 5th class originally had problems with negative S variances - set S@0 for this model, this then resulted in non-convergence. 
* 6th class originally had problems with negative S variances - set S@0 for this model, this then resulted in non-convergence. 

**Significance values for 2 class model:**
Q has significant mean for class 1 (0.044), although only just. Q has non significant mean for class 2 (0.085). 

**Significance values for 3 class model:**
Q has significant mean for class 1 (0.001). Q has significant mean for class 2 (0.002). Q has significant mean for class 3 (0.008).


```{r extract significance values of Q}
# two class
quad.mean.two.class <- si.quad.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.quadratic..isolation_2traj_full_sample_clustered_QUAD.out$parameters$unstandardized %>%
  filter(paramHeader == "Means" & LatentClass < 7) # filter out the means for I, S and Q for each class 

quad.variance.two.class <- si.quad.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.quadratic..isolation_2traj_full_sample_clustered_QUAD.out$parameters$unstandardized %>%
  filter(paramHeader == "Variances" & LatentClass < 7) # filter out the means for I, S and Q for each class 

# three class
quad.mean.three.class <- si.quad.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.quadratic..isolation_3traj_full_sample_clustered_QUAD.out$parameters$unstandardized %>%
  filter(paramHeader == "Means" & LatentClass < 7) # filter out the means for I, S and Q for each class 

quad.variance.three.class <- si.quad.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.quadratic..isolation_3traj_full_sample_clustered_QUAD.out$parameters$unstandardized %>%
  filter(paramHeader == "Variances" & LatentClass < 7) # filter out the means for I, S and Q for each class 
```

### Two class model

```{r means two class}
two.class.mean.var <- rbind(quad.mean.two.class, quad.variance.two.class) %>%
  select(`Mean/variance` = paramHeader, 
         `Intercept/slope/quadratic` = param, 
         `estimate` = est, 
         `standard error` = se, 
         `p value` = pval)
kable(two.class.mean.var)
```

### Three class model

```{r means three class}
three.class.mean.var <- rbind(quad.mean.three.class, quad.variance.three.class) %>%
  select(`Mean/variance` = paramHeader, 
         `Intercept/slope/quadratic` = param, 
         `estimate` = est, 
         `standard error` = se, 
         `p value` = pval)
kable(three.class.mean.var)
```

## Model fit statistics for two and three class models

```{r remove 4,5,6 class models}
# remove 4,5,6 class models
si.two_three.summaries <- si_quad_summaries[-c(3:5),] 
```

```{r create data frame with sumamary variables}
#social isolation summaries 
si.class.summaries <- data.frame(matrix(nrow = 2, 
                                         ncol = 8)) 

#create column names for the variables you will be adding to the empty matrix of si.class.summaries
colnames(si.class.summaries) <- c("Model", 
                                  "Parameters", 
                                  "AIC", 
                                  "BIC", 
                                  "aBIC", 
                                  "Entropy", 
                                  "VLMR LRT p-value", 
                                  "Class Proportions") #or whichever indices you want to compare; do si_summaries$ to see what's in there

#check colnames
#si.class.summaries

#create "Model" variable
si.class.summaries <- si.class.summaries %>%
  mutate(
    Model = 
    as.factor(c("Two Class", #change nrow in data frame above when adding more models
                "Three Class")))  

#add summary information into data frame 
si.class.summaries <- si.class.summaries %>%
  mutate(
    Parameters = 
      si.two_three.summaries$Parameters) %>% #parameters col from original summaries data set
  mutate(
    AIC = 
      si.two_three.summaries$AIC) %>% #AIC col from original summaries data set
  mutate(
    BIC = 
      si.two_three.summaries$BIC) %>% #BIC col from original summaries data set
  mutate(
    aBIC = 
      si.two_three.summaries$aBIC) %>% #aBIC col from original summaries data set
  mutate(
    Entropy = 
      si.two_three.summaries$Entropy) %>% #Entropy col from original summaries data set
  mutate(
    `VLMR LRT p-value` = 
      si.two_three.summaries$T11_VLMR_PValue) #T11_VLMR_PValue col from original summaries data set

#check
#si.class.summaries
```

```{r create class proportions % for each model}
#for each class we want to add in the model estimated class counts - then convert this into a list with the percentage of people in each class

#two classes
si.class.summaries$`Class Proportions`[si.class.summaries$Model == "Two Class"] <- list(c(sprintf("%.0f", si.quad.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.quadratic..isolation_2traj_full_sample_clustered_QUAD.out$class_counts$modelEstimated$proportion*100)))

#three classes
si.class.summaries$`Class Proportions`[si.class.summaries$Model == "Three Class"] <- list(c(sprintf('%.0f', si.quad.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.quadratic..isolation_3traj_full_sample_clustered_QUAD.out$class_counts$modelEstimated$proportion*100)))
```

```{r create class Ns for each model}
#two classes
si.class.summaries$`Class N`[si.class.summaries$Model == "Two Class"] <- list(c(sprintf("%.0f", si.quad.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.quadratic..isolation_2traj_full_sample_clustered_QUAD.out$class_counts$posteriorProb$count)))

#three classes
si.class.summaries$`Class N`[si.class.summaries$Model == "Three Class"] <- list(c(sprintf("%.0f", si.quad.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.quadratic..isolation_3traj_full_sample_clustered_QUAD.out$class_counts$posteriorProb$count)))
```

```{r load summary table for GMM models}
#Look at summary table 
kable(si.class.summaries)
```

## Plots {.tabset .tabset-fade}

### Two Class model
```{r two class GMM model preparation}
#create two class data frame with means and variances
two_class_si <- as.data.frame(
  si.quad.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.quadratic..isolation_2traj_full_sample_clustered_QUAD.out$gh5$means_and_variances_data$y_estimated_means)

#name the variables
names(two_class_si) <- c("Class 1","Class 2")

#create time point column 
two_class_si$timepoint <- c(5, 7, 10, 12)

#convert data to long format 
two_class_si <- two_class_si %>%
  pivot_longer(-timepoint, 
               names_to = "Class", 
               values_to = "social.isolation_score")

#check
#two_class_si
```

```{r plot 2 class GMM  model}
two_traj_plot_clustered_full_sample <- ggplot(two_class_si,
        aes(x = timepoint, 
            y = social.isolation_score, 
            colour = Class)) +
          geom_line(size = 1.5) +
  geom_point(aes(#shape = class, 
                 size = 1)) +
  scale_fill_manual(values = palette2) +
  scale_color_manual(values = palette2) +
  labs(x = "Age (in years)", 
       title = "Two class model",
       color = "Class",
       y ="Social isolation score") +   
  scale_y_continuous(expand = c(0.1,0.1),
                     limits = c(0,12),
                     breaks = seq(0, 12, 2)) +
  scale_x_continuous(expand = c(0.1,0.1),
                     limits = c(5,12),
                     breaks = c(5, 7, 10, 12)) +
  theme(panel.grid.major.y = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray"),
        axis.text = element_text(colour="black", 
                                 size = 12),
        axis.title = element_text(colour="black", 
                                  size = 12),
        panel.background = element_blank(),
        plot.title = element_text(size = 16, hjust = 0.5)) + 
  guides(shape = FALSE, size = FALSE)

two_traj_plot_clustered_full_sample
```

### Three Class model

```{r three class GMM model preparation}
#create three class data frame with means and variances
three_class_si <- as.data.frame(
  si.quad.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.quadratic..isolation_3traj_full_sample_clustered_QUAD.out$gh5$means_and_variances_data$y_estimated_means)

#name the variables
names(three_class_si) <- c("Class 1","Class 2","Class 3")

#create time point column 
three_class_si$timepoint <- c(5, 7, 10, 12)

#convert data to long format 
three_class_si <- three_class_si %>%
  pivot_longer(-timepoint, 
               names_to = "Class", 
               values_to = "social.isolation_score")

#check
#three_class_si
```

```{r plot 3 class GMM model}
three_traj_plot_clustered_full_sample <- ggplot(three_class_si,
        aes(x = timepoint, 
            y = social.isolation_score, 
            colour = Class)) +
          geom_line(size = 1.5) +
  geom_point(aes(size = 1)) +
  scale_fill_manual(values = palette3) +
  scale_color_manual(values = palette3) +
  labs(x = "Age (in years)", 
       title = "Three class model",
       color = "Class",
       y ="Social isolation score") +   
  scale_y_continuous(expand = c(0.1,0.1),
                     limits = c(0,12),
                     breaks = seq(0, 12, 2)) +
  scale_x_continuous(expand = c(0.1,0.1),
                     limits = c(5,12),
                     breaks = c(5, 7, 10, 12)) +
  theme(panel.grid.major.y = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray"),
        axis.text = element_text(colour="black", 
                                 size = 12),
        axis.title = element_text(colour="black", 
                                  size = 12),
        panel.background = element_blank(),
        plot.title = element_text(size = 16, hjust=0.5)) + 
  guides(shape = FALSE, size = FALSE)

three_traj_plot_clustered_full_sample
```

```{r save as JPEG, include=FALSE}
# Two trajectories
ggsave(
  "two_traj_plot_clustered_full_sample_QUAD.jpeg",
  plot = two_traj_plot_clustered_full_sample,
  device = "jpeg",
  path = paste0(graph_save_data_path, "trajectories/"),
  width = 7,
  height = 5
)
# Three trajectories
ggsave(
  "three_traj_plot_clustered_full_sample_QUAD.jpeg",
  plot = three_traj_plot_clustered_full_sample,
  device = "jpeg",
  path = paste0(graph_save_data_path, "trajectories/"),
  width = 7,
  height = 5
)
```

```{r combine graphs into panel figure, include=FALSE}
library(ggpubr) # Needed for combining plots
library(gridExtra)
combined_trajectories_plot_clustered_full_sample_quadratic <- ggarrange(two_traj_plot_clustered_full_sample, 
                                                                        three_traj_plot_clustered_full_sample, 
                                                                        ncol = 2, nrow = 1,
                                                                        widths = c(2.7,2.7), 
                                                                        heights = c(3,3),
                                                                        labels = c("A", "B"),
                                                                        legend = "none"
                                                                        )

ggsave(
  "combined_trajectories_plot_clustered_full_sample_quadratic.jpeg",
  plot = combined_trajectories_plot_clustered_full_sample_quadratic,
  device = "jpeg",
  path = paste0(graph_save_data_path, "trajectories/"),
  width = 9,
  height = 4
)
```

# Allocation differences between linear and quadratic 3 class model

```{r read in prior descriptives data and check variable names}
dat.descriptives <- readRDS(file = paste0(data_path, "preprocessed_isolation_trajectories_Jan2021_full_sample.rds"))
#colnames(dat.descriptives)
```

```{r read in txt file with probabilities LINEAR}
dat.tra.prob.full.LINEAR <- read.table(paste0(mplus_GMM_clustered_full_output_data_path, "GMM_SI_3Cl_full_sample_clustered.txt"), 
                            header = FALSE, 
                            col.names = c("SISOE5", # original social isolation score at age 5
                                         "SISOE7",  # original social isolation score at age 7
                                         "SISOE10", # original social isolation score at age 10
                                         "SISOE12", # original social isolation score at age 12
                                         "I",       # intercept
                                         "S",       # slope
                                         "C_I",     # factor scores based on most likely class membership
                                         "C_S",     # factor scores based on most likely class membership
                                         "CPROB1",  # class probability 1
                                         "CPROB2",  # class probability 2
                                         "CPROB3",  # class probability 3
                                         "C",       # class 
                                         "ID",      # ID
                                         "FAMILYID" # Family ID - twin clustering  
                                       ))   
#check 
#dat.tra.prob.full.LINEAR
```

```{r read in txt file with probabilities QUAD}
dat.tra.prob.full.QUAD <- read.table(paste0(mplus_GMM_clustered_full_output_QUAD_data_path, "GMM_SI_3Cl_full_sample_clustered_QUAD.txt"), 
                            header = FALSE, 
                            col.names = c("SISOE5", # original social isolation score at age 5
                                         "SISOE7",  # original social isolation score at age 7
                                         "SISOE10", # original social isolation score at age 10
                                         "SISOE12", # original social isolation score at age 12
                                         "I",       # intercept
                                         "S",       # slope
                                         "Q",       # quadratic
                                         "C_I",     # factor scores based on most likely class membership
                                         "C_S",     # factor scores based on most likely class membership
                                         "C_Q",     # factor scores based on most likely class membership
                                         "CPROB1",  # class probability 1
                                         "CPROB2",  # class probability 2
                                         "CPROB3",  # class probability 3
                                         "C",       # class 
                                         "ID",      # ID
                                         "FAMILYID" # Family ID - twin clustering  
                                       ))   
#check 
#dat.tra.prob.full.QUAD
```

```{r select relevant variables}
dat.tra.prob.LINEAR <- dat.tra.prob.full.LINEAR %>%
  select(id = ID, 
         prob1.linear = CPROB1,
         prob2.linear = CPROB2,
         prob3.linear = CPROB3,
         class.linear = C
         )
#check 
#dat.tra.prob.LINEAR

dat.tra.prob.QUAD <- dat.tra.prob.full.QUAD %>%
  select(id = ID, 
         prob1.quad = CPROB1,
         prob2.quad = CPROB2,
         prob3.quad = CPROB3,
         class.quad = C
         )
#check 
#dat.tra.prob.QUAD

#create list of data frames to merge
dataframe_list <- list(
  dat.descriptives,
  dat.tra.prob.LINEAR,
  dat.tra.prob.QUAD
)

#merge data frames
dat <- plyr::join_all(
  dataframe_list,
  by = "id" # Alternatively you can join by several columns
  )

#check
#colnames(dat)
```

```{r rename class variable}
# linear
dat <- dat %>%
  mutate(class_renamed.linear =
    recode_factor(class.linear,
      "1" = "Low stable",
      "2" = "Increasing",
      "3" = "Decreasing"
    )
  )
as.data.frame(table(dat$class_renamed.linear)) %>% 
  select(`Linear classes` = Var1,
         Freq)

# quadratic
dat <- dat %>%
  mutate(class_renamed.quad =
    recode_factor(class.quad,
      "1" = "Low stable",
      "2" = "Increasing",
      "3" = "Decreasing"
    )
  )
as.data.frame(table(dat$class_renamed.quad)) %>% 
  select(`Quadratic classes` = Var1,
         Freq)
```

## Matching classification

96.51% of people were classified the same way regardless if the model was linear or quadratic. 

Overall average movement from Linear to Quadratic: 

* 1  more person in the increasing class
* 18 less people in the decreasing class
* 17 more people in the low stable class

```{r matching linear and quadratic}
dat <- dat %>%
  mutate(
    matching.class = 
      case_when(
        class.linear == class.quad ~ "Matching",
        class.linear != class.quad ~ "Not matching"
      )
  )

freq(dat$matching.class,
        cumul = FALSE, 
       display.type = FALSE, 
       headings = FALSE,
       style = "rmarkdown",
       report.nas = FALSE)
```

* 29 in the decreasing class changed classes in the quadratic model. 
* 19 in the increasing class changed classes in the quadratic model.
* 30 in the low stable class changed classes in the quadratic model.
== 78 who are not matching. 

```{r descriptives for nonmatching participants}
dat <- dat %>%
  mutate(matching.isolated =
           case_when(
             matching.class == "Matching" & class_renamed.linear == "Increasing" ~ "Matching and increasing",
             matching.class == "Matching" & class_renamed.linear == "Decreasing" ~ "Matching and decreasing",
             matching.class == "Matching" & class_renamed.linear == "Low stable" ~ "Matching and low stable"
           ))

freq(dat$matching.isolated,
        cumul = FALSE, 
       display.type = FALSE, 
       headings = FALSE,
       style = "rmarkdown",
       report.nas = FALSE)
```

```{r filter non matching}
dat.non.match <- dat %>%
  filter(matching.class == "Not matching")
```

```{r direction of non match}
# from linear to quadratic model
dat.non.match <- dat.non.match %>%
  mutate(
    direction.match =
      case_when(
        class_renamed.linear == "Increasing" & class_renamed.quad == "Low stable" ~ "Increasing to low stable",
        class_renamed.linear == "Decreasing" & class_renamed.quad == "Low stable" ~ "Decreasing to low stable",
        class_renamed.linear == "Increasing" & class_renamed.quad == "Decreasing" ~ "Increasing to decreasing",
        class_renamed.linear == "Decreasing" & class_renamed.quad == "Increasing" ~ "Decreasing to increasing",
        class_renamed.linear == "Low stable" & class_renamed.quad == "Increasing" ~ "Low stable to increasing",
        class_renamed.linear == "Low stable" & class_renamed.quad == "Decreasing" ~ "Low stable to decreasing",
      )
  )

freq(dat.non.match$direction.match,
     cumul = FALSE, 
       display.type = FALSE, 
       headings = FALSE,
       style = "rmarkdown",
       report.nas = FALSE)
```

* 1 person moved from the decreasing to the increasing class.
* 28 people moved from the decreasing to the low stable class.
* 19 people moved from the increasing to the low stable class. 
* 11 people moved from the low stable to the decreasing class.
* 19 people moved from the low stable to the increasing class.
== 78 non matching. 

