---
title: "LCGA statistics and figures"
output:  
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: flatly
    code_folding: hide
    includes:
      after_body: footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = NA,
                      prompt = FALSE,
                      cache = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      results = 'asis')

options(bitmapType = 'quartz') # to render fonts better
```

```{r Clear global environment, include=FALSE}
remove(list = ls())
```

```{r colour palletes for graphs, include=FALSE}
palette1 <- c("#78D9C5")
palette2 <- c("#78D9C5","#F5BE5E")
palette3 <- c("#78D9C5","#F5BE5E","#EEB6E9")
palette4 <- c("#78D9C5","#F5BE5E","#EEB6E9","#DBDB73")
palette5 <- c("#78D9C5","#F5BE5E","#EEB6E9","#DBDB73","#FFED98")
palette6 <- c("#78D9C5","#F5BE5E","#EEB6E9","#DBDB73","#FFED98","#BFD2EB")
```
 
```{r source the data file path, include=FALSE}
#source raw data directory: data_raw and data included
source("../../../isolation_trajectories_data_path.R")
```

```{r Mplus packages, include=FALSE}
library(MplusAutomation)
library(stringr)
library(stringi)
library(filesstrings)
library(texreg)
library(relimp)
library(rhdf5)
library(knitr)
library(ggplot2)
library(tidyverse)
library(plyr) #conflicts with tidyverse for e.g. rename and row_number
```

***

This page displays results from the Latent class growth analysis, which we used as a preliminary analysis step before conducting GMM. LCGA constrains the variance within each class, meaning that the classes that are estimated are trying to account for more variance in the data - thus more classes are often idenfified. 

***

# Linear Latent Class Growth Analysis 

## Model fit statistics

We selected the **four class LCGA model** as the best fitting model. 

```{r read in social isolation model data, results='hide', message=FALSE}
#For class models, read in all output files within your folder that you have all your class models
si.classes.all <- readModels(paste0(mplus_LCGA_clustered_full_output_data_path), recursive = TRUE) #Recursive means it reads data in sub-folders too
```

```{r read in trajectoriesectory summary vairables}
#extract the summary variables from the mplus output files. Assuming there are multiple files in the above directory, model summaries could be retained as a data.frame as follows:
si_summaries <- do.call("rbind.fill",
                        sapply(si.classes.all,
                               "[", 
                               "summaries"))
```

```{r create data frame with sumamary variables}
#social isolation summaries 
si.class.summaries <- data.frame(matrix(nrow = 5, 
                                         ncol = 8)) 

#create column names for the variables you will be adding to the empty matrix of si.class.summaries
colnames(si.class.summaries) <- c("Model", 
                                  "Parameters", 
                                  "AIC", 
                                  "BIC", 
                                  "aBIC", 
                                  "Entropy", 
                                  "VLMR LRT p-value", 
                                  "Class Proportions") #or whichever indices you want to compare; do si_summaries$ to see what's in there

#check colnames
#si.class.summaries

#create "Model" variable
si.class.summaries <- si.class.summaries %>%
  mutate(
    Model = 
    as.factor(c("Two Class", #change nrow in data frame above when adding more models
                "Three Class", 
                "Four Class", 
                "Five Class",
                "Six Class")))  

#add summary information into data frame 
si.class.summaries <- si.class.summaries %>%
  mutate(
    Parameters = 
      si_summaries$Parameters) %>% #parameters col from original summaries data set
  mutate(
    AIC = 
      si_summaries$AIC) %>% #AIC col from original summaries data set
  mutate(
    BIC = 
      si_summaries$BIC) %>% #BIC col from original summaries data set
  mutate(
    aBIC = 
      si_summaries$aBIC) %>% #aBIC col from original summaries data set
  mutate(
    Entropy = 
      si_summaries$Entropy) %>% #Entropy col from original summaries data set
  mutate(
    `VLMR LRT p-value` = 
      si_summaries$T11_VLMR_PValue) #T11_VLMR_PValue col from original summaries data set

#check
# si.class.summaries
```

```{r create class proportions % for each model}
#for each class we want to add in the model estimated class counts - then convert this into a list with the percentage of people in each class

#two classes
si.class.summaries$`Class Proportions`[si.class.summaries$Model == "Two Class"] <- list(c(sprintf("%.0f", si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.LCGA.clustered.full_sample.output..LCGA_isolation_2traj_full_sample_clustered.out$class_counts$modelEstimated$proportion*100)))

#three classes
si.class.summaries$`Class Proportions`[si.class.summaries$Model == "Three Class"] <- list(c(sprintf('%.0f', si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.LCGA.clustered.full_sample.output..LCGA_isolation_3traj_full_sample_clustered.out$class_counts$modelEstimated$proportion*100)))

#four classes
si.class.summaries$`Class Proportions`[si.class.summaries$Model == "Four Class"] <- list(c(sprintf('%.0f', si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.LCGA.clustered.full_sample.output..LCGA_isolation_4traj_full_sample_clustered.out$class_counts$modelEstimated$proportion*100)))

#five classes
si.class.summaries$`Class Proportions`[si.class.summaries$Model == "Five Class"] <- list(c(sprintf('%.0f', si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.LCGA.clustered.full_sample.output..LCGA_isolation_5traj_full_sample_clustered.out$class_counts$modelEstimated$proportion*100)))

#six classes
si.class.summaries$`Class Proportions`[si.class.summaries$Model == "Six Class"] <- list(c(sprintf('%.0f', si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.LCGA.clustered.full_sample.output..LCGA_isolation_6traj_full_sample_clustered.out$class_counts$modelEstimated$proportion*100)))
```

```{r create class Ns for each model}
#two classes
si.class.summaries$`Class N`[si.class.summaries$Model == "Two Class"] <- list(c(sprintf("%.0f", si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.LCGA.clustered.full_sample.output..LCGA_isolation_2traj_full_sample_clustered.out$class_counts$posteriorProb$count)))

#three classes
si.class.summaries$`Class N`[si.class.summaries$Model == "Three Class"] <- list(c(sprintf("%.0f", si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.LCGA.clustered.full_sample.output..LCGA_isolation_3traj_full_sample_clustered.out$class_counts$posteriorProb$count)))

#four classes
si.class.summaries$`Class N`[si.class.summaries$Model == "Four Class"] <- list(c(sprintf("%.0f", si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.LCGA.clustered.full_sample.output..LCGA_isolation_4traj_full_sample_clustered.out$class_counts$posteriorProb$count)))

#five classes
si.class.summaries$`Class N`[si.class.summaries$Model == "Five Class"] <- list(c(sprintf("%.0f", si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.LCGA.clustered.full_sample.output..LCGA_isolation_5traj_full_sample_clustered.out$class_counts$posteriorProb$count)))

#six classes
si.class.summaries$`Class N`[si.class.summaries$Model == "Six Class"] <- list(c(sprintf("%.0f", si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.LCGA.clustered.full_sample.output..LCGA_isolation_6traj_full_sample_clustered.out$class_counts$posteriorProb$count)))
```

```{r load summary table for GMM models}
#Look at summary table 
kable(si.class.summaries)
```
 
*Reminder that LCGA constrains the variance in each class to zero.*

## Classification probabilities 

```{r extract classification probs for 3 class model}
# extract the classification probabilities for the most likely class memberships (column) by latent class (row). This shows the uncertainty rate of the probabilities. If it was perfect - the diagonal would be 1. E.g. for Class 1 - 98.3% of individuals fit that category. 
class.probs <- data.frame(si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.LCGA.clustered.full_sample.output..LCGA_isolation_4traj_full_sample_clustered.out$class_counts$classificationProbs.mostLikely)

# add column names
colnames(class.probs) <- c("Low stable","Decreasing","Low increasing", "High increasing")

class.probs$Class <- c("Low stable","Decreasing","Low increasing", "High increasing")

#as.tibble(class.probs)

class.probs <- class.probs %>%
  tidyr::pivot_longer(-Class, 
               names_to = "class", 
               values_to = "Classification probabilities")

class.probs <- class.probs[-c(2:5,7:10,12:15),-c(2)]
```

```{r classification probabilities table}
kable(class.probs)
```

## Plots {.tabset .tabset-fade}

### Two Class model
```{r two class GMM model preparation}
#create two class data frame with means and variances
two_class_si <- as.data.frame(
  si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.LCGA.clustered.full_sample.output..LCGA_isolation_2traj_full_sample_clustered.out$gh5$means_and_variances_data$y_estimated_means)

#name the variables
names(two_class_si) <- c("Class 1","Class 2")

#create timepoint column 
two_class_si$timepoint <- c(5, 7, 10, 12)

#convert data to long format 
two_class_si <- two_class_si %>%
  pivot_longer(-timepoint, 
               names_to = "Class", 
               values_to = "social.isolation_score")

#check
#two_class_si
```

```{r plot 2 class GMM  model}
two_traj_plot_clustered_full_sample <- ggplot(two_class_si,
        aes(x = timepoint, 
            y = social.isolation_score, 
            colour = Class)) +
          geom_line(size = 1.5) +
  geom_point(aes(#shape = class, 
                 size = 1)) +
  scale_fill_manual(values = palette2) +
  scale_color_manual(values = palette2) +
  labs(x = "Age (in years)", 
       title = "Two class LCGA model of social isolation in E-risk",
       color = "Class",
       y ="Social isolation score") +   
  scale_y_continuous(expand = c(0.1,0.1),
                     limits = c(0,12),
                     breaks = seq(0, 12, 2)) +
  scale_x_continuous(expand = c(0.1,0.1),
                     limits = c(5,12),
                     breaks = c(5, 7, 10, 12)) +
  theme(panel.grid.major.y = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray"),
        axis.text = element_text(colour="black", 
                                 size = 12),
        axis.title = element_text(colour="black", 
                                  size = 12),
        panel.background = element_blank(),
        plot.title = element_text(size = 16, hjust = 0.5)) + 
  guides(shape = FALSE, size = FALSE)

two_traj_plot_clustered_full_sample
```

### Three Class model

```{r three class GMM model preparation}
#create three class data frame with means and variances
three_class_si <- as.data.frame(
  si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.LCGA.clustered.full_sample.output..LCGA_isolation_3traj_full_sample_clustered.out$gh5$means_and_variances_data$y_estimated_means)

#name the variables
names(three_class_si) <- c("Class 1","Class 2","Class 3")

#create time point column 
three_class_si$timepoint <- c(5, 7, 10, 12)

#convert data to long format 
three_class_si <- three_class_si %>%
  pivot_longer(-timepoint, 
               names_to = "Class", 
               values_to = "social.isolation_score")

#check
#three_class_si
```

```{r plot 3 class GMM model}
three_traj_plot_clustered_full_sample <- ggplot(three_class_si,
        aes(x = timepoint, 
            y = social.isolation_score, 
            colour = Class)) +
          geom_line(size = 1.5) +
  geom_point(aes(size = 1)) +
  scale_fill_manual(values = palette3) +
  scale_color_manual(values = palette3) +
  labs(x = "Age (in years)", 
       title = "Three class LCGA model of social isolation in E-risk",
       color = "Class",
       y ="Social isolation score") +   
  scale_y_continuous(expand = c(0.1,0.1),
                     limits = c(0,12),
                     breaks = seq(0, 12, 2)) +
  scale_x_continuous(expand = c(0.1,0.1),
                     limits = c(5,12),
                     breaks = c(5, 7, 10, 12)) +
  theme(panel.grid.major.y = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray"),
        axis.text = element_text(colour="black", 
                                 size = 12),
        axis.title = element_text(colour="black", 
                                  size = 12),
        panel.background = element_blank(),
        plot.title = element_text(size = 16, hjust=0.5)) + 
  guides(shape = FALSE, size = FALSE)

three_traj_plot_clustered_full_sample
```

### Four Class model

```{r four class GMM model preparation}
#create four class data frame with means and variances
four_class_si <- as.data.frame(
  si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.LCGA.clustered.full_sample.output..LCGA_isolation_4traj_full_sample_clustered.out$gh5$means_and_variances_data$y_estimated_means)

#name the variables
names(four_class_si) <- c("Class 1","Class 2","Class 3","Class 4")

#create timepoint column 
four_class_si$timepoint <- c(5, 7, 10, 12)

#convert data to long format 
four_class_si <- four_class_si %>%
  pivot_longer(-timepoint, 
               names_to = "Class", 
               values_to = "social.isolation_score")

#check
#four_class_si
```

```{r plot 4 class GMM model}
four_traj_plot_clustered_full_sample <- ggplot(four_class_si,
        aes(x = timepoint, 
            y = social.isolation_score, 
            colour = Class)) +
          geom_line(size = 1.5) +
  geom_point(aes(size = 1)) +
  scale_fill_manual(values = palette4) +
  scale_color_manual(values = palette4) +
  labs(x = "Age (in years)", 
       title = "Four class LCGA model of social isolation in E-risk",
       color = "Class",
       y ="Social isolation score") +   
  scale_y_continuous(expand = c(0.1,0.1),
                     limits = c(0,12),
                     breaks = seq(0, 12, 2)) +
  scale_x_continuous(expand = c(0.1,0.1),
                     limits = c(5,12),
                     breaks = c(5, 7, 10, 12)) +
  theme(panel.grid.major.y = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray"),
        axis.text = element_text(colour="black", 
                                 size = 12),
        axis.title = element_text(colour="black", 
                                  size = 12),
        panel.background = element_blank(),
        plot.title = element_text(size = 16, hjust=0.5)) + 
  guides(shape = FALSE, size = FALSE)

four_traj_plot_clustered_full_sample
```

### Five Class model

```{r five class GMM  model preparation}
#create five class data frame with means and variances
five_class_si <- as.data.frame(
  si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.LCGA.clustered.full_sample.output..LCGA_isolation_5traj_full_sample_clustered.out$gh5$means_and_variances_data$y_estimated_means)

#name the variables
names(five_class_si) <- c("Class 1","Class 2","Class 3","Class 4","Class 5")

#create timepoint column 
five_class_si$timepoint <- c(5, 7, 10, 12)

#convert data to long format 
five_class_si <- five_class_si %>%
  pivot_longer(-timepoint, 
               names_to = "Class", 
               values_to = "social.isolation_score")

#check
#five_class_si
```

```{r plot 5 class GMM model}
five_traj_plot_clustered_full_sample <- ggplot(five_class_si,
        aes(x = timepoint, 
            y = social.isolation_score, 
            colour = Class)) +
          geom_line(size = 1.5) +
  geom_point(aes(size = 1)) +
  scale_fill_manual(values = palette5) +
  scale_color_manual(values = palette5) +
  labs(x = "Age (in years)", 
       title = "Five class LCGA model of social isolation in E-risk",
       color = "Class",
       y ="Social isolation score") +   
  scale_y_continuous(expand = c(0.1,0.1),
                     limits = c(0,12),
                     breaks = seq(0, 12, 2)) +
  scale_x_continuous(expand = c(0.1,0.1),
                     limits = c(5,12),
                     breaks = c(5, 7, 10, 12)) +
  theme(panel.grid.major.y = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray"),
        axis.text = element_text(colour="black", 
                                 size = 12),
        axis.title = element_text(colour="black", 
                                  size = 12),
        panel.background = element_blank(),
        plot.title = element_text(size = 16, hjust=0.5)) + 
  guides(shape = FALSE, size = FALSE)

five_traj_plot_clustered_full_sample
```

### Six Class model

```{r six class GMM  model preparation}
#create five class data frame with means and variances
six_class_si <- as.data.frame(
  si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.LCGA.clustered.full_sample.output..LCGA_isolation_6traj_full_sample_clustered.out$gh5$means_and_variances_data$y_estimated_means)

#name the variables
names(six_class_si) <- c("Class 1","Class 2","Class 3","Class 4","Class 5","Class 6")

#create timepoint column 
six_class_si$timepoint <- c(5, 7, 10, 12)

#convert data to long format 
six_class_si <- six_class_si %>%
  pivot_longer(-timepoint, 
               names_to = "Class", 
               values_to = "social.isolation_score")

#check
#six_class_si
```

```{r plot 6 class GMM model}
six_traj_plot_clustered_full_sample <- ggplot(six_class_si,
        aes(x = timepoint, 
            y = social.isolation_score, 
            colour = Class)) +
          geom_line(size = 1.5) +
  geom_point(aes(size = 1)) +
  scale_fill_manual(values = palette6) +
  scale_color_manual(values = palette6) +
  labs(x = "Age (in years)", 
       title = "Six class LCGA model of social isolation in E-risk",
       color = "Class",
       y = "Social isolation score") +   
  scale_y_continuous(expand = c(0.1,0.1),
                     limits = c(0,12),
                     breaks = seq(0, 12, 2)) +
  scale_x_continuous(expand = c(0.1,0.1),
                     limits = c(5,12),
                     breaks = c(5, 7, 10, 12)) +
  theme(panel.grid.major.y = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray"),
        axis.text = element_text(colour="black", 
                                 size = 12),
        axis.title = element_text(colour="black", 
                                  size = 12),
        panel.background = element_blank(),
        plot.title = element_text(size = 16, hjust=0.5)) + 
  guides(shape = FALSE, size = FALSE)

six_traj_plot_clustered_full_sample
```

```{r save as JPEG, include=FALSE}
# Two trajectories
ggsave(
  "LCGA_two_traj_plot_clustered_full_sample.jpeg",
  plot = two_traj_plot_clustered_full_sample,
  device = "jpeg",
  path = paste0(graph_save_data_path, "LCGA/"),
  width = 7,
  height = 5
)
# Three trajectories
ggsave(
  "LCGA_three_traj_plot_clustered_full_sample.jpeg",
  plot = three_traj_plot_clustered_full_sample,
  device = "jpeg",
  path = paste0(graph_save_data_path, "LCGA/"),
  width = 7,
  height = 5
)
# Four trajectories
ggsave(
  "LCGA_four_traj_plot_clustered_full_sample.jpeg",
  plot = four_traj_plot_clustered_full_sample,
  device = "jpeg",
  path = paste0(graph_save_data_path, "LCGA/"),
  width = 7,
  height = 5
)
# Five trajectories
ggsave(
  "LCGA_five_traj_plot_clustered_full_sample.jpeg",
  plot = five_traj_plot_clustered_full_sample,
  device = "jpeg",
  path = paste0(graph_save_data_path, "LCGA/"),
  width = 7,
  height = 5
)
# Six trajectories
ggsave(
  "LCGA_six_traj_plot_clustered_full_sample.jpeg",
  plot = six_traj_plot_clustered_full_sample,
  device = "jpeg",
  path = paste0(graph_save_data_path, "LCGA/"),
  width = 7,
  height = 5
)
```

# Combined LCGA trajectory plot

```{r rename plots, include=FALSE}
# rename all plot titles
two_traj_plot_clustered_full_sample <- two_traj_plot_clustered_full_sample +
  labs(title = "Two class model", color = "") +
  theme(plot.title = element_text(size = 12,face = "bold"),
        axis.title.x = element_text(size = 10, face = "bold"), 
        legend.position="none")

three_traj_plot_clustered_full_sample <- three_traj_plot_clustered_full_sample +
  labs(title = "Three class model", color = "") +
  theme(plot.title = element_text(size = 12,face = "bold"),
        axis.title.x = element_text(size = 10, face = "bold"),
        legend.position="none")

four_traj_plot_clustered_full_sample <- four_traj_plot_clustered_full_sample +
  labs(title = "Four class model", color = "") +
  theme(plot.title = element_text(size = 12,face = "bold"),
        axis.title.x = element_text(size = 10, face = "bold"),
        legend.position="none")

five_traj_plot_clustered_full_sample <- five_traj_plot_clustered_full_sample +
  labs(title = "Five class model", color = "") +
  theme(plot.title = element_text(size = 12,face = "bold"),
        axis.title.x = element_text(size = 10, face = "bold"),
        legend.position="none")

six_traj_plot_clustered_full_sample <- six_traj_plot_clustered_full_sample +
  labs(title = "Six class model", color = "") +
  theme(plot.title = element_text(size = 12,face = "bold"),
        axis.title.x = element_text(size = 10, face = "bold"),
        legend.position="none")
```

```{r combine trajectory plots}
#combine all plots
library(ggpubr) # Needed for combining plots
library(gridExtra)
combined_trajectories_plot_clustered_full_sample <- ggarrange( 
                                        two_traj_plot_clustered_full_sample, 
                                        three_traj_plot_clustered_full_sample, 
                                        four_traj_plot_clustered_full_sample, 
                                        five_traj_plot_clustered_full_sample,
                                        six_traj_plot_clustered_full_sample,
                                        ncol = 2, nrow = 3,
                                        widths = c(2.7,2.7,2.7,2.7), 
                                        heights = c(3,3,3,3),
                                        labels = c("A", "B", "C", "D", "E", "F"),
                                        legend = "bottom")

# ggsave(
#   "LCGA_combined_trajectories_plot_clustered_full_sample.jpeg",
#   plot = combined_trajectories_plot_clustered_full_sample,
#   device = "jpeg",
#   path = paste0(graph_save_data_path, "LCGA/"),
#   width = 9,
#   height = 10
# )
```

```{r combined plot, fig.height=10, fig.width=8.5}
combined_trajectories_plot_clustered_full_sample
```
