---
title: "Preliminary descriptives for social isolation in E-Risk"
author: "Katherine N Thompson"
date: "13th Jan 2021"
output:  
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: flatly
code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      comment = NA,
                      prompt = FALSE,
                      cache = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      results = 'asis')

#st_options(plain.ascii = FALSE, style = "rmarkdown")
options(bitmapType = 'quartz') # to render fonts better
```

```{r Clear global environment}
remove(list = ls())
```

```{r Current date}
date = Sys.Date()
```

```{r palletes}
palette2 <- c("#78D9C5","#F5BE5E")
palette3 <- c("#78D9C5","#F5BE5E","#EEB6E9")
palette4 <- c("#78D9C5","#F5BE5E","#EEB6E9","#DBDB73")
palette5 <- c("#78D9C5","#F5BE5E","#EEB6E9","#DBDB73","#FFED98")
palette6 <- c("#78D9C5","#F5BE5E","#EEB6E9","#DBDB73","#FFED98","#BFD2EB")
```

```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(skimr)
library(ggpubr)
library(ggplot2)
library(tidyr)
library(gplots)
library(Rmisc)
library(tidyverse)
```

```{r source the data file path}
#source raw data directory: data_raw and data included
source("../../../../../isolation_trajectories_data_path.R")
```

```{r read in data and check variable names, include=FALSE}
dat <- readRDS(file = paste0(data_path, "preprocessed_isolation_trajectories_Jan2021_full_sample.rds"))
colnames(dat)
```

```{r prepare item lists}
#All lists for isolation and grouping variables
#Only using elder twin variables as each twin is included in the data set as an individual. Without doing twin analyses, the elder and younger variables show the same information.
combined_social_isolation <- c("isolation_combined_05",
                              "isolation_combined_07",
                              "isolation_combined_10",
                              "isolation_combined_12")

teacher_social_isolation <- c("isolation_teacher_05",
                              "isolation_teacher_07",
                              "isolation_teacher_10",
                              "isolation_teacher_12")

mother_social_isolation <- c("isolation_mother_05",
                              "isolation_mother_07",
                              "isolation_mother_10",
                              "isolation_mother_12")
```

```{r Half-violin plot theme}
#source the half/flat function for the violin plot
source("https://raw.githubusercontent.com/datavizpyr/data/master/half_flat_violinplot.R")

#theme for ggplotting
theme_personal <-  theme(
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    axis.title.y = element_blank(),
    legend.background = element_blank(),
    legend.box.background = element_blank(),
    panel.background = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(
      colour = "gray",
      linetype = "dashed",
      size = 0.2
      ),
    axis.ticks = element_blank()
    )
```

# Missingness descriptives

## Frequency tables social isolation {.tabset .tabset-fade}

### NAs per person
```{r missingness Nas per person}
dat %>%
  rename(`Social isolation NAs per person` = na.per.person.si) %>%
  freq(`Social isolation NAs per person`, 
       cumul = FALSE, 
       display.type = FALSE, 
       headings = FALSE,
       style = "rmarkdown",
       report.nas = FALSE)
```

### Frequencies at age 5
```{r NAs at age 5}
dat %>%
  rename(`Social isolation age 5` = isolation_combined_05) %>%
  freq(`Social isolation age 5`, 
       cumul = FALSE, 
       display.type = FALSE, 
       headings = FALSE, 
       style = "rmarkdown")
```
Zero NAs at age 5. 

### Frequencies at age 7
```{r NAs at age 7}
dat %>%
  rename(`Social isolation age 7` = isolation_combined_07) %>%
  freq(`Social isolation age 7`, 
       cumul = FALSE, 
       display.type = FALSE, 
       headings = FALSE, 
       style = "rmarkdown")
```
54 NAs at age 7. 

### Frequencies at age 10
```{r NAs at age 10}
dat %>%
  rename(`Social isolation age 10` = isolation_combined_10) %>%
  freq(`Social isolation age 10`, 
       cumul = FALSE, 
       display.type = FALSE, 
       headings = FALSE, 
       style = "rmarkdown")
```
94 NAs at age 10. 

### Frequencies at age 12
```{r NAs at age 12}
freq(dat$isolation_combined_12, 
       cumul = FALSE, 
       display.type = FALSE, 
       headings = FALSE, 
       style = "rmarkdown")
```
87 NAs at age 12. 

## Groups for participants missing 2 or 3 time points

```{r create variables for different group combinations, echo=TRUE}
# group with 18 missing 3 time points vs everyone else
dat <- dat %>%
  mutate(
    missing_3_time_points = 
      if_else(
        na.per.person.si > 2, 
        1,
        0) %>%
      recode_factor(
        "0" = "No",
        "1" = "Yes"
      )
  )

# create group of just 18 who have 3 missing time points
dat.18 <- filter(dat, missing_3_time_points == "Yes")

# excluding the 18 missing 3 time points, those with 2 missing time points vs every one else
dat.missing.3time <- filter(dat,
                      na.per.person.si != 3)

dat.missing.3time <- dat.missing.3time %>%
  mutate(
    missing_2_time_points = 
      if_else(
        na.per.person.si > 1 & na.per.person.si < 3, 
        1,
        0) %>%
      recode_factor(
        "0" = "No",
        "1" = "Yes"
      )
  )
```

# Differences for those who have missing social isolation data {.tabset .tabset-fade}

Results from t-tests and Chi square tests assessing significant group differences for those who were missing three time points, and for those who were missing two time pionts. 46 people were missing two of the four time points. 18 people were missing three out of four time points. 

```{r functions}
# This function creates a results sentence for all t-tests and cho square tests to follow. 
test.sentence <- function(test="t"|"X-squared", test.data, time.point.missing = "two"|"three", variable){
  part1 <- cat(paste0(variable, " at age five was not significantly different for those missing ", time.point.missing, " time points, compared to everyone else (", test, " =", round(test.data$statistic, 3), ", df=", round(test.data$parameter, 3), ", p=", round(test.data$p.value, 3), "). "))
  if (test== "X-squared"){print(part1)}
  if (test=="t"){
  part1
  cat(paste0(part1, " Mean for group **not** missing ", time.point.missing, " time points = ", round(test.data$estimate[1], 3), ", mean for group missing ",time.point.missing, " three time points = ", round(test.data$estimate[2], 3), ". "))
  }
}
```

## Social isolation at age 5
```{r differences in isolation based on missingness, results='asis'}
ttest_SI_3time <- t.test( 
  isolation_combined_05 ~ missing_3_time_points,
  data = dat,
  alternative = "two.sided",
  conf.level = 0.95)
test.sentence(test = "t", test.data = ttest_SI_3time, time.point.missing = "three", variable = "Social isolation")

ttest_SI_2time <- t.test( 
  isolation_combined_05 ~ missing_2_time_points,
  data = dat.missing.3time,
  alternative = "two.sided",
  conf.level = 0.95)
test.sentence(test = "t", test.data = ttest_SI_2time, time.point.missing = "two", variable = "Social isolation")
```

## SES at age 5
```{r differences in SES based on missingness}
chisq_SES_3time <- chisq.test(dat$SES,dat$missing_3_time_points) # nonsig
test.sentence(test = "X-squared", test.data = chisq_SES_3time, time.point.missing = "three", variable = "Socioeconomic status")

chisq_SES_2time <- chisq.test(dat.missing.3time$SES, dat.missing.3time$missing_2_time_points) # nonsig
test.sentence(test = "X-squared", test.data = chisq_SES_2time, time.point.missing = "two", variable = "Socioeconomic status")
```

## IQ at age 5
```{r differences in IQ based on missingness}
ttest_IQ_3time <- t.test( #nonsig
  IQ_05 ~ missing_3_time_points,
  data = dat,
  alternative = "two.sided",
  conf.level = 0.95) 
test.sentence(test = "t", test.data = ttest_IQ_3time, time.point.missing = "three", variable = "IQ")

ttest_IQ_2time <- t.test( #SIGNIFICANT - missing two time points has lower IQ
  IQ_05 ~ missing_2_time_points,
  data = dat.missing.3time,
  alternative = "two.sided",
  conf.level = 0.95) 
paste0("IQ at age five was **significantly** different for those missing two time points, compared to everyone else (t=", round(ttest_IQ_2time$statistic, 3), ", df=", round(ttest_IQ_2time$parameter, 3), ", p=", round(ttest_IQ_2time$p.value, 3), "). Mean for group **not** missing two time points = ", round(ttest_IQ_2time$estimate[1], 3), ", mean for group missing two time points = ", round(ttest_IQ_2time$estimate[2], 3))
```

## Internalising problems at age 5
```{r differences in internalising problems at age 5 based on missingness}
ttest_int_3time <- t.test( #nonsig
  internalising_combined_excl_sis_05 ~ missing_3_time_points,
  data = dat,
  alternative = "two.sided",
  conf.level = 0.95) 
test.sentence(test = "t", test.data = ttest_int_3time, time.point.missing = "three", variable = "Internalising behaviours")

ttest_int_2time <- t.test( #nonsig
  internalising_combined_excl_sis_05 ~ missing_2_time_points,
  data = dat.missing.3time,
  alternative = "two.sided",
  conf.level = 0.95) 
test.sentence(test = "t", test.data = ttest_int_2time, time.point.missing = "two", variable = "Internalising behaviours")
```

## Externalising problems at age 5
```{r differences in externalising problems at age 5 based on missingness}
ttest_ext_3time <- t.test( #nonsig
  externalising_combined_05 ~ missing_3_time_points,
  data = dat,
  alternative = "two.sided",
  conf.level = 0.95) 
test.sentence(test = "t", test.data = ttest_ext_3time, time.point.missing = "three", variable = "Externalising behaviours")

ttest_ext_2time <- t.test( #nonsig
  externalising_combined_05 ~ missing_2_time_points,
  data = dat.missing.3time,
  alternative = "two.sided",
  conf.level = 0.95) 
test.sentence(test = "t", test.data = ttest_ext_2time, time.point.missing = "two", variable = "Externalising behaviours")
```

# Missingness differences at age 18

We found no differences between those who did and did not take part at age 18 in terms of socioeconomic status (SES) assessed when the cohort was initially defined (χ2 = 0.86, p = 0.65), age-5 IQ scores (t = 0.98, p = 0.33), or age-5 emotional or behavioural problems (t = 0.40, p = 0.69 and t = 0.41, p = 0.68, respectively). There were 2066 complete responses at age 18 (92.56% retention). 
 
```{r took part age age 18 - social isolation}
# check the freq 
employment.freq <- freq(dat$not_in_employment_education_18,
     cumul = FALSE)

# create data set for those who did and did not take part at age 18 
## 166 participants
dat_not_age_18 <- dat %>%
  filter(is.na(not_in_employment_education_18))
## 2066 participants
dat_age_18 <- dat %>%
  filter(!is.na(not_in_employment_education_18))

# create variable in original data set based on cut off
dat <- dat %>%
  mutate(
    took_part_at_18 = 
      if_else(
        !is.na(not_in_employment_education_18), 
        1,
        0) %>%
      recode_factor(
        "0" = "No",
        "1" = "Yes"
      )
  )

# check
# table(dat$took_part_at_18)

# Social isolation - t-test - nonsig
ttest_SI_age18 <- t.test( 
  isolation_combined_05 ~ took_part_at_18,
  data = dat,
  alternative = "two.sided",
  conf.level = 0.95) 

# IQ - t-test - nonsig
ttest_IQ_age18 <- t.test( 
  IQ_05 ~ took_part_at_18,
  data = dat,
  alternative = "two.sided",
  conf.level = 0.95) 

# externalising - t-test - nonsig
ttest_ext_age18 <- t.test( 
  externalising_combined_05 ~ took_part_at_18,
  data = dat,
  alternative = "two.sided",
  conf.level = 0.95) 

# internalising - t-test - nonsig
ttest_int_age18 <- t.test( 
  internalising_combined_excl_sis_05 ~ took_part_at_18,
  data = dat,
  alternative = "two.sided",
  conf.level = 0.95) 

# SES - chi square
chisq_SES_age18 <- chisq.test(dat$SES,
                              dat$took_part_at_18) # nonsig

```

# Means across all time points {.tabset .tabset-fade}

```{r convert data to long format, include=FALSE}
#convert data to long format for graphs
dat_long<- dat %>%
  gather(
    key = "time_point_raw",
    value = "social_isolation",
      all_of(combined_social_isolation)) %>%
  select(
    id,
    sex,
    zygosity_binary,
    SES,
    SES_ordered,
    time_point_raw,
    social_isolation)

head(dat_long)

#create variable that has just the number for the time point
dat_long <- dat_long %>%
  mutate(time_point =
  recode_factor(time_point_raw, 
                "isolation_combined_05" = "5", 
                "isolation_combined_07" = "7", 
                "isolation_combined_10" = "10", 
                "isolation_combined_12" = "12")) %>%
  select(
    id,
    sex,
    zygosity_binary,
    SES,
    SES_ordered,
    time_point,
    social_isolation)

head(dat_long)
```

```{r means across time plot}
#summarize the data fir the plot 
dat_summary.isolation <- summarySE(dat_long, 
                            measurevar = "social_isolation", 
                            groupvars = c("time_point"),
                            na.rm=TRUE)
#create plot
diff.plot <- ggplot(dat_summary.isolation, 
       aes(x = time_point, 
           y = social_isolation,
           group = 1)) + 
    geom_errorbar(aes(ymin = social_isolation - se, 
                      ymax = social_isolation + se), 
                  colour = "black", 
                  width = 0.1) +
    geom_line(colour = "red") +
    geom_point(size = 3, 
               shape = 21, 
               fill = "white") + # 21 is filled circle
    xlab("Time point") +
    ylab("Social isolation") +
    ggtitle("Average differences in social isolation at time point 5, 7, 10 and 12") +
    expand_limits(y = 0) +                     
    scale_y_continuous(expand = c(0.1,0.1),
                     limits = c(0,12),
                     breaks = seq(0, 12, 2)) +
  theme_bw()

#create scaled plot
diff.plot.scaled <- diff.plot+ 
  scale_y_continuous(expand = c(0.1,0.1),
                     limits = c(0,2),
                     breaks = seq(0, 2, 0.25)) +
   ggtitle("Scaled average differences in social isolation at time point 5, 7, 10 and 12")
diff.plot.scaled
```

## Combined mother/teacher report of social isolation
```{r means for social isolation combined, message=FALSE, warning=FALSE}
dat %>%
  select(`Age five` = isolation_combined_05,
         `Age seven` = isolation_combined_07,
         `Age ten` = isolation_combined_10,
         `Age twelve` = isolation_combined_12) %>%
  descr( 
      headings = FALSE,
      stats =  c("mean", "sd", "min", "max", "skewness", "kurtosis", "n.valid", "pct.valid"), 
      style = "rmarkdown") 
```

## Mother report of social isolation
```{r means for social isolation mother report, message=FALSE, warning=FALSE}
dat %>%
  select(`Age five` = isolation_mother_05,
         `Age seven` = isolation_mother_07,
         `Age ten` = isolation_mother_10,
         `Age twelve` = isolation_mother_12) %>%
  descr( 
      headings = FALSE,
      stats =  c("mean", "sd", "min", "max", "skewness", "kurtosis", "n.valid", "pct.valid"), 
      style = "rmarkdown") 
```

## Teacher report
```{r means for social isolation teacher report, message=FALSE}
dat %>%
  select(`Age five` = isolation_teacher_05,
         `Age seven` = isolation_teacher_07,
         `Age ten` = isolation_teacher_10,
         `Age twelve` = isolation_teacher_12) %>%
  descr( 
      headings = FALSE,
      stats =  c("mean", "sd", "min", "max", "skewness", "kurtosis", "n.valid", "pct.valid"), 
      style = "rmarkdown") 
```


