---
title: "Preliminary descriptives of E-risk social isolation"
author: "Katherine N Thompson"
date: "13th Jan 2021"
output:  
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: cerulean
code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      comment = NA,
                      prompt = FALSE,
                      cache = FALSE,
                      message = FALSE,
                      warning = FALSE)


options(bitmapType = 'quartz') # to render fonts better
```

```{r Clear global environment}
remove(list = ls())
```

```{r Current date}
date = Sys.Date()
```

```{r palletes}
palette2 <- c("#78D9C5","#F5BE5E")
palette3 <- c("#78D9C5","#F5BE5E","#EEB6E9")
palette4 <- c("#78D9C5","#F5BE5E","#EEB6E9","#DBDB73")
palette5 <- c("#78D9C5","#F5BE5E","#EEB6E9","#DBDB73","#FFED98")
palette6 <- c("#78D9C5","#F5BE5E","#EEB6E9","#DBDB73","#FFED98","#BFD2EB")
```

# Packages 

```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(skimr)
library(ggpubr)
library(ggplot2)
library(tidyr)
library(gplots)
library(Rmisc)
library(tidyverse)
```

# Source data path

```{r source the data file path}
#source raw data directory: data_raw and data included
source("../../../../../isolation_trajectories_data_path.R")
```

# Read in data - preprocessed 

```{r read in data and check variable names, include=FALSE}
dat <- readRDS(file = paste0(data_path, "preprocessed_isolation_trajectories_Jan2021_full_sample.rds"))
colnames(dat)
```

```{r prepare item lists}
#All lists for isolation and grouping variables
#Only using elder twin variables as each twin is included in the data set as an individual. Without doing twin analyses, the elder and younger variables show the same information.
combined_social_isolation <- c("isolation_combined_05",
                              "isolation_combined_07",
                              "isolation_combined_10",
                              "isolation_combined_12")

teacher_social_isolation <- c("isolation_teacher_05",
                              "isolation_teacher_07",
                              "isolation_teacher_10",
                              "isolation_teacher_12")

mother_social_isolation <- c("isolation_mother_05",
                              "isolation_mother_07",
                              "isolation_mother_10",
                              "isolation_mother_12")
```

```{r Half-violin plot}
#source the half/flat function for the violin plot
source("https://raw.githubusercontent.com/datavizpyr/data/master/half_flat_violinplot.R")

#theme for ggplotting
theme_personal <-  theme(
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    axis.title.y = element_blank(),
    legend.background = element_blank(),
    legend.box.background = element_blank(),
    panel.background = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(
      colour = "gray",
      linetype = "dashed",
      size = 0.2
      ),
    axis.ticks = element_blank()
    )
```

# Missingness descriptives

```{r social isolation missingness}
#look at the frequency of missingness
freq(dat$na.per.person.si,
     cumul = FALSE)

#look at the frequency at 5 years only
freq(dat$isolation_combined_05,
     cumul = FALSE)

#look at the frequency at 7 years only
freq(dat$isolation_combined_07,
     cumul = FALSE)

#look at the frequency at 10 years only
freq(dat$isolation_combined_10,
     cumul = FALSE)

#look at the frequency at 12 years only
freq(dat$isolation_combined_12,
     cumul = FALSE)
```

## Create groups for those missing 3 time points or 2 time points. 

```{r create variables for different group combinations}
# group with 18 missing 3 time points vs everyone else
dat <- dat %>%
  mutate(
    missing_3_time_points = 
      if_else(
        na.per.person.si > 2, 
        1,
        0) %>%
      recode_factor(
        "0" = "No",
        "1" = "Yes"
      )
  )

# create group of just 18 who have 3 missing time points
dat.18 <- filter(dat, missing_3_time_points == "Yes")

# excluding the 18 missing 3 time points, those with 2 missing time points vs every one else
dat.missing.3time <- filter(dat,
                      na.per.person.si != 3)

dat.missing.3time <- dat.missing.3time %>%
  mutate(
    missing_2_time_points = 
      if_else(
        na.per.person.si > 1 & na.per.person.si < 3, 
        1,
        0) %>%
      recode_factor(
        "0" = "No",
        "1" = "Yes"
      )
  )
```

# Differences on social isolation, SES, IQ, internalising problems, behavioural problems for those who had data missing and those who didn't.

```{r check differences in missingness between the 4 groups}
# Social isolation at age 5
ttest_SI_3time <- t.test( 
  isolation_combined_05 ~ missing_3_time_points,
  data = dat,
  alternative = "two.sided",
  conf.level = 0.95)

ttest_SI_2time <- t.test( 
  isolation_combined_05 ~ missing_2_time_points,
  data = dat.missing.3time,
  alternative = "two.sided",
  conf.level = 0.95) 

# SES - chi square
chisq_SES_3time <- chisq.test(dat$SES,dat$missing_3_time_points) # nonsig
chisq_SES_2time <- chisq.test(dat.missing.3time$SES, dat.missing.3time$missing_2_time_points) # nonsig

# IQ - t-test
ttest_IQ_3time <- t.test( #nonsig
  IQ_05 ~ missing_3_time_points,
  data = dat,
  alternative = "two.sided",
  conf.level = 0.95) 

ttest_IQ_2time <- t.test( #SIGNIFICANT - missing two time points has lower IQ
  IQ_05 ~ missing_2_time_points,
  data = dat.missing.3time,
  alternative = "two.sided",
  conf.level = 0.95) 

# internalizing problems - t-test
ttest_int_3time <- t.test( #nonsig
  internalising_combined_excl_sis_05 ~ missing_3_time_points,
  data = dat,
  alternative = "two.sided",
  conf.level = 0.95) 

ttest_int_2time <- t.test( #nonsig
  internalising_combined_excl_sis_05 ~ missing_2_time_points,
  data = dat.missing.3time,
  alternative = "two.sided",
  conf.level = 0.95) 

# Behavioural problems - externalising?? - t-test
ttest_ext_3time <- t.test( #nonsig
  externalising_combined_05 ~ missing_3_time_points,
  data = dat,
  alternative = "two.sided",
  conf.level = 0.95) 

ttest_ext_2time <- t.test( #nonsig
  externalising_combined_05 ~ missing_2_time_points,
  data = dat.missing.3time,
  alternative = "two.sided",
  conf.level = 0.95) 
```

# Missingness differences at age 18

What Matthews found for differences between those who did and did not take part at age 18: There were no differences between those who did and did not take part at age 18 in terms of  socioeconomic status (SES) assessed when the cohort was initially defined (χ2 = 0.86, p = 0.65), age-5 IQ scores (t = 0.98, p = 0.33), or age-5 emotional or behavioural problems (t = 0.40, p = 0.69 and t = 0.41, p = 0.68, respectively)
 - Here I'm using not_in_employment_education_18 as the indicator variable for if they took part at age 18 or not (2066 complete responses)

```{r took part age age 18 - social isolation}
# check the freq 
freq(dat$not_in_employment_education_18,
     cumul = FALSE)

# create data set for those who did and did not take part at age 18 
## 166 participants
dat_not_age_18 <- dat %>%
  
  filter(is.na(not_in_employment_education_18))
## 2066 participants
dat_age_18 <- dat %>%
  filter(!is.na(not_in_employment_education_18))

# create variable in original data set based on cut off
dat <- dat %>%
  mutate(
    took_part_at_18 = 
      if_else(
        !is.na(not_in_employment_education_18), 
        1,
        0) %>%
      recode_factor(
        "0" = "No",
        "1" = "Yes"
      )
  )

# check
table(dat$took_part_at_18)

# Social isolation - t-test - nonsig
ttest_SI_age18 <- t.test( 
  isolation_combined_05 ~ took_part_at_18,
  data = dat,
  alternative = "two.sided",
  conf.level = 0.95) 

# IQ - t-test - nonsig
ttest_IQ_age18 <- t.test( 
  IQ_05 ~ took_part_at_18,
  data = dat,
  alternative = "two.sided",
  conf.level = 0.95) 

# externalising - t-test - nonsig
ttest_ext_age18 <- t.test( 
  externalising_combined_05 ~ took_part_at_18,
  data = dat,
  alternative = "two.sided",
  conf.level = 0.95) 

# internalising - t-test - nonsig
ttest_int_age18 <- t.test( 
  internalising_combined_excl_sis_05 ~ took_part_at_18,
  data = dat,
  alternative = "two.sided",
  conf.level = 0.95) 

# SES - chi square
chisq_SES_age18 <- chisq.test(dat$SES,
                              dat$took_part_at_18) # nonsig

```


# Convert data to long format

```{r convert data to long format, include=FALSE}
#convert to long format 
dat_long<- dat %>%
  gather(
    key = "time_point_raw",
    value = "social_isolation",
      all_of(combined_social_isolation)) %>%
  select(
    id,
    sex,
    zygosity_binary,
    SES,
    SES_ordered,
    time_point_raw,
    social_isolation)

head(dat_long)

#create variable that has just the number for the time point
dat_long <- dat_long %>%
  mutate(time_point =
  recode_factor(time_point_raw, 
                "isolation_combined_05" = "5", 
                "isolation_combined_07" = "7", 
                "isolation_combined_10" = "10", 
                "isolation_combined_12" = "12")) %>%
  select(
    id,
    sex,
    zygosity_binary,
    SES,
    SES_ordered,
    time_point,
    social_isolation)

head(dat_long)
```

# Means across all time points {.tabset .tabset-fade}

```{r means across time plot}
#summarize the data fir the plot 
dat_summary.isolation <- summarySE(dat_long, 
                            measurevar = "social_isolation", 
                            groupvars = c("time_point"),
                            na.rm=TRUE)
#create plot
diff.plot <- ggplot(dat_summary.isolation, 
       aes(x = time_point, 
           y = social_isolation,
           group = 1)) + 
    geom_errorbar(aes(ymin = social_isolation - se, 
                      ymax = social_isolation + se), 
                  colour = "black", 
                  width = 0.1) +
    geom_line(colour = "red") +
    geom_point(size = 3, 
               shape = 21, 
               fill = "white") + # 21 is filled circle
    xlab("Time point") +
    ylab("Social isolation") +
    ggtitle("Average differences in social isolation at time point 5, 7, 10 and 12") +
    expand_limits(y = 0) +                     
    scale_y_continuous(expand = c(0.1,0.1),
                     limits = c(0,12),
                     breaks = seq(0, 12, 2)) +
  theme_bw()
diff.plot

#create scaled plot
diff.plot.scaled <- diff.plot+ 
  scale_y_continuous(expand = c(0.1,0.1),
                     limits = c(0,2),
                     breaks = seq(0, 2, 0.25)) +
   ggtitle("Scaled average differences in social isolation at time point 5, 7, 10 and 12")
diff.plot.scaled
```

## Combined report
```{r means for social isolation combined}
descr(x = dat[,combined_social_isolation], 
      headings = FALSE,
      stats =  c("mean", "sd", "min", "med", "max", "skewness", "kurtosis", "n.valid", "pct.valid")) %>%
  print(method = "render")
```

## Mother report
```{r means for social isolation mother report}
descr(x = dat[,mother_social_isolation], 
      headings = FALSE,
      stats =  c("mean", "sd", "min", "med", "max", "skewness", "kurtosis", "n.valid", "pct.valid")) %>%
  print(method = "render") 
```

## Teacher report
```{r means for social isolation teacher report}
descr(x = dat[,teacher_social_isolation], 
      headings = FALSE,
      stats =  c("mean", "sd", "min", "med", "max", "skewness", "kurtosis", "n.valid", "pct.valid")) %>%
  print(method = "render")
```


