---
title: 'Trajectory descriptives'
author: "Katherine N Thompson"
date: "02 Feb 2021"
output:
  word_document:
    toc: yes
    toc_depth: '2'
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
    number_sections: no
    highlight: monochrome
    theme: cerulean
code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      comment = NA,
                      prompt = FALSE,
                      cache = FALSE,
                      message = FALSE,
                      warning = FALSE)


options(bitmapType = 'quartz') # to render fonts better
```

```{r Clear global environment}
remove(list = ls())
```

```{r Current date}
date = Sys.Date()
```

```{r Load packages}
library(knitr)
library(summarytools)
library(MplusAutomation)
library(tidyr)
library(tidyverse)
library(plyr)
library(dplyr) #conflicts with tidyverse for e.g. rename and row_number
```

```{r palette}
palette3 <- c("#78D9C5","#F5BE5E","#EEB6E9")
```

```{r source the data file path}
#source raw data directory: data_raw and data included
source("../../../../../isolation_trajectories_data_path.R")
```

# Read in descriptives output file

```{r read in prior descriptives data and check variable names, include=FALSE}
dat.descriptives <- readRDS(file = paste0(data_path, "preprocessed_isolation_trajectories_Jan2021_full_sample.rds"))
colnames(dat.descriptives)
```

# Read in class probability file 

```{r read in txt file with probabilities, include=FALSE}
dat.tra.prob.full <- read.table(paste0(mplus_GMM_clustered_full_output_data_path, "GMM_SI_3Cl_full_sample_clustered.txt"), 
                            header = FALSE, 
                            col.names = c("SISOE5", # original social isolation score at age 5
                                         "SISOE7",  # original social isolation score at age 7
                                         "SISOE10", # original social isolation score at age 10
                                         "SISOE12", # original social isolation score at age 12
                                         "I",       # intercept
                                         "S",       # slope
                                         "C_I",     # factor scores based on most likely class membership
                                         "C_S",     # factor scores based on most likely class membership
                                         "CPROB1",  # class probability 1
                                         "CPROB2",  # class probability 2
                                         "CPROB3",  # class probability 3
                                         "C",       # class 
                                         "ID",      # ID
                                         "FAMILYID" # Family ID - twin clustering  
                                       ))   
#check 
dat.tra.prob.full
```

## Select the variables needed from the Mplus output file 

```{r select relevant variables, include=FALSE}
dat.tra.prob <- dat.tra.prob.full %>%
  select(id = ID, 
         prob1 = CPROB1,
         prob2 = CPROB2,
         prob3 = CPROB3,
         class = C
         )
#check 
dat.tra.prob

#create list of data frames to merge
dataframe_list <- list(
  dat.descriptives,
  dat.tra.prob
)

#merge data frames
dat <- plyr::join_all(
  dataframe_list,
  by = "id" # Alternatively you can join by several columns
  )

#check
colnames(dat)
```

# Rename class variable to shape of trajectory

```{r rename class variable}
dat <- dat %>%
  mutate(class_renamed =
    recode_factor(class,
      "1" = "Low stable",
      "2" = "Increasing",
      "3" = "Decreasing"
    )
  )
table(dat$class_renamed)
```

# Spaghetti plot of social isolation over time

## Convert data to long format

```{r convert data to long format, include=FALSE}
combined_social_isolation <- c("isolation_combined_05",
                              "isolation_combined_07",
                              "isolation_combined_10",
                              "isolation_combined_12")

#convert to long format 
dat_long <- dat %>%
  gather(
    key = "time_point_raw",
    value = "social_isolation",
      all_of(combined_social_isolation)) %>%
  select(
    id,
    sex,
    time_point_raw,
    social_isolation,
    class_renamed)

head(dat_long)

#create variable that has just the number for the time point
dat_long <- dat_long %>%
  mutate(time_point =
  recode_factor(time_point_raw, 
                "isolation_combined_05" = "5", 
                "isolation_combined_07" = "7", 
                "isolation_combined_10" = "10", 
                "isolation_combined_12" = "12")) %>%
  select(
    id,
    sex,
    time_point,
    social_isolation,
    class_renamed)

head(dat_long)
```

```{r spaghetti plot}
spaghetti_plot <- ggplot(data = dat_long,
       aes(x = time_point,
           y = social_isolation,
           group = factor(id),
           color = class_renamed)) +
  geom_line() +
  scale_color_manual(name = "Class", values = palette3) +
  labs(y = "Social isolation score",
       x = "Age (years)") +
  scale_x_discrete(expand = c(0.05,0.05), breaks = c("5", "7", "10", "12")) +
  scale_y_continuous(expand = c(0.05,0.05), breaks = c(0,2,4,6,8,10,12)) +
  theme(
        panel.grid.major.y = element_line(size = 0.2,linetype = 'dashed',colour = "gray"),
        axis.title.x = element_text(size = 12,face="bold"),
        axis.title.y = element_text(size = 12,face="bold"),
        axis.text.x = element_text(colour = "black", size = 10),
        axis.text.y = element_text(colour = "black", size = 10),
        panel.background = element_blank(),
        legend.position = "bottom")

ggsave(
  "spaghetti_plot_three_class_colour_coded.jpeg",
  plot = spaghetti_plot,
  device = "jpeg",
  path = paste0(graph_save_data_path, "trajectories/"),
  width = 7,
  height = 5
)
```

# Descriptives by class

```{r ethnicity descriptives by class, include=FALSE}
# calculate Ns and %s for each class by ethnicity
ethnicity_descriptives_stable <- dat %>% #create object to bind later on
  dplyr::count(class_renamed, ethnicity) %>%
  filter(class_renamed == "Low stable") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

ethnicity_descriptives_decreasing <- dat %>% #create object to bind later on
  dplyr::count(class_renamed, ethnicity) %>%
  filter(class_renamed == "Decreasing") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

ethnicity_descriptives_increasing <- dat %>% #create object to bind later on
  dplyr::count(class_renamed, ethnicity) %>%
  filter(class_renamed == "Increasing") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

eth_df <- list(ethnicity_descriptives_stable, ethnicity_descriptives_decreasing, ethnicity_descriptives_increasing)
ethnicity_descriptives <- join_all(eth_df,
                                   by = "ethnicity",
                                   type = "left")


# to avoid NAs in later descriptives table - change Na to 0
ethnicity_descriptives$`n_Increasing` <- ethnicity_descriptives$`n_Increasing` %>% replace_na(0)
ethnicity_descriptives$`Prop_Increasing` <- ethnicity_descriptives$`Prop_Increasing` %>% replace_na(0)

ethnicity_descriptives <- ethnicity_descriptives %>%
  mutate(Total = `n_Increasing` + `n_Decreasing` + `n_Low stable`,
         Prop_total = round(Total/2232*100, digits = 2))

# rename columns 
ethnicity_descriptives <- ethnicity_descriptives %>%
  select(
    "Descriptive" = ethnicity,
    "Low stable" = `n_Low stable`,
    "% Low stable" = `Prop_Low stable`,
    "Increasing" = `n_Increasing`,
    "% Increasing" = `Prop_Increasing`,
    "Decreasing" = `n_Decreasing`,
    "% Decreasing" = `Prop_Decreasing`,
    "Total" = Total,
    "% Total" = Prop_total)
ethnicity_descriptives
```

## SES descriptives stratified by class 

```{r SES descriptives by class, include=FALSE}
SES_descriptives_stable <- dat %>% 
  dplyr::count(class_renamed, SES) %>%
  filter(class_renamed == "Low stable") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

SES_descriptives_decreasing <- dat %>% 
  dplyr::count(class_renamed, SES) %>%
  filter(class_renamed == "Decreasing") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

SES_descriptives_increasing <- dat %>% 
  dplyr::count(class_renamed, SES) %>%
  filter(class_renamed == "Increasing") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

ses_df <- list(SES_descriptives_stable, SES_descriptives_decreasing, SES_descriptives_increasing)
SES_descriptives <- join_all(ses_df,
                                   by = "SES",
                                   type = "left")

SES_descriptives <- SES_descriptives %>%
  mutate(Total = `n_Increasing` + `n_Decreasing` + `n_Low stable`,
         Prop_total = round(Total/2232*100, digits = 2)) %>% 
  select(
    "Descriptive" = SES,
    "Low stable" = `n_Low stable`,
    "% Low stable" = `Prop_Low stable`,
    "Increasing" = `n_Increasing`,
    "% Increasing" = `Prop_Increasing`,
    "Decreasing" = `n_Decreasing`,
    "% Decreasing" = `Prop_Decreasing`,
    "Total" = Total,
    "% Total" = Prop_total)
SES_descriptives
```

## Zygosity descriptives stratified by class 

```{r zygosity descriptives by class, include=FALSE}
zygosity_binary_descriptives_stable <- dat %>% 
  dplyr::count(class_renamed, zygosity_binary) %>%
  filter(class_renamed == "Low stable") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

zygosity_binary_descriptives_decreasing <- dat %>% 
  dplyr::count(class_renamed, zygosity_binary) %>%
  filter(class_renamed == "Decreasing") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

zygosity_binary_descriptives_increasing <- dat %>% 
  dplyr::count(class_renamed, zygosity_binary) %>%
  filter(class_renamed == "Increasing") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

zyg_df <- list(zygosity_binary_descriptives_stable, zygosity_binary_descriptives_decreasing, zygosity_binary_descriptives_increasing)
zygosity_binary_descriptives <- join_all(zyg_df,
                                   by = "zygosity_binary",
                                   type = "left")

zygosity_binary_descriptives <- zygosity_binary_descriptives %>%
  mutate(Total = `n_Increasing` + `n_Decreasing` + `n_Low stable`,
         Prop_total = round(Total/2232*100, digits = 2)) %>% 
  select(
    "Descriptive" = zygosity_binary,
    "Low stable" = `n_Low stable`,
    "% Low stable" = `Prop_Low stable`,
    "Increasing" = `n_Increasing`,
    "% Increasing" = `Prop_Increasing`,
    "Decreasing" = `n_Decreasing`,
    "% Decreasing" = `Prop_Decreasing`,
    "Total" = Total,
    "% Total" = Prop_total)
zygosity_binary_descriptives
```

## Sex descriptives stratified by class 

```{r sex descriptives by class, include=FALSE}
sex_descriptives_stable <- dat %>% 
  dplyr::count(class_renamed, sex) %>%
  filter(class_renamed == "Low stable") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop))

sex_descriptives_decreasing <- dat %>% 
  dplyr::count(class_renamed, sex) %>%
  filter(class_renamed == "Decreasing") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop))

sex_descriptives_increasing <- dat %>% 
  dplyr::count(class_renamed, sex) %>%
  filter(class_renamed == "Increasing") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop))

sex_df <- list(sex_descriptives_stable, sex_descriptives_decreasing, sex_descriptives_increasing)
sex_descriptives <- join_all(sex_df,
                                   by = "sex",
                                   type = "left")

sex_descriptives <- sex_descriptives %>%
  mutate(Total = `n_Increasing` + `n_Decreasing` + `n_Low stable`,
         Prop_total = round(Total/2232*100, digits = 2)) %>% 
  select(
    "Descriptive" = sex,
    "Low stable" = `n_Low stable`,
    "% Low stable" = `Prop_Low stable`,
    "Increasing" = `n_Increasing`,
    "% Increasing" = `Prop_Increasing`,
    "Decreasing" = `n_Decreasing`,
    "% Decreasing" = `Prop_Decreasing`,
    "Total" = Total,
    "% Total" = Prop_total)
sex_descriptives
```

## Combine all descriptives into one table

```{r combine descriptives, include=FALSE}
class.descriptives <- bind_rows(
  sex_descriptives,
  zygosity_binary_descriptives,
  SES_descriptives,
  ethnicity_descriptives
)

class.descriptives <- class.descriptives %>%
  mutate(
    Category = 
      c("Sex", "",
        "Zygosity", "", 
        "SES", "", "",
        "Ethnicity", "", "", "", "")) %>%
  select(
    Category,
    Descriptive,
    `Low stable`,
    `% Low stable`,
    `Increasing`,
    `% Increasing`,
    `Decreasing`,
    `% Decreasing`,
    Total,
    `% Total`
  )
```

```{r all descriptives by class}
#final descriptives by class table
kable(class.descriptives)
```

# *Adapt variables for regression analysis*

Need to relevel the class variable to use the Low Stable class as the reference group in all regressions. 

```{r relevel class, include=FALSE}
library(nnet)
# check the biggest group in the class to be the reference group
table(dat$class_renamed)

# relevel class_factor
dat <- dat %>% 
  mutate(
    class_reordered =
      relevel(
        class_renamed,
        ref = "Low stable",
        first = TRUE, #levels in ref come first
        collapse = "+", #String used when constructing names for combined factor levels
        xlevels = TRUE #levels maintained even if not actually occurring
        
      )
  )
# check
table(dat$class_reordered)
```

# Class probability statistics

```{r read in data for all traj models, include=FALSE}
class.all <- readModels(paste0(mplus_GMM_clustered_full_output_data_path), recursive = TRUE)
```

## Classification probabilities 

For the diagonal, this represents the probability given that the an observation is a part of a specific latent class that you will classify it as said class.

*Note* this is not average probabilities - this table wasn't available in Mplus version 8. 

```{r extract classification probs for 3 class model, include=FALSE}
# extract the classification probabilities for the most likely class memberships (column) by latent class (row). This shows the uncertainty rate of the probabilities. If it was perfect - the diagonal would be 1. E.g. for Class 1 - 98.3% of individuals fit that category. 
class.probs <- data.frame(class.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.output..isolation_3traj_full_sample_clustered.out$class_counts$classificationProbs.mostLikely)

# add column names
colnames(class.probs) <- c("Low stable","Increasing","Decreasing")

class.probs$Class <- c("Low stable","Increasing","Decreasing")

as.tibble(class.probs)

class.probs <- class.probs %>%
  tidyr::pivot_longer(-Class, 
               names_to = "class", 
               values_to = "Classification probabilities")

class.probs <- class.probs[-c(2:4,6:8),-c(2)]
```

```{r classification probabilities table, include=FALSE}
kable(class.probs)
```

## Class means

```{r extract means (TECH7 output) for each class}
# extract means for each class 
class.1.means <- as.tibble(class.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.output..isolation_3traj_full_sample_clustered.out$tech7$CLASS.1$classSampMeans)

class.2.means <- as.tibble(class.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.output..isolation_3traj_full_sample_clustered.out$tech7$CLASS.2$classSampMeans)

class.3.means <- as.tibble(class.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.output..isolation_3traj_full_sample_clustered.out$tech7$CLASS.3$classSampMeans)

class.means <- rbind(class.1.means,class.2.means,class.3.means)

class.means$Class <- c("Low stable","Increasing","Decreasing")

class.means <- class.means %>%
  select(
    Class,
    `Mean at age 5` = SISOE5,
    `Mean at age 7` = SISOE7,
    `Mean at age 10` = SISOE10,
    `Mean at age 12` = SISOE12
  )
```

```{r class means table}
kable(class.means)
```

# Na per person split by class

```{r na per person slipt by class}
ctable(dat$class_renamed, dat$na.per.person.si)
```

# Export the joined preprocessed and class data

```{r export new data files}
# save R data file
saveRDS(object = dat, file = paste0(data_path, "class_joined_preprocessed_isolation_data_full_sample.rds"))
```


