---
title: "GMM trajectory descriptives"
output:  
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: flatly
    code_folding: hide
    includes:
      after_body: footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = NA,
                      prompt = FALSE,
                      cache = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      results = 'asis')

options(bitmapType = 'quartz') # to render fonts better
```

```{r Clear global environment, include=FALSE}
remove(list = ls())
```

```{r Current date, include=FALSE}
date = Sys.Date()
```

```{r Load packages, include=FALSE}
library(knitr)
library(summarytools)
library(MplusAutomation)
library(tidyr)
library(tidyverse)
library(plyr)
library(dplyr) #conflicts with tidyverse for e.g. rename and row_number
```

```{r palette, include=FALSE}
palette3 <- c("#78D9C5","#F5BE5E","#EEB6E9")
palette_spag <- c(`Low stable` = "#78D9C5", Increasing = "#F5BE5E", Decreasing ="#EEB6E9", `Class 1` = "#19A487", `Class 2` = "#DE9310", `Class 3` = "#E570DA")
```

```{r source the data file path, include=FALSE}
#source raw data directory: data_raw and data included
source("../../../isolation_trajectories_data_path.R")
```

***

This page contains the code to combine the original Rmd dataset and the new Mplus output. We also include relevant descriptive statistics for the classes/trajectories. We have used the hard classification variable given by Mplus for our analyses. We conduct sensitivity analyses using the 3STEP approach, data is prepped for this at the end of this page. *Please note, if you are running this analysis for yourself, you will need to do this step after running your Mplus models and before analysing the data in Rmd*

***

# Combine Mplus and Rmd data

We have read in the processed R data and the Mplus model information to show who is has identified into which class. 

```{r read in prior descriptives data and check variable names}
dat.descriptives <- readRDS(file = paste0(data_path, "preprocessed_isolation_trajectories_Jan2021_full_sample.rds"))
```

The Mplus output file contains the raw original data, slope estimate, intercept estimate, factor scores based on the slope and intercept, class probabilities for each person per class, and the hard class classification. 

```{r read in txt file with probabilities}
dat.tra.prob.full <- read.table(paste0(mplus_GMM_clustered_full_output_data_path, "GMM_SI_3Cl_full_sample_clustered.txt"), 
                            header = FALSE, 
                            col.names = c("SISOE5", # original social isolation score at age 5
                                         "SISOE7",  # original social isolation score at age 7
                                         "SISOE10", # original social isolation score at age 10
                                         "SISOE12", # original social isolation score at age 12
                                         "I",       # intercept
                                         "S",       # slope
                                         "C_I",     # factor scores based on most likely class membership
                                         "C_S",     # factor scores based on most likely class membership
                                         "CPROB1",  # class probability 1
                                         "CPROB2",  # class probability 2
                                         "CPROB3",  # class probability 3
                                         "C",       # class 
                                         "ID",      # ID
                                         "FAMILYID" # Family ID - twin clustering  
                                       ))   
#check 
head(dat.tra.prob.full)
```

We will only need the classification probabilities and the hard classification variable. This is then combined with the rest of our Rmd data.

```{r select relevant variables}
dat.tra.prob <- dat.tra.prob.full %>%
  select(id = ID, 
         prob1 = CPROB1,
         prob2 = CPROB2,
         prob3 = CPROB3,
         class = C
         )
#check 
# dat.tra.prob

#create list of data frames to merge
dataframe_list <- list(
  dat.descriptives,
  dat.tra.prob
)

#merge data frames
dat <- plyr::join_all(
  dataframe_list,
  by = "id" # Alternatively you can join by several columns
  )

#check
# colnames(dat)
```

The class variable is then renamed to reflect the pattern seen in the classes identified. Here, we name them "Low stable", "Increasing", and "Decreasing" to reflect the pattern of each of the linear curves. 

```{r rename class variable}
dat <- dat %>%
  mutate(class_renamed =
    recode_factor(class,
      "1" = "Low stable",
      "2" = "Increasing",
      "3" = "Decreasing"
    )
  )

data.frame(table(dat$class_renamed))
```
***

# Spaghetti plot showing all social isolation trajectories over time

```{r convert data to long format}
combined_social_isolation <- c("isolation_combined_05",
                              "isolation_combined_07",
                              "isolation_combined_10",
                              "isolation_combined_12")

#convert to long format 
dat_long <- dat %>%
  gather(
    key = "time_point_raw",
    value = "social_isolation",
      all_of(combined_social_isolation)) %>%
  select(
    id,
    sex,
    time_point_raw,
    social_isolation,
    class_renamed)

#head(dat_long)

#create variable that has just the number for the time point
dat_long <- dat_long %>%
  mutate(time_point =
  recode_factor(time_point_raw, 
                "isolation_combined_05" = "5", 
                "isolation_combined_07" = "7", 
                "isolation_combined_10" = "10", 
                "isolation_combined_12" = "12")) %>%
  select(
    id,
    sex,
    time_point,
    social_isolation,
    class_renamed)

#head(dat_long)
```

```{r spaghetti plot A}
spaghetti_plot <- ggplot(data = dat_long,
       aes(x = time_point,
           y = social_isolation,
           group = factor(id),
           color = class_renamed)) +
  geom_line() +
  scale_color_manual(name = "Class", values = palette3) +
  labs(y = "Social isolation score",
       x = "Age (years)") +
  scale_x_discrete(expand = c(0.05,0.05), breaks = c("5", "7", "10", "12")) +
  scale_y_continuous(expand = c(0.05,0.05), breaks = c(0,2,4,6,8,10,12)) +
  theme(
        panel.grid.major.y = element_line(size = 0.2,linetype = 'dashed',colour = "gray"),
        axis.title.x = element_text(size = 12,face="bold"),
        axis.title.y = element_text(size = 12,face="bold"),
        axis.text.x = element_text(colour = "black", size = 10),
        axis.text.y = element_text(colour = "black", size = 10),
        panel.background = element_blank(),
        legend.position = "bottom")
```

The trajectory of each participant is plotted separately and colour coded depending on which class they were classified. Decreasing is indicated in pink, Increasing in Orange, and Low stable in blue. The average class trajectories are then plotted over the top. 

```{r read in data for spag plot, results='hide', message=FALSE}
#For class models, read in all output files within your folder that you have all your class models
si.classes.all <- readModels(paste0(mplus_GMM_clustered_full_output_data_path), recursive = TRUE) #Recursive means it reads data in sub-folders too
```

```{r spaghetti plot B prep}
#create three class data frame with means and variances
three_class_si <- as.data.frame(
  si.classes.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.output..isolation_3traj_full_sample_clustered.out$gh5$means_and_variances_data$y_estimated_means)

#name the variables
names(three_class_si) <- c("Class 1","Class 2","Class 3")

#create time point column 
three_class_si$time_point <- factor(c("5","7","10","12"))

#convert data to long format 
three_class_si <- three_class_si %>%
  pivot_longer(-time_point, 
               names_to = "class_renamed", 
               values_to = "social_isolation")

# create factor for Class variable
three_class_si <- three_class_si %>%
  mutate(class_renamed = 
              factor(class_renamed, 
                     levels = c("Class 1","Class 2","Class 3"))) 
```

```{r spaghetti plot B}
spaghetti_plotB <- ggplot(data = dat_long,
                         aes(x = time_point,
                             y = social_isolation,
                             color = class_renamed)) +
                    geom_line(aes(group = factor(id)), alpha = 0.9, size = 0.5) +
                    geom_line(data = three_class_si, aes(group = class_renamed), size = 1, alpha = 1) +
                    scale_color_manual(name = "class_renamed", values = palette_spag) +
                    labs(y = "Social isolation score",
                         x = "Age (years)") +
                    scale_x_discrete(expand = c(0.05,0.05), breaks = c("5", "7", "10", "12")) +
                    scale_y_continuous(expand = c(0.05,0.05), breaks = c(0,2,4,6,8,10,12)) +
                    theme(
                          panel.grid.major.y = element_line(size = 0.2,linetype = 'dashed',colour = "gray"),
                          axis.title.x = element_text(size = 12,face="bold"),
                          axis.title.y = element_text(size = 12,face="bold"),
                          axis.text.x = element_text(colour = "black", size = 10),
                          axis.text.y = element_text(colour = "black", size = 10),
                          panel.background = element_blank(),
                          legend.position = "none") 
spaghetti_plotB
```

```{r save spag plot}
# ggsave(
#   "spaghetti_plot_three_class_colour_coded.jpeg",
#   plot = spaghetti_plotB,
#   device = "jpeg",
#   path = paste0(graph_save_data_path, "trajectories/"),
#   width = 7,
#   height = 5
# )
```
***

# Descriptives by class

The code to create these descriptives is long, so has been hidden here, please see raw code on Github if interested. The descriptives include ethnicity, SES, zygosity, and sex. 

```{r ethnicity descriptives by class, include=FALSE}
# calculate Ns and %s for each class by ethnicity
ethnicity_descriptives_stable <- dat %>% #create object to bind later on
  dplyr::count(class_renamed, ethnicity) %>%
  filter(class_renamed == "Low stable") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

ethnicity_descriptives_decreasing <- dat %>% #create object to bind later on
  dplyr::count(class_renamed, ethnicity) %>%
  filter(class_renamed == "Decreasing") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

ethnicity_descriptives_increasing <- dat %>% #create object to bind later on
  dplyr::count(class_renamed, ethnicity) %>%
  filter(class_renamed == "Increasing") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

eth_df <- list(ethnicity_descriptives_stable, ethnicity_descriptives_decreasing, ethnicity_descriptives_increasing)
ethnicity_descriptives <- join_all(eth_df,
                                   by = "ethnicity",
                                   type = "left")


# to avoid NAs in later descriptives table - change Na to 0
ethnicity_descriptives$`n_Increasing` <- ethnicity_descriptives$`n_Increasing` %>% replace_na(0)
ethnicity_descriptives$`Prop_Increasing` <- ethnicity_descriptives$`Prop_Increasing` %>% replace_na(0)

ethnicity_descriptives <- ethnicity_descriptives %>%
  mutate(Total = `n_Increasing` + `n_Decreasing` + `n_Low stable`,
         Prop_total = round(Total/2232*100, digits = 2))

# rename columns 
ethnicity_descriptives <- ethnicity_descriptives %>%
  select(
    "Descriptive" = ethnicity,
    "Low stable" = `n_Low stable`,
    "% Low stable" = `Prop_Low stable`,
    "Increasing" = `n_Increasing`,
    "% Increasing" = `Prop_Increasing`,
    "Decreasing" = `n_Decreasing`,
    "% Decreasing" = `Prop_Decreasing`,
    "Total" = Total,
    "% Total" = Prop_total)
ethnicity_descriptives
```

```{r SES descriptives by class, include=FALSE}
SES_descriptives_stable <- dat %>% 
  dplyr::count(class_renamed, SES) %>%
  filter(class_renamed == "Low stable") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

SES_descriptives_decreasing <- dat %>% 
  dplyr::count(class_renamed, SES) %>%
  filter(class_renamed == "Decreasing") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

SES_descriptives_increasing <- dat %>% 
  dplyr::count(class_renamed, SES) %>%
  filter(class_renamed == "Increasing") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

ses_df <- list(SES_descriptives_stable, SES_descriptives_decreasing, SES_descriptives_increasing)
SES_descriptives <- join_all(ses_df,
                                   by = "SES",
                                   type = "left")

SES_descriptives <- SES_descriptives %>%
  mutate(Total = `n_Increasing` + `n_Decreasing` + `n_Low stable`,
         Prop_total = round(Total/2232*100, digits = 2)) %>% 
  select(
    "Descriptive" = SES,
    "Low stable" = `n_Low stable`,
    "% Low stable" = `Prop_Low stable`,
    "Increasing" = `n_Increasing`,
    "% Increasing" = `Prop_Increasing`,
    "Decreasing" = `n_Decreasing`,
    "% Decreasing" = `Prop_Decreasing`,
    "Total" = Total,
    "% Total" = Prop_total)
SES_descriptives
```

```{r zygosity descriptives by class, include=FALSE}
zygosity_binary_descriptives_stable <- dat %>% 
  dplyr::count(class_renamed, zygosity_binary) %>%
  filter(class_renamed == "Low stable") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

zygosity_binary_descriptives_decreasing <- dat %>% 
  dplyr::count(class_renamed, zygosity_binary) %>%
  filter(class_renamed == "Decreasing") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

zygosity_binary_descriptives_increasing <- dat %>% 
  dplyr::count(class_renamed, zygosity_binary) %>%
  filter(class_renamed == "Increasing") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop)) 

zyg_df <- list(zygosity_binary_descriptives_stable, zygosity_binary_descriptives_decreasing, zygosity_binary_descriptives_increasing)
zygosity_binary_descriptives <- join_all(zyg_df,
                                   by = "zygosity_binary",
                                   type = "left")

zygosity_binary_descriptives <- zygosity_binary_descriptives %>%
  mutate(Total = `n_Increasing` + `n_Decreasing` + `n_Low stable`,
         Prop_total = round(Total/2232*100, digits = 2)) %>% 
  select(
    "Descriptive" = zygosity_binary,
    "Low stable" = `n_Low stable`,
    "% Low stable" = `Prop_Low stable`,
    "Increasing" = `n_Increasing`,
    "% Increasing" = `Prop_Increasing`,
    "Decreasing" = `n_Decreasing`,
    "% Decreasing" = `Prop_Decreasing`,
    "Total" = Total,
    "% Total" = Prop_total)
zygosity_binary_descriptives
```

```{r sex descriptives by class, include=FALSE}
sex_descriptives_stable <- dat %>% 
  dplyr::count(class_renamed, sex) %>%
  filter(class_renamed == "Low stable") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop))

sex_descriptives_decreasing <- dat %>% 
  dplyr::count(class_renamed, sex) %>%
  filter(class_renamed == "Decreasing") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop))

sex_descriptives_increasing <- dat %>% 
  dplyr::count(class_renamed, sex) %>%
  filter(class_renamed == "Increasing") %>%
  mutate(Prop = round(n/sum(n)*100, digits = 2)) %>%
  pivot_wider(names_from = class_renamed, values_from = c(n, Prop))

sex_df <- list(sex_descriptives_stable, sex_descriptives_decreasing, sex_descriptives_increasing)
sex_descriptives <- join_all(sex_df,
                                   by = "sex",
                                   type = "left")

sex_descriptives <- sex_descriptives %>%
  mutate(Total = `n_Increasing` + `n_Decreasing` + `n_Low stable`,
         Prop_total = round(Total/2232*100, digits = 2)) %>% 
  select(
    "Descriptive" = sex,
    "Low stable" = `n_Low stable`,
    "% Low stable" = `Prop_Low stable`,
    "Increasing" = `n_Increasing`,
    "% Increasing" = `Prop_Increasing`,
    "Decreasing" = `n_Decreasing`,
    "% Decreasing" = `Prop_Decreasing`,
    "Total" = Total,
    "% Total" = Prop_total)
sex_descriptives
```

```{r combine descriptives, include=FALSE}
class.descriptives <- bind_rows(
  sex_descriptives,
  zygosity_binary_descriptives,
  SES_descriptives,
  ethnicity_descriptives
)

class.descriptives <- class.descriptives %>%
  mutate(
    Category = 
      c("Sex", "",
        "Zygosity", "", 
        "SES", "", "",
        "Ethnicity", "", "", "", "")) %>%
  select(
    Category,
    Descriptive,
    `Low stable`,
    `% Low stable`,
    `Increasing`,
    `% Increasing`,
    `Decreasing`,
    `% Decreasing`,
    Total,
    `% Total`
  )
```

```{r all descriptives by class}
#final descriptives by class table
kable(class.descriptives)
```

***

# Adapt variables for regression analysis

We need to relevel the class variable to use the Low Stable class as the reference group in all subsequent regression analyses.  

```{r relevel class variable}
library(nnet)
# relevel class_factor
dat <- dat %>% 
  mutate(
    class_reordered =
      relevel(
        class_renamed,
        ref = "Low stable",
        first = TRUE, #levels in ref come first
        collapse = "+", #String used when constructing names for combined factor levels
        xlevels = TRUE #levels maintained even if not actually occurring
        
      )
  )
```
***

# Class statistics {.tabset .tabset-fade}

Next, we calculate the classification probabilities, class means and NAs per class. 

```{r read in data for all traj models, include=FALSE}
class.all <- readModels(paste0(mplus_GMM_clustered_full_output_data_path), recursive = TRUE)
```

## Classification probabilities 

For the diagonal, this represents the probability given that the an observation is a part of a specific latent class that you will classify it as said class.

*Note* this does not represent average probabilities as this table wasn't available in Mplus version 8. 

```{r extract classification probs for 3 class model}
# extract the classification probabilities for the most likely class memberships (column) by latent class (row). This shows the uncertainty rate of the probabilities. If it was perfect - the diagonal would be 1. E.g. for Class 1 - 98.3% of individuals fit that category. 
class.probs <- data.frame(class.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.output..isolation_3traj_full_sample_clustered.out$class_counts$classificationProbs.mostLikely)

# add column names
colnames(class.probs) <- c("Low stable","Increasing","Decreasing")

class.probs$Class <- c("Low stable","Increasing","Decreasing")

# as.tibble(class.probs)

class.probs <- class.probs %>%
  tidyr::pivot_longer(-Class, 
               names_to = "class", 
               values_to = "Classification probabilities")

class.probs <- class.probs[-c(2:4,6:8),-c(2)]
```

```{r classification probabilities table}
kable(class.probs)
```

## Class means

```{r extract means (TECH7 output) for each class}
# extract means for each class 
class.1.means <- as.tibble(class.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.output..isolation_3traj_full_sample_clustered.out$tech7$CLASS.1$classSampMeans)

class.2.means <- as.tibble(class.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.output..isolation_3traj_full_sample_clustered.out$tech7$CLASS.2$classSampMeans)

class.3.means <- as.tibble(class.all$X.Users.katiethompson.Documents.PhD.LISS.DTP_Louise_and_Tim.Social.isolation.trajectories_Paper.1.data_analysis.mplus.GMM.clustered.full_sample.output..isolation_3traj_full_sample_clustered.out$tech7$CLASS.3$classSampMeans)

class.means <- rbind(class.1.means,class.2.means,class.3.means)

class.means$Class <- c("Low stable","Increasing","Decreasing")

class.means <- class.means %>%
  select(
    Class,
    `Mean at age 5` = SISOE5,
    `Mean at age 7` = SISOE7,
    `Mean at age 10` = SISOE10,
    `Mean at age 12` = SISOE12
  )
```

```{r class means table}
kable(class.means)
```

## Na per person split by class

```{r na per person slipt by class}
ctable(dat$class_renamed, dat$na.per.person.si)
```


***

# Export data 

## Main analyses

```{r export new data files}
# save R data file
saveRDS(object = dat, file = paste0(data_path, "class_joined_preprocessed_isolation_data_full_sample.rds"))
```

## 3STEP method sensitivity analysis 

This 3STEP method that will be conducted in Mplus uses the ID variable, the class membership variable. 

*Remember to manually delete the headers in the csv file and replace NA to . so that Mplus can read the file*

Will need to combine the raw data file (to get numeric values and short names to the class statistics)

```{r read in raw data and combine with classes}
library(foreign)
dat.raw <- read.spss(file = paste0(data.raw_path, "Katie_18Dec20.sav"),
                  use.value.labels = FALSE, #convert value labels into factors with those labels
                  to.data.frame = TRUE) #return data frame 

#colnames(dat.raw) # check colnames

dat.tra.prob.3step <- dat.tra.prob %>%
  select(atwinid = id,
         prob1,
         prob2,
         prob3,
         class)

#colnames(dat.tra.prob.3step) # check colnames

#create list of data frames to merge
dataframe_list <- list(
  dat.raw,
  dat.tra.prob.3step
)

#merge data frames
dat.3step <- plyr::join_all(
  dataframe_list,
  by = "atwinid" # Alternatively you can join by several columns
  )

#check
#colnames(dat.3step)
```

### Variable REcode based on analysis decisions {.tabset .tabset-fade}

Variables that were created in the preprocessing script need to be recomputed here for the 3STEP sensitivity analysis.

#### ACORN recode

Moderate Means and Hard Pressed were combined, Wealthy Achievers, Urban Prosperity, and Comfortably Off were combined.

```{r ACORN recode}
dat.3step <- dat.3step %>%
  mutate(P5CACORNCategoryrecoded =
           if_else(
             P5CACORNCategory == 4 | #Moderate Means and Hard Pressed were combined, Wealthy Achievers, Urban Prosperity, and Comfortably Off were combined
               P5CACORNCategory == 5,
             1, # Deprived
             0  # Relatively affluent
           ))
#table(dat.3step$P5CACORNCategoryrecoded)
```

#### Harm recode 

Possible harm and Definite harm were combined to give a binary variable of Harm VS No harm.

```{r child harm recode, include=FALSE}
dat.3step <- dat.3step %>%
  mutate(harm3em5recoded =
           if_else(
             harm3em5 == 1 |
               harm3em5 == 2,
             1,
             0
           ))
         
#table(dat.3step$harm3em5recoded)
```

#### Highest education recode

GSCE grades were collapsed: No qualification and Level 1 (GCSE at grades D-G) were combined. Level 2 (GCSE at grades A*-C) and Level 3 (A Level) were combined. 

```{r highest education recode, include=FALSE}
dat.3step <- dat.3step %>%
  mutate(educachve18recoded =
           if_else(
             educachve18 == 2 |
               educachve18 == 3,
             1,
             0
           ))
         
#table(dat.3step$educachve18recoded)
```

#### Parental antisocial bhaviour and alcoholism

We combined maternal and paternal reports to form parental report for antisocial behaviour and alcoholism. 

```{r recode antisocial behaviour for parent}
# create total score for mum and data to make parent
dat.3step <- dat.3step %>%
  mutate(antisocialparent =
           asbmm5 + asbfm5) # mum plus dad scores
```

```{r recode alcoholism for parent}
# create total score for mum and data to make parent
dat.3step <- dat.3step %>%
  mutate(alcoholismparent =
           alcmm5 + alcfm5) # mum plus dad scores
```

## Check column names and export the data 

```{r check column names}
colnames(dat.3step)
```

```{r export files for 3step sensitivity analysis}
# save csv file 
write_csv(x = dat.3step, path = paste0(data_path, "FOR_MPLUS_preprocessed_isolation_trajectories_Sep2021_3STEP.csv")) # save data in data file
write_csv(x = dat.3step, path = "/Users/katiethompson/Desktop/FOR_MPLUS_preprocessed_isolation_trajectories_Sep2021_3STEP.csv") # save data on desktop for Mplus 
```

